\input{../../../preamble/preamble_exams.tex}

\begin{document}
\ExamTitleBlock{11th grade}{Term 3 --- Lesson 1 Practice Activity (C6) with Solutions}{}

\section*{Contents}
\addcontentsline{toc}{section}{Contents}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{1}
\tableofcontents

\ExamSection{Evaluated criteria}
\begin{ExamCriteria}
    \item C6: Interprets decisions from hypothesis tests in terms of the context of the problem.
\end{ExamCriteria}

\ExamSection{Problems}
\begin{ExamProblems}

\item
\section{Claim about average fuel inflation}
Claim: the mean monthly fuel inflation is above 4.0\%. A one-tailed test uses \(H_0\colon \mu=4.0\), \(H_1\colon \mu>4.0\), with \(\alpha=0.05\). The reported decision is \textit{reject \(H_0\)}. Interpret this decision and state what cannot be concluded.

\subsection*{C6}
\begin{itemize}
    \item The decision supports the claim that the population mean inflation is greater than 4.0\%.
    \item It does not prove with certainty that \(\mu>4.0\); there is still Type I error risk.
    \item It does not say every product category has inflation above 4.0\%; the result is about the population mean.
\end{itemize}

\item
\section{Claim about unemployment duration}
Claim: the average unemployment duration is less than 11 weeks. Test setup: \(H_0\colon \mu=11\), \(H_1\colon \mu<11\), \(\alpha=0.10\). Outcome: \textit{fail to reject \(H_0\)}. Interpret the result.

\subsection*{C6}
\begin{itemize}
    \item The data do not provide sufficient evidence to support the claim that mean duration is below 11 weeks.
    \item This does not prove \(\mu=11\); it only means evidence was not strong enough at the 10\% level.
    \item Some individuals may still be unemployed for fewer than 11 weeks; the test concerns the population mean.
\end{itemize}

\item
\section{Claim about export defect proportion}
Claim: fewer than 5\% of exported parts are defective. Setup: \(H_0\colon p=0.05\), \(H_1\colon p<0.05\), \(\alpha=0.01\). Outcome: \textit{fail to reject \(H_0\)}. Explain the interpretation.

\subsection*{C6}
\begin{itemize}
    \item There is insufficient evidence at 1\% to conclude that the defect proportion is below 5\%.
    \item This does not prove the defect proportion is exactly 5\%.
    \item A Type II error is possible: the true defect rate could be below 5\%, but the test did not detect it.
\end{itemize}

\item
\section{Claim about household digital payment use}
Claim: more than 70\% of households use digital payments weekly. Setup: \(H_0\colon p=0.70\), \(H_1\colon p>0.70\), \(\alpha=0.05\). Outcome: \textit{reject \(H_0\)}. Interpret carefully.

\subsection*{C6}
\begin{itemize}
    \item The decision indicates sufficient evidence that the population proportion exceeds 70\%.
    \item It does not guarantee every household uses digital payments weekly.
    \item The conclusion is probabilistic, not absolute certainty.
\end{itemize}

\item
\section{Claim about average customs wait time}
Claim: mean customs wait time is different from 2.5 hours. Setup: \(H_0\colon \mu=2.5\), \(H_1\colon \mu\ne 2.5\), \(\alpha=0.05\). Outcome: \textit{reject \(H_0\)}. What does this mean and not mean?

\subsection*{C6}
\begin{itemize}
    \item Evidence supports that the population mean wait time is not 2.5 hours.
    \item The result does not identify direction unless sample information is checked separately.
    \item It does not imply all shipments wait longer or shorter than 2.5 hours.
\end{itemize}

\item
\section{Claim about school meal participation}
Claim: at least 62\% of students join the school meal program. Setup: \(H_0\colon p=0.62\), \(H_1\colon p<0.62\), \(\alpha=0.10\). Outcome: \textit{reject \(H_0\)}. Provide interpretation.

\subsection*{C6}
\begin{itemize}
    \item The data provide sufficient evidence that participation is below 62\%.
    \item This does not prove the participation rate for each school is below 62\%.
    \item There remains a chance of Type I error due to using \(\alpha=0.10\).
\end{itemize}

\item
\section{Claim about mean wage growth}
Claim: average annual wage growth equals 3.5\%. Setup: \(H_0\colon \mu=3.5\), \(H_1\colon \mu\ne 3.5\), \(\alpha=0.01\). Outcome: \textit{fail to reject \(H_0\)}. Interpret the decision.

\subsection*{C6}
\begin{itemize}
    \item There is not enough evidence at 1\% to conclude wage growth differs from 3.5\%.
    \item This does not prove the mean is exactly 3.5\%.
    \item A non-significant result can happen when sample evidence is weak or variability is high.
\end{itemize}

\item
\section{Claim about inflation-awareness training success}
Claim: more than 80\% of trainees pass an inflation-awareness certification. Setup: \(H_0\colon p=0.80\), \(H_1\colon p>0.80\), \(\alpha=0.05\). Outcome: \textit{fail to reject \(H_0\)}. Explain in context.

\subsection*{C6}
\begin{itemize}
    \item The evidence is insufficient to support that pass proportion is above 80\%.
    \item This does not prove the pass rate is 80\% or below 80\%.
    \item The conclusion refers to the population proportion, not a guarantee for each cohort.
\end{itemize}

\item
\section{Claim about average farm loan size}
Claim: the mean approved farm loan is less than 18{,}000 dollars. Setup: \(H_0\colon \mu=18000\), \(H_1\colon \mu<18000\), \(\alpha=0.05\). Outcome: \textit{reject \(H_0\)}. Interpret this conclusion.

\subsection*{C6}
\begin{itemize}
    \item There is sufficient evidence that the population mean approved loan is below 18{,}000 dollars.
    \item It does not imply every loan is below 18{,}000 dollars.
    \item The claim is supported statistically, but not proven with absolute certainty.
\end{itemize}

\item
\section{Claim about broadband access proportion}
Claim: the proportion of households with broadband is 76\%. Setup: \(H_0\colon p=0.76\), \(H_1\colon p\ne 0.76\), \(\alpha=0.05\). Outcome: \textit{fail to reject \(H_0\)}. State the interpretation and limitation.

\subsection*{C6}
\begin{itemize}
    \item At the 5\% level, evidence is insufficient to conclude the broadband proportion differs from 76\%.
    \item The decision does not prove the true proportion equals 76\%.
    \item A meaningful difference may still exist but remain undetected (possible Type II error).
\end{itemize}

\end{ExamProblems}
\end{document}
