\makeatletter
\def\input@path{{./}{../}{../../}{../../../}{../../../../}{preamble/}{../preamble/}{../../preamble/}{../../../preamble/}}
\makeatother

\input{preamble.tex}

\title{Grade 10 - Term 3 \\ Learning Evidence 4 Activity (C10)}
\author{}
\date{}

\begin{document}
	\maketitle
	
	\section*{Learning Objective and Assessed Criterion}
	\textbf{Learning objective.} Use an interactive webpage to model real or simulated data with continuous probability distributions, compare the fit of different PDFs, calculate probabilities from each model, and communicate conclusions in a structured report.
	
	\vspace{0.2cm}
	\textbf{Assessed criterion.}
	\begin{itemize}
		\item \textbf{C10.} Uses continuous probability models to analyze data, compare fitted distributions, calculate probabilities, and write evidence-based conclusions.
	\end{itemize}

	\section{Background from C1--C9}
	In C1--C4, you modeled continuous probabilities with uniform, linear, triangular, and piecewise density functions. In C5--C7, you studied the normal curve and calculated probabilities with normal distributions. In C8, you created an interactive webpage to graph continuous PDFs and calculate areas under the curve. In C9, you used AI support to build a script that estimates parameters from data and compares fitted PDFs with histograms.
	
	In C10, you will use those same tools to model complete datasets and evaluate which probability model gives the most reasonable fit.

	\section{Task Overview}
	You will analyze \textbf{three datasets}. For each dataset, you must fit \textbf{three models} using your C8--C9 webpage:
	\begin{enumerate}
		\item Piecewise uniform distribution
		\item Linear distribution
		\item Normal distribution
	\end{enumerate}
	
	Then you will compute required probabilities from each fitted model and write a structured comparative report.
	
	\textbf{Total required work:} 9 model fits and 18 probability calculations.

	\section{Required Datasets}
	Use the following CSV files from this folder:
	\begin{enumerate}
		\item \texttt{G10\_T3\_L4\_C10\_dataset\_1.csv}
		\item \texttt{G10\_T3\_L4\_C10\_dataset\_2.csv}
		\item \texttt{G10\_T3\_L4\_C10\_dataset\_3.csv}
	\end{enumerate}
	
	\textbf{Important note:} These datasets are simulated for educational purposes, but they are designed to represent realistic contexts.
	
	Each dataset includes:
	\begin{itemize}
		\item a context/group identifier,
		\item one numerical variable to model,
		\item enough observations to build a meaningful histogram.
	\end{itemize}

	\section{Modeling Requirements}
	For \textbf{each dataset}, complete all steps below:
	\begin{enumerate}
		\item Upload the dataset into your C8--C9 webpage.
		\item Display the dataset histogram.
		\item Fit a piecewise uniform model and plot its PDF on the same graph as the histogram.
		\item Fit a linear model and plot its PDF on the same graph as the histogram.
		\item Fit a normal model and plot its PDF on the same graph as the histogram.
		\item Compare visually which model appears to fit better.
		\item Explain one strength and one limitation of each model for that dataset.
	\end{enumerate}

	\section{Probability-Calculation Requirements}
	For \textbf{each dataset} and \textbf{each fitted model}, calculate:
	\[
	P(\mu-\sigma < X < \mu+\sigma)
	\]
	and
	\[
	P(X<\mu-2\sigma\;\text{or}\;X>\mu+2\sigma).
	\]
	
	Use probabilities from the \textbf{fitted model} (area under the model PDF), not only point-counting from raw data.
	
	You must complete all combinations:
	\begin{itemize}
		\item Dataset 1 with piecewise uniform, linear, and normal models.
		\item Dataset 2 with piecewise uniform, linear, and normal models.
		\item Dataset 3 with piecewise uniform, linear, and normal models.
	\end{itemize}

	\section{Report Requirements}
	Write a structured report with these sections:
	\begin{enumerate}
		\item Title
		\item Introduction
		\item Description of the three datasets
		\item Methodology
		\item Model fitting results
		\item Probability calculations
		\item Comparison of models
		\item Discussion
		\item Conclusion
		\item Appendix (screenshots and datasets)
	\end{enumerate}
	
	For \textbf{each dataset}, include:
	\begin{enumerate}
		\item screenshot of histogram,
		\item screenshot of piecewise uniform fit,
		\item screenshot of linear fit,
		\item screenshot of normal fit,
		\item estimated parameters for all three models,
		\item both required probabilities for all three models,
		\item a short evidence-based statement about which model fits better and why.
	\end{enumerate}

	\section{AI-Use Requirements}
	You may use AI to improve your script or support interpretation, but you must document responsible use.
	
	Include in your report:
	\begin{enumerate}
		\item AI prompt(s) used,
		\item what the AI helped with,
		\item what you verified or corrected,
		\item a statement that final interpretation is your responsibility.
	\end{enumerate}

	\section{Deliverables}
	Submit all items below:
	\begin{enumerate}
		\item Report (PDF)
		\item Editable report source file
		\item Three CSV datasets
		\item Screenshots of all model fits
		\item Webpage files used for the analysis
		\item Short AI-use reflection
	\end{enumerate}

	\section{Evaluation Checklist}
	Your evidence is complete only if all statements are true:
	\begin{itemize}
		\item[\(\square\)] I analyzed all three datasets.
		\item[\(\square\)] I modeled each dataset with piecewise uniform, linear, and normal PDFs.
		\item[\(\square\)] I showed each fitted PDF on the same graph as the histogram.
		\item[\(\square\)] I calculated \(P(\mu-\sigma<X<\mu+\sigma)\) for every model.
		\item[\(\square\)] I calculated \(P(X<\mu-2\sigma\;\text{or}\;X>\mu+2\sigma)\) for every model.
		\item[\(\square\)] I compared models using graph evidence and probability results.
		\item[\(\square\)] My conclusions are consistent with my data and fitted models.
	\end{itemize}

	\section{Submission Instructions}
	\begin{enumerate}
		\item Organize files in clearly named folders (report, datasets, screenshots, webpage, AI-reflection).
		\item Verify that your webpage and scripts run correctly before submission.
		\item Ensure all screenshots are readable and labeled by dataset and model.
		\item Submit all required files by the teacher's deadline.
	\end{enumerate}
	
	\vspace{0.2cm}
	\textbf{Reminder:} This is an activity sheet. Do not submit pre-solved answers or copied conclusions. Show your own modeling process and justified interpretation.

\end{document}
