\input{../../preamble/preamble_exams.tex}
\begin{document}
	\ExamTitleBlock{11th grade}{Practice Activity 2.1: C6 Practice}{}
	
	\ExamSection{Evaluated criteria}
	\begin{ExamCriteria}
		\item C6: Determine the required sample size to achieve a target margin of error.
	\end{ExamCriteria}
	
	\ExamSection{Problems}
	\begin{ExamProblems}
		\item
		\subsection*{Problem description}
		A regional logistics agency wants to estimate the mean time (in minutes) for cargo to clear port inspections. Historical monitoring shows the population standard deviation is $\sigma = 12$ minutes. The agency wants a 95\% confidence estimate with a maximum error of $E = 3$ minutes. Determine the minimum sample size needed.
		
		\subsection*{C6}
		This solution establishes the sample size so the error will not exceed a specified amount by identifying $E$, selecting $z^*$, substituting into the formula, and rounding up to the minimum integer sample size.
		
		Step 1: Identify the target error and confidence level. The maximum error is $E = 3$ minutes and the confidence level is 95\%, so $z^* = 1.96$.
		
		Step 2: Apply the sample size formula.
		\[
		n = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{1.96 \cdot 12}{3}\right)^2
		= \left(7.84\right)^2
		= 61.47
		\]
		
		Step 3: Round up and verify Criterion 6. Because the sample size must ensure the error does not exceed $E$, we round up: $n = 62$. With $n = 62$, the margin of error is at most 3 minutes, satisfying Criterion 6.

		\item
		\subsection*{Problem description}
		A development agency wants to estimate the mean monthly food spending for households in a region. A pilot survey suggests the population standard deviation is about $\sigma \approx 45$ currency units. The agency wants a 90\% confidence estimate with a target interval width of $W = 16$ currency units. The width is defined as $W = \text{Upper} - \text{Lower} = 2E$. Determine the minimum sample size needed.
		
		\subsection*{C6}
		This solution establishes the sample size so the interval width will not exceed a specified amount by converting $W$ to $E$, selecting the correct $z^*$, substituting into the formula, and rounding up to the minimum safe integer.
		
		Step 1: Identify the target width and confidence level. The target width is $W = 16$, so $E = W/2 = 8$. The confidence level is 90\%, so $z^* = 1.645$.
		
		Step 2: Apply the sample size formula.
		\[
		n = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{1.645 \cdot 45}{8}\right)^2
		= \left(9.253\right)^2
		\approx 85.62
		\]
		
		Step 3: Round up and verify Criterion 6. We must round up to keep the width within $W$, so $n = 86$. This minimum sample size guarantees the interval width is no more than 16 currency units.

		\item
		\subsection*{Problem description}
		A shipping cooperative wants to estimate the mean fuel cost per container on a trade route. Company records indicate the population standard deviation is $\sigma = 18$ dollars. The cooperative wants the maximum margin of error to be no more than $E = 5$ dollars. Determine the minimum sample size needed for 90\%, 95\%, and 98\% confidence.
		
		\subsection*{C6}
		This solution calculates the minimum sample size needed to keep the margin of error at or below $E = 5$ dollars for three confidence levels.
		
		Step 1: Identify the confidence levels and $z^*$ values. Use $z^* = 1.645$ for 90\%, $z^* = 1.96$ for 95\%, and $z^* = 2.326$ for 98\% confidence.
		
		Step 2: Apply the sample size formula for each confidence level.
		\[
		\text{90\%: } n = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{1.645 \cdot 18}{5}\right)^2
		= \left(5.922\right)^2
		\approx 35.07
		\]
		\[
		\text{95\%: } n = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{1.96 \cdot 18}{5}\right)^2
		= \left(7.056\right)^2
		\approx 49.79
		\]
		\[
		\text{98\%: } n = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{2.326 \cdot 18}{5}\right)^2
		= \left(8.374\right)^2
		\approx 70.12
		\]
		
		Step 3: Round up to ensure the error does not exceed 5 dollars. The required sample sizes are $n = 36$ for 90\%, $n = 50$ for 95\%, and $n = 71$ for 98\% confidence.

		\item
		\subsection*{Problem description}
		A trade corridor study tracks the mean daily transit time (in minutes) for cross-border trucks. The population standard deviation is known to be $\sigma = 12$ minutes. The planners will use a 95\% confidence level. Determine the minimum sample size needed for each target margin of error: $E = 3$ minutes, $E = 4$ minutes, and $E = 5$ minutes.
		
		\subsection*{C6}
		This solution determines the minimum sample size for a fixed 95\% confidence level with three different target margins of error.
		
		Step 1: Identify the confidence level and $z^*$. For 95\% confidence, $z^* = 1.96$.
		
		Step 2: Apply the sample size formula for each margin of error.
		\[
		\text{For } E = 3: \quad n = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{1.96 \cdot 12}{3}\right)^2
		= \left(7.84\right)^2
		= 61.47
		\]
		\[
		\text{For } E = 4: \quad n = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{1.96 \cdot 12}{4}\right)^2
		= \left(5.88\right)^2
		= 34.57
		\]
		\[
		\text{For } E = 5: \quad n = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{1.96 \cdot 12}{5}\right)^2
		= \left(4.704\right)^2
		= 22.13
		\]
		
		Step 3: Round up to ensure each error target is met. The minimum sample sizes are $n = 62$ for $E = 3$, $n = 35$ for $E = 4$, and $n = 23$ for $E = 5$.

		\item
		\subsection*{Problem description}
		A regional logistics agency is updating its estimate of the mean time (in minutes) for cargo to clear port inspections. Historical monitoring shows the population standard deviation is $\sigma = 10$ minutes. The current study already includes $n_{\text{current}} = 40$ observations, with sample summaries $\bar{x} = 47.8$ minutes and $s = 10.9$ minutes (for context only). Because $\sigma$ is known, planning must use $\sigma$. The agency wants a 95\% confidence estimate with target margin of error $E = 2.5$ minutes. How many additional observations are required?
		
		\subsection*{C6}
		Step 1: Identify $E$ and $z^*$ from the confidence level. The target margin is $E = 2.5$ minutes. For 95\% confidence, $z^* = 1.96$.
		
		Step 2: Compute $n_{\text{required}}$ using the formula.
		\[
		n_{\text{required}} = \left(\frac{z^*\,\sigma}{E}\right)^2
		= \left(\frac{1.96 \cdot 10}{2.5}\right)^2
		= \left(7.84\right)^2
		= 61.47
		\]
		
		Step 3: Round up $n_{\text{required}}$ to the next integer. $n_{\text{required}} = 62$.
		
		Step 4: Compute additional observations. $n_{\text{required}} - n_{\text{current}} = 62 - 40 = 22$.
		
		Conclusion: The agency needs $22$ additional observations.

		\item
		\subsection*{Problem description}
		An economic survey team is refining an estimate of mean monthly household food spending (in currency units). Prior studies show the population standard deviation is $\sigma = 40$ units. The current dataset has $n_{\text{current}} = 95$ households, with sample summaries $\bar{x} = 512$ and $s = 42$ (reported for context only). Because $\sigma$ is known, the planning calculation uses $\sigma$. For 90\% confidence, the team wants the confidence interval width to be at most $W_{\text{target}} = 12$ units. How many additional observations are required?
		
		\subsection*{C6}
		Step 1: Identify the target width and $z^*$ from the confidence level. The target width is $W_{\text{target}} = 12$ units. For 90\% confidence, $z^* = 1.645$.
		
		Step 2: Convert width to the target margin of error.
		\[
		E_{\text{target}} = \frac{W_{\text{target}}}{2}
		= \frac{12}{2}
		= 6
		\]
		
		Step 3: Compute $n_{\text{required}}$ using the sample size formula.
		\[
		n_{\text{required}} = \left(\frac{z^*\,\sigma}{E_{\text{target}}}\right)^2
		= \left(\frac{1.645 \cdot 40}{6}\right)^2
		= \left(10.967\right)^2
		= 120.27
		\]
		
		Step 4: Round up $n_{\text{required}}$ to the next integer. $n_{\text{required}} = 121$.
		
		Step 5: Compute additional observations. $n_{\text{required}} - n_{\text{current}} = 121 - 95 = 26$.
		
		Conclusion: The survey needs $26$ additional household observations to keep the interval width at or below 12 units.

		\item
		\subsection*{Problem description}
		A shipping cooperative is extending a study of mean fuel cost per container on a trade route. Accounting records indicate the population standard deviation is $\sigma = 22$ dollars, and the team has already collected $n_{\text{current}} = 100$ observations. The cooperative wants the maximum margin of error to be no more than $E = 4$ dollars. For 90\%, 95\%, and 98\% confidence, determine how many additional observations are required.
		
		\subsection*{C6}
		Step 1: Identify the fixed target margin, current sample size, and the $z^*$ values. Use $E = 4$ dollars, $n_{\text{current}} = 100$, with $z^* = 1.645$ for 90\%, $z^* = 1.96$ for 95\%, and $z^* = 2.326$ for 98\% confidence.
		
		Step 2: Use the sample size formula to find $n_{\text{required}}$ for each confidence-level case.
		\[
		n_{\text{required}} = \left(\frac{z^*\,\sigma}{E}\right)^2
		\]
		\[
		\text{90\%: } n_{\text{required}} = \left(\frac{1.645 \cdot 22}{4}\right)^2
		= \left(9.048\right)^2
		\approx 81.86 \Rightarrow 82
		\]
		\[
		\text{95\%: } n_{\text{required}} = \left(\frac{1.96 \cdot 22}{4}\right)^2
		= \left(10.78\right)^2
		\approx 116.21 \Rightarrow 117
		\]
		\[
		\text{98\%: } n_{\text{required}} = \left(\frac{2.326 \cdot 22}{4}\right)^2
		= \left(12.793\right)^2
		\approx 163.67 \Rightarrow 164
		\]
		
		Step 3: Compute additional observations for each case.
		\[
		n_{\text{additional}} = n_{\text{required}} - n_{\text{current}}
		\]
		\[
		\text{90\%: } n_{\text{additional}} = 82 - 100 = -18 \Rightarrow 0
		\]
		At 90\% confidence, no additional observations are required.
		\[
		\text{95\%: } n_{\text{additional}} = 117 - 100 = 17
		\]
		\[
		\text{98\%: } n_{\text{additional}} = 164 - 100 = 64
		\]
		Additional observations needed: $0$ (90\%), $17$ (95\%), and $64$ (98\%).

		\item
		\subsection*{Problem description}
		A trade corridor team is updating its estimate of mean daily cross-border transit time (in minutes) for trucks. Historical analysis gives a known population standard deviation of $\sigma = 15$ minutes, and the team has already collected $n_{\text{current}} = 80$ observations. At 95\% confidence, determine how many additional observations are required for each target margin of error: $E_1 = 2.5$ minutes, $E_2 = 3.5$ minutes, and $E_3 = 4.5$ minutes.
		
		\subsection*{C6}
		Step 1: Identify the confidence level, current sample size, and $z^*$. For 95\% confidence, $z^* = 1.96$, with $n_{\text{current}} = 80$.
		
		Step 2: Use the sample size formula to find $n_{\text{required}}$ for each margin-of-error case.
		\[
		n_{\text{required}} = \left(\frac{z^*\,\sigma}{E}\right)^2
		\]
		\[
		E_1 = 2.5: \quad n_{\text{required}} = \left(\frac{1.96 \cdot 15}{2.5}\right)^2
		= \left(11.76\right)^2
		\approx 138.30 \Rightarrow 139
		\]
		\[
		E_2 = 3.5: \quad n_{\text{required}} = \left(\frac{1.96 \cdot 15}{3.5}\right)^2
		= \left(8.4\right)^2
		= 70.56 \Rightarrow 71
		\]
		\[
		E_3 = 4.5: \quad n_{\text{required}} = \left(\frac{1.96 \cdot 15}{4.5}\right)^2
		= \left(6.533\right)^2
		\approx 42.68 \Rightarrow 43
		\]
		
		Step 3: Compute additional observations for each margin target.
		\[
		n_{\text{additional}} = n_{\text{required}} - n_{\text{current}}
		\]
		\[
		E_1: \quad n_{\text{additional}} = 139 - 80 = 59
		\]
		\[
		E_2: \quad n_{\text{additional}} = 71 - 80 = -9 \Rightarrow 0
		\]
		For $E_2$, no additional observations are required.
		\[
		E_3: \quad n_{\text{additional}} = 43 - 80 = -37 \Rightarrow 0
		\]
		For $E_3$, no additional observations are required.
		Additional observations needed: $59$ for $E_1 = 2.5$, $0$ for $E_2 = 3.5$, and $0$ for $E_3 = 4.5$.
		
	\end{ExamProblems}
\end{document}
