\input{../../preamble/preamble_exams.tex}
\begin{document}
	\ExamTitleBlock{11th grade}{Learning evidence 2.1: Confidence Interval Planning (Practice Solutions)}{}
	
	\ExamSection{Evaluated criteria}
	\begin{ExamCriteria}
		\item C1: Compute the sample mean and sample standard deviation.
		\item C2: Distinguish between sample statistics and population parameters.
		\item C3: Explain why interval estimation is preferred over a single point estimate.
		\item C5: Construct a X\% confidence interval using known population standard deviation.
		\item C4: Interpret the meaning of a X\% confidence interval in context.
	\end{ExamCriteria}
	
	\ExamSection{Problems}
	\begin{ExamProblems}
		\item
		\subsection*{Problem description}
		A distribution center monitors the time (in minutes) required to load outbound pallets.
		A sample of 75 loading times is grouped into class intervals below. The center has a historical estimate of the population standard deviation, \(\sigma = 6.0\) minutes.
		For classroom reference, suppose the true population mean loading time is \(\mu = 33\) minutes.
		The grouped intervals and frequencies are: 18--22 minutes (10), 23--27 minutes (16), 28--32 minutes (19), 33--37 minutes (17), and 38--42 minutes (13).
		Use the grouped data to estimate the sample mean and sample standard deviation.
		Construct and interpret a 95\% confidence interval for the population mean loading time.
		Construct and interpret a 98\% confidence interval for the population mean loading time.

		\subsection*{C1}
		Using grouped midpoints $x$ with frequencies $f$. \\
		First, we calculate the sample mean $\overline{x}$
		\[
		n = \sum f \qquad \longrightarrow \qquad n = 10 + 16 + 19 + 17 + 13 = 75
		\]
		\[
		\bar{x} = \frac{\sum f x}{n} \qquad \longrightarrow \qquad \sum f x = 10(20) + 16(25) + 19(30) + 17(35) + 13(40) = 2285 \qquad \longrightarrow \qquad \bar{x} = \frac{2285}{75} \approx 30.47
		\]
		Second, we calculate the sample standard deviation $s$ or the sample variance $s^2$
		\[
		s^2 = \frac{\sum f(x-\bar{x})^2}{n-1} \quad \longrightarrow \quad \sum f(x-\bar{x})^2 \approx 10(20-30.47)^2 + 16(25-30.47)^2 + 19(30-30.47)^2 + 17(35-30.47)^2 + 13(40-30.47)^2
		\]
		\[
		\sum f(x-\bar{x})^2 \approx 1095.51 + 478.15 + 4.14 + 349.37 + 1181.50 = 3108.67 \quad \longrightarrow \quad s = \sqrt{\frac{\sum f(x-\bar{x})^2}{n-1}} = \sqrt{\frac{3108.67}{75-1}} \approx 6.48
		\]

		\subsection*{C2}
		The sample statistics are $\bar{x} \approx 30.47$ and $s \approx 6.48$, while the population parameters are $\mu$ (the true mean) and $\sigma = 6.0$ minutes. The sample values estimate, but do not equal, the population values.

		\subsection*{C3}
		A single point estimate like $\bar{x}$ ignores sampling variability. A confidence interval gives a range of plausible values for $\mu$, which is more informative for planning in a real setting.

		\subsection*{C5}
		With known $\sigma = 6.0$ and $n=75$, the standard error is
		\[
		SE = \frac{\sigma}{\sqrt{n}} \qquad \longrightarrow \qquad SE = \frac{6.0}{\sqrt{75}} \approx 0.69
		\]
		For 95\% confidence, $z_{0.025}=1.96$.
		\[
		E = z_{0.025} \cdot SE \qquad \longrightarrow \qquad E = 1.96(0.69) \approx 1.36
		\]
		\[
		\mu \in \bar{x} \pm E \qquad \longrightarrow \qquad \mu \in 30.47 \pm 1.36 = [30.47 - 1.36, 30.47 + 1.36] = [29.11, 31.82]
		\]
		For 98\% confidence, $z_{0.01}=2.33$.
		\[
		E = z_{0.01} \cdot SE \qquad \longrightarrow \qquad E = 2.33(0.69) \approx 1.61
		\]
		\[
		\mu \in \bar{x} \pm E \qquad \longrightarrow \qquad \mu \in 30.47 \pm 1.61 = [30.47 - 1.61, 30.47 + 1.61] = [28.85, 32.08]
		\]

		\subsection*{C4}
		We are 95\% confident the true mean loading time is between about 29.11 and 31.82 minutes, and 98\% confident it is between about 28.85 and 32.08 minutes. The 98\% interval is wider because higher confidence requires a larger margin of error.

		\item
		\subsection*{Problem description}
		A school cafeteria tracks the number of sandwiches sold during lunch to plan inventory.
		A random sample of 20 lunch periods is recorded below. The district provides a historical population standard deviation of \(\sigma = 5.5\) sandwiches.

		\begin{itemize}
			\item Sample (\(n = 20\)): 42, 38, 45, 47, 40, 44, 39, 41, 46, 43, 37, 48, 40, 45, 42, 44, 39, 46, 41, 43.
		\end{itemize}

		Use the sample to compute the mean and sample standard deviation.
		Construct and interpret a 95\% confidence interval for the population mean sandwiches sold.
		Construct and interpret a 98\% confidence interval for the population mean sandwiches sold.

		\subsection*{C1}
		First, we calculate the sample mean $\overline{x}$
		\[
		\bar{x} = \frac{\sum x}{n} \quad \longrightarrow \quad \sum x = 42 + 38 + 45 + 47 + 40 + 44 + 39 + 41 + 46 + 43 + 37 + 48 + 40 + 45 + 42 + 44 + 39 + 46 + 41 + 43 = 850 \] 
		\[\quad \longrightarrow \quad \bar{x} = \frac{850}{20} = 42.5
		\]
		Second, we calculate the sample standard deviation $s$ or the sample variance $s^2$
		\[
		\begin{aligned}
			s^2 &= \frac{\sum (x-\bar{x})^2}{n-1}
			\quad \longrightarrow \quad
			\sum (x-\bar{x})^2 = {}\\
			&\; 2(42-42.5)^2 + 2(40-42.5)^2 + 2(45-42.5)^2 + 2(44-42.5)^2 \\
			&\; + 2(39-42.5)^2 + 2(46-42.5)^2 + 2(41-42.5)^2 + 2(43-42.5)^2 \\
			&\; + (38-42.5)^2 + (47-42.5)^2 + (37-42.5)^2 + (48-42.5)^2 = 185
		\end{aligned}
		\]

		\[
		s = \sqrt{\frac{\sum (x-\bar{x})^2}{n-1}} \qquad \longrightarrow \qquad s = \sqrt{\frac{185}{20-1}} = \sqrt{\frac{185}{19}} \approx 3.12
		\]

		\subsection*{C2}
		The sample statistics are $\bar{x} = 42.5$ and $s \approx 3.12$, while the population mean $\mu$ and population standard deviation $\sigma = 5.5$ are parameters for all lunch periods. The sample measures are estimates of the population values.

		\subsection*{C3}
		Because only 20 lunch periods are observed, a point estimate may be off. Confidence intervals quantify that uncertainty and show a plausible range for $\mu$ instead of a single guess.

		\subsection*{C5}
		With known $\sigma = 5.5$ and $n=20$,
		\[
		SE = \frac{\sigma}{\sqrt{n}} \qquad \longrightarrow \qquad SE = \frac{5.5}{\sqrt{20}} \approx 1.23
		\]
		For 95\% confidence, $z_{0.025}=1.96$.
		\[
		E = z_{0.025} \cdot SE \qquad \longrightarrow \qquad E = 1.96(1.23) \approx 2.41
		\]
		\[
		\mu \in \bar{x} \pm E \qquad \longrightarrow \qquad \mu \in 42.5 \pm 2.41 = [42.5 - 2.41, 42.5 + 2.41] = [40.09, 44.91]
		\]
		For 98\% confidence, $z_{0.01}=2.33$.
		\[
		E = z_{0.01} \cdot SE \qquad \longrightarrow \qquad E = 2.33(1.23) \approx 2.87
		\]
		\[
		\mu \in \bar{x} \pm E \qquad \longrightarrow \qquad \mu \in 42.5 \pm 2.87 = [42.5 - 2.87, 42.5 + 2.87] = [39.63, 45.37]
		\]

		\subsection*{C4}
		We are 95\% confident the true mean sandwiches sold per lunch period is between about 40.09 and 44.91, and 98\% confident it is between about 39.63 and 45.37. The wider 98\% interval offers more assurance at the cost of precision.
	\end{ExamProblems}
\end{document}
