\input{../../../preamble/preamble_exams.tex}

\begin{document}
\ExamTitleBlock{11th grade}{Term 3 --- Lesson 1 Summary of Practice Activities}{ }

\section*{Contents}
\addcontentsline{toc}{section}{Contents}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{1}
\tableofcontents

\section{Summary of \texttt{G11\_T3\_L1\_C1C2C3\_practice\_activity1.tex}}
This activity develops foundational work for Criteria C1--C3 by helping students identify whether a context uses a population mean or proportion, assign correct symbols, and write hypotheses in all three possible directions (lower, higher, different). Its purpose is introductory: students first build a stable template for parameter notation and then connect each alternative-hypothesis direction to left-, right-, or two-tailed testing. The reasoning focus is classification and structure recognition before full procedural testing.

\begin{itemize}
    \item \textbf{Problem 1:} Introductory proportion problem where students map context values to \(p, p_0, n, x, \alpha\), then write all three directional hypothesis pairs and classify the corresponding tail types.
    \item \textbf{Problem 2:} Parallel mean-based setup (with known \(\sigma\)) that reinforces the same three-direction framework, now transferring symbolic reasoning from proportions to means.
    \item \textbf{Problem 3:} Proportion-focused quality-control context that strengthens directional hypothesis writing under a stricter significance level, emphasizing precision in claim wording.
    \item \textbf{Problem 4:} Mean-context consolidation problem in which students repeat the full C1--C3 cycle and verify that tail classification is determined by the alternative-hypothesis direction, not by topic context.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C1C2C3\_practice\_activity2.tex}}
This redesigned follow-up activity structures C1--C3 work as a clear progression from contextual direction selection to evidence-based conclusion making. Students first identify the correct one-direction formulation from context and complete full hypothesis-test decisions using reject \(H_0\) or fail to reject \(H_0\). They then examine how conclusions can change when significance levels change while hypotheses and tail direction remain fixed. The activity ends with paired comparisons on identical data to show explicitly how choosing a one-sided versus two-sided alternative can change p-values and decisions.

\begin{itemize}
    \item \textbf{Problem 1:} Lower-tailed mean context where students identify \(\mu\), write \(H_0: \mu=\mu_0\) and \(H_a: \mu<\mu_0\), compute test evidence, and reach a reject-\(H_0\) decision.
    \item \textbf{Problem 2:} Higher-tailed proportion context where students map \(p\)-notation, test \(H_a: p>p_0\), and obtain a reject-\(H_0\) decision from the sample evidence.
    \item \textbf{Problem 3:} Two-tailed mean context for a ``different'' claim, producing a fail-to-reject-\(H_0\) outcome so students see contrasting decisions across Stage 1.
    \item \textbf{Problem 4:} Lower-tailed mean test where hypotheses and tail stay fixed, but conclusions are compared across multiple \(\alpha\) levels motivated by routine monitoring versus stricter accountability.
    \item \textbf{Problem 5:} Higher-tailed proportion test where the same evidence is interpreted under internal-planning and public-reporting significance standards, showing threshold-sensitive decisions.
    \item \textbf{Problem 6:} Two-tailed mean quality-control test that keeps the ``different'' formulation fixed while contrasting conclusions across looser and stricter quality thresholds.
    \item \textbf{Problem 7 (same-evidence lower vs different):} Same context, same data, and same \(\alpha\) comparison of left-tailed and two-tailed formulations to highlight how alternative-hypothesis framing changes p-values and can affect rejection strength.
    \item \textbf{Problem 8 (same-evidence higher vs different):} Same-data comparison of right-tailed and two-tailed formulations at one significance level where both tests fail to reject \(H_0\), reinforcing meaningful non-significance interpretation under different alternatives.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C4\_practice\_activity.tex}}
This activity is now the dedicated C4 ladder for one-sample proportion hypothesis testing. It develops a full proportion-testing workflow from foundational setup to direction and significance-level sensitivity, then to same-evidence comparative framing, and finally to synthesis cases that classify robust, threshold-sensitive, and near-boundary outcomes using reject-\(H_0\)/fail-to-reject-\(H_0\) language in context.

\begin{itemize}
    \item \textbf{General Problem (Stage 1 --- proportion-only full procedure):} Establishes parameter identification \((p)\), benchmark setup \((p_0)\), sample proportion \((\hat p)\), condition checks, one-sample proportion \(z\)-statistic, p-value by tail type, decision rule, and contextual conclusion template.
    \item \textbf{Problem 1 (Stage 1 --- introductory lower-tail proportion test):} First direct lower-tail proportion application with full step-by-step computation and conclusion.
    \item \textbf{Problem 2 (Stage 1 --- introductory higher-tail proportion test):} First direct higher-tail proportion application reinforcing the same full workflow.
    \item \textbf{Problem 3 (Stage 1 --- introductory two-tailed proportion test):} First direct two-tailed proportion application for ``different'' claims.
    \item \textbf{Problem 4 (Stage 2 --- lower-tail \(\alpha\)-sensitivity):} Holds sample evidence fixed and compares decisions across 5\% and 1\% thresholds.
    \item \textbf{Problem 5 (Stage 2 --- higher-tail \(\alpha\)-sensitivity):} Demonstrates threshold-sensitive decisions from the same higher-tail proportion evidence.
    \item \textbf{Problem 6 (Stage 2 --- two-tailed \(\alpha\)-sensitivity):} Two-tailed proportion case where decision changes with significance level.
    \item \textbf{Problem 7 (Stage 3 --- lower vs higher directional comparison):} Same data compared under opposite one-tailed alternatives to show tail-dependent p-values and contrasting decisions.
    \item \textbf{Problem 8 (Stage 3 --- lower vs different directional comparison):} Same proportion evidence compared between lower-tail and two-tailed framing.
    \item \textbf{Problem 9 (Stage 3 --- higher vs different directional comparison):} Same evidence comparison where higher-tail and two-tailed tests both reject but with different p-values.
    \item \textbf{Problem 10 (Stage 3 --- lower vs higher vs different comparison):} Same evidence tested under all three alternatives, yielding consistent fail-to-reject outcomes.
    \item \textbf{Problem 11 (Stage 4 --- near-boundary triple comparison):} Advanced proportion case emphasizing careful interpretation when evidence is near decision boundaries.
    \item \textbf{Problem 12 (Stage 4 --- one-direction-only rejection synthesis):} Shows a pattern where only one directional test rejects while opposite-direction and two-tailed tests do not.
    \item \textbf{Problem 13 (Stage 4 --- strongest directional synthesis):} Strong-evidence proportion case where one direction and the two-tailed test reject, but the opposite direction fails to reject.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C5\_practice\_activity.tex}}
This activity is now the dedicated C5 ladder for one-sample mean inference with known population standard deviation. It develops the mean \(z\)-test procedure from foundational setup to direction and significance-level sensitivity, then to same-evidence directional comparisons, and finally to synthesis structures that distinguish robust, near-boundary, and threshold-sensitive decisions using precise hypothesis-testing language.

\begin{itemize}
    \item \textbf{General Problem (Stage 1 --- mean-only full procedure):} Establishes parameter identification \((\mu)\), benchmark setup \((\mu_0)\), sample mean \((\bar x)\), known \(\sigma\), one-sample mean \(z\)-statistic, p-value logic, decision rule, and contextual conclusion template.
    \item \textbf{Problem 1 (Stage 1 --- introductory lower-tail mean test):} First direct lower-tail mean \(z\)-test with full worked inference sequence.
    \item \textbf{Problem 2 (Stage 1 --- introductory higher-tail mean test):} First direct higher-tail mean \(z\)-test reinforcing benchmark-above reasoning.
    \item \textbf{Problem 3 (Stage 1 --- introductory two-tailed mean test):} First direct two-tailed mean \(z\)-test for ``different'' claims.
    \item \textbf{Problem 4 (Stage 2 --- lower-tail \(\alpha\)-sensitivity):} Same mean evidence evaluated at 5\% and 1\% to show threshold-dependent decisions.
    \item \textbf{Problem 5 (Stage 2 --- higher-tail \(\alpha\)-sensitivity):} Same higher-tail mean evidence produces different conclusions under different \(\alpha\) levels.
    \item \textbf{Problem 6 (Stage 2 --- two-tailed \(\alpha\)-sensitivity):} Two-tailed mean case that highlights how fixed evidence can cross one threshold but not a stricter one.
    \item \textbf{Problem 7 (Stage 3 --- lower vs higher directional comparison):} Same data under opposite one-tailed alternatives to illustrate tail and decision reversal.
    \item \textbf{Problem 8 (Stage 3 --- lower vs different directional comparison):} Same evidence compared between lower-tail and two-tailed framing.
    \item \textbf{Problem 9 (Stage 3 --- higher vs different directional comparison):} Same evidence compared between higher-tail and two-tailed tests with both rejecting.
    \item \textbf{Problem 10 (Stage 3 --- lower vs higher vs different comparison):} Three-way same-evidence mean comparison with coherent fail-to-reject outcomes.
    \item \textbf{Problem 11 (Stage 4 --- robust non-rejection synthesis):} Advanced mean case where all alternatives fail to reject under strict significance control.
    \item \textbf{Problem 12 (Stage 4 --- one-direction-only near-boundary synthesis):} Near-boundary mean evidence used to interpret directional framing and non-rejection coherence.
    \item \textbf{Problem 13 (Stage 4 --- strongest directional synthesis):} Strong-signal mean case where one directional test and two-tailed test reject, while the opposite direction fails to reject.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C6\_practice\_activity.tex}}
This redesigned C6 activity is organized as a deliberate five-stage learning ladder that develops interpretation from foundational decision language to advanced policy synthesis. The progression keeps sample-size and margin-of-error awareness, p-value versus threshold reading, and decision-theoretic communication in context, while making Type I and Type II error consequences the central thread connecting every problem.

\begin{itemize}
    \item \textbf{Problem 1 (Stage 1 --- general framework for decision interpretation):} Establishes reject \(H_0\) versus fail to reject \(H_0\), clarifies that decisions are probabilistic rather than certain, defines Type I/Type II errors, and sets the rule that the immediate error concern depends on the reported decision.
    \item \textbf{Problem 2 (Stage 2):} Introductory single-test case with a reported reject-\(H_0\) outcome, requiring contextual interpretation of the supported claim and explicit Type I error interpretation for decision communication.
    \item \textbf{Problem 3 (Stage 2):} Introductory single-test case with fail-to-reject-\(H_0\), focusing on insufficient evidence language and corresponding Type II risk under non-rejection interpretation.
    \item \textbf{Problem 4 (Stage 3 --- same-evidence significance-level comparison):} Same p-value evaluated at \(\alpha=0.05\) and \(\alpha=0.01\), showing rejection at both thresholds and interpreting robust conclusions with persistent Type I caution.
    \item \textbf{Problem 5 (Stage 3 --- threshold-sensitive significance-level comparison):} Same p-value interpreted at strict and planning thresholds, producing reject at 5\% but fail to reject at 1\%, and highlighting the shift between Type I protection and Type II exposure.
    \item \textbf{Problem 6 (Stage 3 --- stable non-rejection across significance levels):} Same p-value interpreted at two thresholds with fail-to-reject outcomes at both, emphasizing stable non-rejection and continued Type II concern.
    \item \textbf{Problem 7 (Stage 3 to Stage 4 bridge --- Type I vs Type II prioritization):} Contextual comparison of Type I versus Type II consequences to justify why stricter or looser \(\alpha\) may be appropriate given different consequence priorities.
    \item \textbf{Problem 8 (Stage 4):} Context-driven error-cost analysis requiring interpretation of what Type I and Type II errors imply for resource allocation and missed-improvement consequences.
    \item \textbf{Problem 9 (Stage 4 --- precision and error-severity interpretation):} Interprets decisions while linking sample size and margin of error to evidence precision, then compares practical severity of false positive versus false negative conclusions.
    \item \textbf{Problem 10 (Stage 4 --- two-test dashboard synthesis):} Multi-test reading task requiring separation of hypotheses, decisions, and immediate error concerns for each metric in one policy report.
    \item \textbf{Problem 11 (Stage 4 --- three-test dashboard synthesis):} Extends dashboard reasoning across robust and threshold-sensitive findings, asking students to assign distinct error-priority cautions by indicator.
    \item \textbf{Problem 12 (Stage 5 --- cross-threshold synthesis):} Advanced cross-threshold task classifying results as robust, moderate, or tentative and linking each category to Type I/Type II consequence communication.
    \item \textbf{Problem 13 (Stage 5 --- final synthesis problem):} Strongest synthesis problem comparing policy choices under stricter versus more lenient false-positive tolerance, with explicit justification of which error is more costly in each domain.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C7\_practice\_activity.tex}}
This activity develops C7 as a progression in hypothesis-test interpretation by contrasting the p-value (evidence from data) with the reference probability \(\alpha\) (the chosen decision threshold). The sequence moves from foundational decision language to single-case comparisons, then to same-evidence directional framing, threshold sensitivity with fixed p-values, and final multi-case synthesis that classifies evidence patterns and consequences in context.

\begin{itemize}
    \item \textbf{General Problem (Stage 1 --- foundations of p-value comparison):} Defines p-value meaning, defines \(\alpha\) as reference probability, applies the rule ``if p-value \(\le \alpha\), reject \(H_0\); if p-value \(>\alpha\), fail to reject \(H_0\),'' and interprets the decision contextually.
    \item \textbf{Problem 1 (Stage 2 --- single-test clear reject case):} Compares a clearly small p-value with \(\alpha\), concludes reject \(H_0\), and interprets strong evidence against the null in context.
    \item \textbf{Problem 2 (Stage 2 --- single-test clear fail-to-reject case):} Compares a clearly large p-value with \(\alpha\), concludes fail to reject \(H_0\), and interprets non-significant evidence.
    \item \textbf{Problem 3 (Stage 2 --- near-boundary single-test case):} Uses a p-value close to \(\alpha\) to emphasize tentative evidence and careful wording of conclusions.
    \item \textbf{Problem 4 (Stage 3 --- same evidence, lower-tail framing):} Uses fixed sample evidence under \(H_a\!:\,\mu<\mu_0\), computes the corresponding p-value, compares to the same \(\alpha\), and interprets the directional conclusion.
    \item \textbf{Problem 5 (Stage 3 --- same evidence, higher-tail framing):} Uses the same data under \(H_a\!:\,\mu>\mu_0\), contrasts the changed p-value with the same \(\alpha\), and interprets why the decision differs from lower-tail framing.
    \item \textbf{Problem 6 (Stage 3 --- same evidence, two-tailed framing):} Uses the same data under \(H_a\!:\,\mu\ne\mu_0\), compares the two-tailed p-value to the same \(\alpha\), and explains conclusion stability or change.
    \item \textbf{Problem 7 (Stage 4 --- fixed p-value across \(\alpha=0.10,0.05,0.01\)):} Keeps p-value fixed and shows how changing reference probability changes reject/fail-to-reject decisions.
    \item \textbf{Problem 8 (Stage 4 --- threshold sensitivity in quality control):} Interprets one fixed p-value under operational and strict compliance thresholds and explains why decisions can differ.
    \item \textbf{Problem 9 (Stage 4 --- threshold sensitivity in public policy reporting):} Uses a fixed p-value with multiple significance standards to justify context-dependent threshold choice.
    \item \textbf{Problem 10 (Stage 5 --- multi-case classification dashboard I):} Compares several p-value/\(\alpha\) pairs and classifies each as robust rejection, threshold-sensitive rejection, borderline evidence, or clear non-significance.
    \item \textbf{Problem 11 (Stage 5 --- multi-case classification dashboard II):} Synthesizes decisions across service, finance, and education indicators and interprets practical consequences of each p-value/\(\alpha\) comparison.
\end{itemize}

\end{document}
