\input{../../../preamble/preamble_exams.tex}

\begin{document}
\ExamTitleBlock{11th grade}{Term 3 --- Lesson 1 Practice Activity (C4) with Solutions}{}

\section*{Contents}
\addcontentsline{toc}{section}{Contents}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{1}
\tableofcontents

\ExamSection{Evaluated criteria}
\begin{ExamCriteria}
    \item C4: Applies the general procedure for hypothesis testing with fixed probability in contextual situations.
\end{ExamCriteria}

\section{General Problem (Full Procedure Template)}
A benchmark claim may refer to either a population proportion or a population mean (with known population standard deviation). A hypothesis test is conducted at a given significance level \(\alpha\) to evaluate whether sample evidence supports a lower, higher, or different claim.

\textbf{Template A (proportion):} A benchmark proportion \(p_0\) is given. A random sample of size \(n\) yields \(x\) successes. Test \(H_0\colon p=p_0\) against a directional or two-tailed alternative.

\textbf{Template B (mean):} A benchmark mean \(\mu_0\) is given. A random sample of size \(n\) has sample mean \(\bar x\), with known population standard deviation \(\sigma\). Test \(H_0\colon \mu=\mu_0\) against a directional or two-tailed alternative.

\textbf{Questions}
\begin{enumerate}
    \item[(a)] Write hypotheses for lower, higher, and different alternatives in both templates.
    \item[(b)] State the condition checks required for a one-sample proportion \(z\)-test.
    \item[(c)] State the test statistic formulas for proportion and mean \(z\)-tests.
    \item[(d)] Explain how tail type changes p-value computation for lower-tail, higher-tail, and two-tailed tests.
    \item[(e)] State the decision rule using p-value and \(\alpha\), and write a contextual conclusion format.
\end{enumerate}

\subsection*{C4}
\begin{itemize}
    \item \textbf{Parameter identification:} Proportion template uses \(p\); mean template uses \(\mu\).
    \item \textbf{Hypotheses:}
    \[
    H_0\colon p=p_0,\quad H_a\colon p<p_0\ \text{or}\ p>p_0\ \text{or}\ p\ne p_0,
    \]
    \[
    H_0\colon \mu=\mu_0,\quad H_a\colon \mu<\mu_0\ \text{or}\ \mu>\mu_0\ \text{or}\ \mu\ne \mu_0.
    \]
    \item \textbf{Conditions for proportion test:} Verify \(np_0\ge 10\) and \(n(1-p_0)\ge 10\).
    \item \textbf{Test statistics:}
    \[
    z=\frac{\hat p-p_0}{\sqrt{\frac{p_0(1-p_0)}{n}}},\qquad
    z=\frac{\bar x-\mu_0}{\sigma/\sqrt{n}}.
    \]
    \item \textbf{Tail and p-value logic:} Lower-tail \(p\)-value is \(P(Z\le z_{obs})\); higher-tail is \(P(Z\ge z_{obs})\); two-tailed is \(2P(Z\ge |z_{obs}|)\).
    \item \textbf{Decision rule:} Reject \(H_0\) if p-value \(\le \alpha\); otherwise fail to reject \(H_0\).
    \item \textbf{Context conclusion format:} ``There is sufficient evidence'' (reject) or ``There is not enough evidence'' (fail to reject) for the claim stated in \(H_a\).
\end{itemize}

\section{Food delivery completion rate}
A local delivery app states that 64\% of assigned orders are completed without cancellation. In a random sample of 220 assigned orders, 129 are completed without cancellation. Using a significance level of 5\%, test whether the true completion proportion is lower than 64\%.

\subsection*{C4}
\begin{itemize}
    \item \textbf{Parameter:} \(p\) = true completion proportion.
    \item \textbf{Hypotheses:} \(H_0\colon p=0.64\), \(H_a\colon p<0.64\).
    \item \textbf{Conditions:} \(np_0=220(0.64)=140.8\ge 10\), \(n(1-p_0)=220(0.36)=79.2\ge 10\).
    \item \textbf{Sample statistic:} \(\hat p=129/220\approx 0.586\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{0.586-0.64}{\sqrt{\frac{0.64(0.36)}{220}}}\approx -1.66.
    \]
    \item \textbf{\(p\)-value:} \(P(Z\le -1.66)\approx 0.0485\).
    \item \textbf{Decision:} Reject \(H_0\) since \(0.0485<0.05\).
    \item \textbf{Conclusion:} There is sufficient evidence that the true completion rate is lower than 64\%.
\end{itemize}

\section{Average one-way commute time for office staff}
A company dashboard sets 42 minutes as the benchmark average one-way commute time for staff. A random sample of 49 employees has an average one-way commute of 43.0 minutes, and the population standard deviation is known to be 8 minutes. Using a significance level of 5\%, test whether the true average commute time is higher than 42 minutes.

\subsection*{C4}
\begin{itemize}
    \item \textbf{Parameter:} \(\mu\) = true average one-way commute time.
    \item \textbf{Hypotheses:} \(H_0\colon \mu=42\), \(H_a\colon \mu>42\).
    \item \textbf{Sample statistic:} \(\bar x=43.0\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{43.0-42}{8/\sqrt{49}}=\frac{1.0}{1.143}\approx 0.88.
    \]
    \item \textbf{\(p\)-value:} \(P(Z\ge 0.88)\approx 0.1894\).
    \item \textbf{Decision:} Fail to reject \(H_0\) since \(0.1894>0.05\).
    \item \textbf{Conclusion:} There is not enough evidence that the true average commute time is higher than 42 minutes.
\end{itemize}

\section{Public health screening attendance benchmark (lower vs higher)}
A district health office reports that 58\% of invited adults attend a free screening. In a random sample of 300 invited adults, 156 attend.

\textbf{Questions}
\begin{enumerate}
    \item[(a)] At \(\alpha=5\%\), test whether the true attendance proportion is lower than 58\%.
    \item[(b)] Using the same sample evidence and the same \(\alpha\), test whether the true attendance proportion is higher than 58\%.
\end{enumerate}

\subsection*{C4}
\textbf{Common setup}
\begin{itemize}
    \item \(p\) = true attendance proportion; \(p_0=0.58\); \(\hat p=156/300=0.52\).
    \item Conditions: \(np_0=174\ge10\), \(n(1-p_0)=126\ge10\).
    \item Test statistic:
    \[
    z=\frac{0.52-0.58}{\sqrt{\frac{0.58(0.42)}{300}}}\approx -2.11.
    \]
\end{itemize}

\textbf{Part (a): Lower-tail test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon p=0.58\), \(H_a\colon p<0.58\).
    \item \textbf{\(p\)-value:} \(P(Z\le -2.11)\approx 0.0174\).
    \item \textbf{Decision:} Reject \(H_0\) since \(0.0174<0.05\).
    \item \textbf{Conclusion:} There is sufficient evidence that attendance is lower than 58\%.
\end{itemize}

\textbf{Part (b): Higher-tail test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon p=0.58\), \(H_a\colon p>0.58\).
    \item \textbf{\(p\)-value:} \(P(Z\ge -2.11)\approx 0.9826\).
    \item \textbf{Decision:} Fail to reject \(H_0\) since \(0.9826>0.05\).
    \item \textbf{Conclusion:} There is not enough evidence that attendance is higher than 58\%.
\end{itemize}

\textbf{Comparison note:} The sample evidence is identical in both parts; only the direction in \(H_a\) changes the tail area and therefore the decision.

\section{Average packaging time in a school enterprise lab (lower vs higher)}
A school enterprise club monitors packaging time for student-made products. The benchmark mean packaging time is 12.0 minutes. A random sample of 64 packages has mean time 11.0 minutes, with known population standard deviation 3.2 minutes.

\textbf{Questions}
\begin{enumerate}
    \item[(a)] At \(\alpha=5\%\), test whether the true mean packaging time is lower than 12.0 minutes.
    \item[(b)] Using the same data, test whether the true mean packaging time is higher than 12.0 minutes.
\end{enumerate}

\subsection*{C4}
\textbf{Common setup}
\begin{itemize}
    \item \(\mu\) = true mean packaging time, \(\mu_0=12.0\), \(\bar x=11.0\), \(\sigma=3.2\), \(n=64\).
    \item Test statistic:
    \[
    z=\frac{11.0-12.0}{3.2/\sqrt{64}}=\frac{-1.0}{0.4}=-2.50.
    \]
\end{itemize}

\textbf{Part (a): Lower-tail test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon \mu=12.0\), \(H_a\colon \mu<12.0\).
    \item \textbf{\(p\)-value:} \(P(Z\le -2.50)\approx 0.0062\).
    \item \textbf{Decision:} Reject \(H_0\) since \(0.0062<0.05\).
    \item \textbf{Conclusion:} There is sufficient evidence that the true mean packaging time is lower than 12.0 minutes.
\end{itemize}

\textbf{Part (b): Higher-tail test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon \mu=12.0\), \(H_a\colon \mu>12.0\).
    \item \textbf{\(p\)-value:} \(P(Z\ge -2.50)\approx 0.9938\).
    \item \textbf{Decision:} Fail to reject \(H_0\) since \(0.9938>0.05\).
    \item \textbf{Conclusion:} There is not enough evidence that the true mean packaging time is higher than 12.0 minutes.
\end{itemize}

\textbf{Comparison note:} Tail direction determines whether the observed negative \(z\)-value gives small evidence or very large p-value.

\section{Average monthly public transport spending benchmark (same data, two \(\alpha\) levels)}
A mobility report sets 185 dollars as the benchmark average monthly public transport spending. A random sample of 50 commuters has mean spending 191.2 dollars, and the population standard deviation is known to be 20 dollars. Test whether the true mean is different from 185 dollars at \(\alpha=5\%\) and \(\alpha=1\%\).

\subsection*{C4}
\begin{itemize}
    \item \textbf{Parameter:} \(\mu\) = true mean monthly spending.
    \item \textbf{Hypotheses:} \(H_0\colon \mu=185\), \(H_a\colon \mu\ne 185\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{191.2-185}{20/\sqrt{50}}\approx 2.19.
    \]
    \item \textbf{\(p\)-value:} \(2P(Z\ge 2.19)\approx 0.0286\).
    \item \textbf{Decision at 5\%:} Reject \(H_0\) because \(0.0286<0.05\).
    \item \textbf{Decision at 1\%:} Fail to reject \(H_0\) because \(0.0286>0.01\).
    \item \textbf{Conclusion:} The evidence is unchanged; only \(\alpha\) changes. A policy using 5\% finds a difference, while a stricter 1\% policy does not.
\end{itemize}

\section{Weekly side-income amount from freelance tasks (two-tailed, robust across \(\alpha\))}
A financial planning blog uses 310 dollars per week as a benchmark for average side-income from freelance tasks. A random sample of 64 workers has mean side-income 296 dollars, and the population standard deviation is known to be 40 dollars. Test whether the true mean differs from 310 dollars at \(\alpha=5\%\) and \(\alpha=1\%\).

\subsection*{C4}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon \mu=310\), \(H_a\colon \mu\ne 310\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{296-310}{40/\sqrt{64}}=\frac{-14}{5}=-2.80.
    \]
    \item \textbf{\(p\)-value:} \(2P(Z\le -2.80)\approx 0.0051\).
    \item \textbf{Decision at 5\%:} Reject \(H_0\) because \(0.0051<0.05\).
    \item \textbf{Decision at 1\%:} Reject \(H_0\) because \(0.0051<0.01\).
    \item \textbf{Conclusion:} There is sufficient evidence at both levels that mean side-income differs from 310 dollars.
\end{itemize}

\section{Use of monthly budgeting reminders: lower vs higher}
A finance app claims that 52\% of users keep monthly budgeting reminders turned on. In a random sample of 250 users, 133 have reminders turned on. Use \(\alpha=5\%\) for both parts.

\textbf{Questions}
\begin{enumerate}
    \item[(a)] Test whether the true proportion is lower than 52\%.
    \item[(b)] Test whether the true proportion is higher than 52\%.
\end{enumerate}

\subsection*{C4}
\textbf{Common setup}
\begin{itemize}
    \item \(p_0=0.52\), \(\hat p=133/250=0.532\), \(np_0=130\ge10\), \(n(1-p_0)=120\ge10\).
    \item \(z=\dfrac{0.532-0.52}{\sqrt{\frac{0.52(0.48)}{250}}}\approx 0.38\).
\end{itemize}

\textbf{Part (a): Lower-tail test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon p=0.52\), \(H_a\colon p<0.52\).
    \item \textbf{\(p\)-value:} \(P(Z\le 0.38)\approx 0.6480\).
    \item \textbf{Decision:} Fail to reject \(H_0\).
    \item \textbf{Conclusion:} There is not enough evidence that the true proportion is lower than 52\%.
\end{itemize}

\textbf{Part (b): Higher-tail test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon p=0.52\), \(H_a\colon p>0.52\).
    \item \textbf{\(p\)-value:} \(P(Z\ge 0.38)\approx 0.3520\).
    \item \textbf{Decision:} Fail to reject \(H_0\).
    \item \textbf{Conclusion:} There is not enough evidence that the true proportion is higher than 52\%.
\end{itemize}

\textbf{Comparison note:} This is an all-fail-to-reject comparison case for opposite one-tailed alternatives.

\section{Average evening study time: lower vs different}
An academic support program uses 2.6 hours as the benchmark mean evening study time. A random sample of 100 students has mean evening study time 2.5 hours, with known population standard deviation 0.8 hours. Use \(\alpha=5\%\).

\textbf{Questions}
\begin{enumerate}
    \item[(a)] Test whether the true mean study time is lower than 2.6 hours.
    \item[(b)] Using the same data, test whether the true mean study time is different from 2.6 hours.
\end{enumerate}

\subsection*{C4}
\textbf{Common setup}
\begin{itemize}
    \item \(\mu_0=2.6\), \(\bar x=2.5\), \(\sigma=0.8\), \(n=100\).
    \item \(z=\dfrac{2.5-2.6}{0.8/\sqrt{100}}=\dfrac{-0.1}{0.08}=-1.25\).
\end{itemize}

\textbf{Part (a): Lower-tail test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon \mu=2.6\), \(H_a\colon \mu<2.6\).
    \item \textbf{\(p\)-value:} \(P(Z\le -1.25)\approx 0.1056\).
    \item \textbf{Decision:} Fail to reject \(H_0\) since \(0.1056>0.05\).
\end{itemize}

\textbf{Part (b): Two-tailed test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon \mu=2.6\), \(H_a\colon \mu\ne 2.6\).
    \item \textbf{\(p\)-value:} \(2P(Z\le -1.25)\approx 0.2112\).
    \item \textbf{Decision:} Fail to reject \(H_0\) since \(0.2112>0.05\).
\end{itemize}

\textbf{Conclusion:} Both framings fail to reject \(H_0\), but the two-tailed p-value is larger because it counts both tails.

\section{Customer support wait-time benchmark: higher vs different}
A service center sets 6.0 minutes as the benchmark mean wait time. A random sample of 100 customer calls has mean wait time 6.50 minutes, with known population standard deviation 1.5 minutes. Use \(\alpha=5\%\).

\textbf{Questions}
\begin{enumerate}
    \item[(a)] Test whether the true mean wait time is higher than 6.0 minutes.
    \item[(b)] Using the same data, test whether the true mean wait time is different from 6.0 minutes.
\end{enumerate}

\subsection*{C4}
\textbf{Common setup}
\begin{itemize}
    \item \(\mu_0=6.0\), \(\bar x=6.50\), \(\sigma=1.5\), \(n=100\).
    \item \(z=\dfrac{6.50-6.0}{1.5/\sqrt{100}}=\dfrac{0.50}{0.15}=3.33\).
\end{itemize}

\textbf{Part (a): Higher-tail test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon \mu=6.0\), \(H_a\colon \mu>6.0\).
    \item \textbf{\(p\)-value:} \(P(Z\ge 3.33)\approx 0.0004\).
    \item \textbf{Decision:} Reject \(H_0\) since \(0.0004<0.05\).
\end{itemize}

\textbf{Part (b): Two-tailed test}
\begin{itemize}
    \item \textbf{Hypotheses:} \(H_0\colon \mu=6.0\), \(H_a\colon \mu\ne 6.0\).
    \item \textbf{\(p\)-value:} \(2P(Z\ge 3.33)\approx 0.0008\).
    \item \textbf{Decision:} Reject \(H_0\) since \(0.0008<0.05\).
\end{itemize}

\textbf{Conclusion:} The directional and two-tailed tests both reject \(H_0\), while the practical interpretation remains that average wait time is above the benchmark.

\section{Use of monthly budgeting reminders with near-boundary evidence (higher, lower, different)}
A finance app claims that 52\% of users keep monthly budgeting reminders turned on. In a random sample of 300 users, 142 have reminders turned on. Use \(\alpha=5\%\) for all parts.

\textbf{Questions}
\begin{enumerate}
    \item[(a)] Test whether the true proportion is higher than 52\%.
    \item[(b)] Test whether the true proportion is lower than 52\%.
    \item[(c)] Test whether the true proportion is different from 52\%.
\end{enumerate}

\subsection*{C4}
\textbf{Common setup}
\begin{itemize}
    \item \(p_0=0.52\), \(\hat p=142/300\approx 0.4733\).
    \item Conditions: \(np_0=156\ge10\), \(n(1-p_0)=144\ge10\).
    \item \(z=\dfrac{0.4733-0.52}{\sqrt{\frac{0.52(0.48)}{300}}}\approx -1.62\).
\end{itemize}

\textbf{Part (a): Higher-tail test}
\begin{itemize}
    \item \(H_0\colon p=0.52\), \(H_a\colon p>0.52\).
    \item \(p\)-value \(=P(Z\ge -1.62)\approx 0.9471\).
    \item Decision: Fail to reject \(H_0\).
\end{itemize}

\textbf{Part (b): Lower-tail test}
\begin{itemize}
    \item \(H_0\colon p=0.52\), \(H_a\colon p<0.52\).
    \item \(p\)-value \(=P(Z\le -1.62)\approx 0.0529\).
    \item Decision: Fail to reject \(H_0\) (near boundary, since \(0.0529\) is slightly above \(0.05\)).
\end{itemize}

\textbf{Part (c): Two-tailed test}
\begin{itemize}
    \item \(H_0\colon p=0.52\), \(H_a\colon p\ne 0.52\).
    \item \(p\)-value \(=2P(Z\le -1.62)\approx 0.1058\).
    \item Decision: Fail to reject \(H_0\).
\end{itemize}

\textbf{Conclusion:} This near-boundary case illustrates careful p-value versus \(\alpha\) comparison; all three tests fail to reject \(H_0\).

\section{Battery-life compliance synthesis under multiple framings}
A school technology center claims that a new battery model has mean runtime 10.0 hours. In a random sample of \(n=100\) batteries, the mean runtime is 10.30 hours, and the population standard deviation is known to be 1.2 hours.

\textbf{Questions}
\begin{enumerate}
    \item[(a)] At \(\alpha=5\%\), test \(H_a\colon \mu>10.0\).
    \item[(b)] At \(\alpha=5\%\), test \(H_a\colon \mu<10.0\).
    \item[(c)] At \(\alpha=5\%\), test \(H_a\colon \mu\ne 10.0\).
    \item[(d)] Re-evaluate the two-tailed test from part (c) at \(\alpha=1\%\).
\end{enumerate}

\subsection*{C4}
\textbf{Common computation}
\begin{itemize}
    \item \(\mu_0=10.0\), \(\bar x=10.30\), \(\sigma=1.2\), \(n=100\).
    \item \(z=\dfrac{10.30-10.0}{1.2/\sqrt{100}}=\dfrac{0.30}{0.12}=2.50\).
\end{itemize}

\textbf{Part (a): Higher-tail at 5\%}
\begin{itemize}
    \item \(H_0\colon \mu=10.0\), \(H_a\colon \mu>10.0\).
    \item \(p\)-value \(=P(Z\ge 2.50)\approx 0.0062\).
    \item Decision: Reject \(H_0\) since \(0.0062<0.05\).
\end{itemize}

\textbf{Part (b): Lower-tail at 5\%}
\begin{itemize}
    \item \(H_0\colon \mu=10.0\), \(H_a\colon \mu<10.0\).
    \item \(p\)-value \(=P(Z\le 2.50)\approx 0.9938\).
    \item Decision: Fail to reject \(H_0\).
\end{itemize}

\textbf{Part (c): Two-tailed at 5\%}
\begin{itemize}
    \item \(H_0\colon \mu=10.0\), \(H_a\colon \mu\ne 10.0\).
    \item \(p\)-value \(=2P(Z\ge 2.50)\approx 0.0124\).
    \item Decision: Reject \(H_0\) since \(0.0124<0.05\).
\end{itemize}

\textbf{Part (d): Two-tailed at 1\%}
\begin{itemize}
    \item Same evidence and same p-value \(0.0124\), but \(\alpha=0.01\).
    \item Decision: Fail to reject \(H_0\) since \(0.0124>0.01\).
\end{itemize}

\textbf{Synthesis conclusion:} Direction is essential: higher-tail rejects while lower-tail does not for the same data. At 5\%, both the higher-tail and corresponding two-tailed test reject; at 1\%, the stricter standard changes the two-tailed decision to fail to reject.

\end{document}
