\input{../../../preamble/preamble_exams.tex}

\begin{document}
\ExamTitleBlock{11th grade}{Term 3 --- Lesson 1 Practice Activity (C5) with Solutions}{}

\section*{Contents}
\addcontentsline{toc}{section}{Contents}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{1}
\tableofcontents

\ExamSection{Evaluated criteria}
\begin{ExamCriteria}
    \item C5: Calculates hypothesis tests on the mean of a population.
\end{ExamCriteria}


\section{General Problem}
A benchmark claim may refer to a population mean. A hypothesis test is conducted at a given significance level \(\alpha\) to evaluate whether the sample evidence supports or contradicts the benchmark value.

\textbf{Mean template:} A benchmark mean value \(\mu_0\) is given for a population mean. A sample of size \(n\) is collected with sample mean \(\bar x\). The population standard deviation \(\sigma\) is known. Using a significance level \(\alpha\), test whether the true population mean is lower than, higher than, or different from the benchmark mean.

\textbf{Questions}
\begin{enumerate}
	\item[(a)] Identify the population parameter and write hypotheses for the case where the true parameter is lower than the benchmark.
	\item[(b)] Write hypotheses for the case where the true parameter is higher than the benchmark.
	\item[(c)] Write hypotheses for the case where the true parameter is different from the benchmark.
	\item[(d)] Write the one-sample mean \(z\)-test statistic and explain how the p-value is obtained for each tail type.
	\item[(e)] State the decision rule using \(\alpha\), and write a contextual conclusion template for reject/fail-to-reject outcomes.
\end{enumerate}

\subsection*{C5}
\begin{itemize}
	\item Let \(\mu\) denote the true population mean for the context.
	\item \textbf{Hypotheses by direction:}
	\[
	H_0\colon \mu=\mu_0,
	\qquad
	H_1\colon \mu<\mu_0\ \text{or}\ \mu>\mu_0\ \text{or}\ \mu\ne\mu_0.
	\]
	\item \textbf{Test statistic (known \(\sigma\)):}
	\[
	z=\frac{\bar x-\mu_0}{\sigma/\sqrt{n}}.
	\]
	\item \textbf{p-value rules:}
	\begin{itemize}
		\item Left-tailed: \(p=P(Z\le z_{\text{obs}})\).
		\item Right-tailed: \(p=P(Z\ge z_{\text{obs}})\).
		\item Two-tailed: \(p=2P(Z\ge |z_{\text{obs}}|)\).
	\end{itemize}
	\item \textbf{Decision rule:} Reject \(H_0\) if \(p\le\alpha\); otherwise fail to reject \(H_0\).
	\item \textbf{Conclusion template:}
	\begin{itemize}
		\item Reject \(H_0\): ``There is sufficient evidence at significance level \(\alpha\) to support \(H_1\) in context.''
		\item Fail to reject \(H_0\): ``There is insufficient evidence at significance level \(\alpha\) to support \(H_1\) in context.''
	\end{itemize}
\end{itemize}

\section{Average monthly spending on phone plans}
A mobile provider says customers spend an average of 45 dollars per month on phone plans. A random sample of 64 customers has an average of 46.2 dollars per month. The known long-run standard deviation is 8 dollars. Use a 5\% significance level to test whether the true mean spending is higher than 45 dollars.

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean monthly spending on phone plans.
    \item \textbf{Hypotheses:} \(H_0\colon \mu=45\), \(H_1\colon \mu>45\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{46.2-45}{8/\sqrt{64}}=\frac{1.2}{1}=1.20.
    \]
    \item \textbf{\(p\)-value:} \(P(Z\ge 1.20)\approx 0.1151\).
    \item \textbf{Decision:} Fail to reject \(H_0\) since \(0.1151\ge 0.05\).
    \item \textbf{Conclusion:} There is insufficient evidence that the true average monthly phone-plan spending is above 45 dollars.
\end{itemize}

\section{Average weekly grocery budget}
A budgeting app reports that users spend 180 dollars per week on groceries on average. A random sample of 100 users shows an average of 176.4 dollars per week. The known long-run standard deviation is 18 dollars. Use a 5\% significance level to test whether the true mean is lower than 180 dollars.

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean weekly grocery spending.
    \item \textbf{Hypotheses:} \(H_0\colon \mu=180\), \(H_1\colon \mu<180\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{176.4-180}{18/\sqrt{100}}=\frac{-3.6}{1.8}=-2.00.
    \]
    \item \textbf{\(p\)-value:} \(P(Z\le -2.00)\approx 0.0228\).
    \item \textbf{Decision:} Reject \(H_0\) since \(0.0228<0.05\).
    \item \textbf{Conclusion:} There is sufficient evidence that the true average weekly grocery spending is below 180 dollars.
\end{itemize}

\section{Average monthly commuting cost}
A workplace program states that employees spend 220 dollars per month on commuting on average. A random sample of 64 employees has an average of 224 dollars per month. The known long-run standard deviation is 15 dollars. Use a 5\% significance level to test whether the true mean differs from 220 dollars.

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean monthly commuting cost.
    \item \textbf{Hypotheses:} \(H_0\colon \mu=220\), \(H_1\colon \mu\ne 220\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{224-220}{15/\sqrt{64}}=\frac{4}{1.875}\approx 2.13.
    \]
    \item \textbf{\(p\)-value:} \(2P(Z\ge 2.13)\approx 0.0332\).
    \item \textbf{Decision:} Reject \(H_0\) since \(0.0332<0.05\).
    \item \textbf{Conclusion:} There is sufficient evidence that the true mean monthly commuting cost is different from 220 dollars.
\end{itemize}

\section{Average sodium intake in ready meals}
A health committee uses 730 mg as a benchmark for average sodium intake per ready meal. A random sample of 100 meals has a mean sodium content of 727.36 mg. The population standard deviation is known to be 12 mg. Test whether the true mean sodium intake is lower than 730 mg using two significance levels: 5\% and 1\%.

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean sodium content per ready meal.
    \item \textbf{Hypotheses:} \(H_0\colon \mu=730\), \(H_1\colon \mu<730\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{727.36-730}{12/\sqrt{100}}=\frac{-2.64}{1.2}=-2.20.
    \]
    \item \textbf{\(p\)-value:} \(P(Z\le -2.20)\approx 0.0139\).
    \item \textbf{Decision at 5\%:} Reject \(H_0\) since \(0.0139<0.05\).
    \item \textbf{Decision at 1\%:} Fail to reject \(H_0\) since \(0.0139\ge 0.01\).
    \item \textbf{Interpretation of sensitivity:} The sample evidence and \(z\)-statistic are unchanged across both tests; only \(\alpha\) changes, and the conclusion changes because the p-value lies between the two significance thresholds.
    \item \textbf{Conclusion:} There is evidence at 5\% (but not at 1\%) that mean sodium intake is below 730 mg.
\end{itemize}

\section{Average monthly savings in emergency funds}
A financial planner claims that workers save 240 dollars per month in emergency funds on average. A random sample of 64 workers has an average monthly saving of 243.3 dollars. The population standard deviation is known to be 12 dollars. Test whether the true mean is higher than 240 dollars at two significance levels: 5\% and 1\%.

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean monthly emergency-fund savings.
    \item \textbf{Hypotheses:} \(H_0\colon \mu=240\), \(H_1\colon \mu>240\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{243.3-240}{12/\sqrt{64}}=\frac{3.3}{1.5}=2.20.
    \]
    \item \textbf{\(p\)-value:} \(P(Z\ge 2.20)\approx 0.0139\).
    \item \textbf{Decision at 5\%:} Reject \(H_0\) since \(0.0139<0.05\).
    \item \textbf{Decision at 1\%:} Fail to reject \(H_0\) since \(0.0139\ge 0.01\).
    \item \textbf{Interpretation of sensitivity:} With fixed sample evidence, the conclusion changes because the same evidence is compared with a stricter threshold.
    \item \textbf{Conclusion:} There is evidence at 5\% (but not at 1\%) that mean monthly emergency-fund savings exceed 240 dollars.
\end{itemize}

\section{Average mathematics benchmark score}
A district benchmark for a mathematics test is 70 points. A random sample of 196 students has an average score of 71.4 points. The population standard deviation is known to be 9.8 points. Test whether the true mean score is different from 70 points using two significance levels: 5\% and 1\%.

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean mathematics score.
    \item \textbf{Hypotheses:} \(H_0\colon \mu=70\), \(H_1\colon \mu\ne 70\).
    \item \textbf{Test statistic:}
    \[
    z=\frac{71.4-70}{9.8/\sqrt{196}}=\frac{1.4}{0.7}=2.00.
    \]
    \item \textbf{\(p\)-value:} \(2P(Z\ge 2.00)\approx 0.0456\).
    \item \textbf{Decision at 5\%:} Reject \(H_0\) since \(0.0456<0.05\).
    \item \textbf{Decision at 1\%:} Fail to reject \(H_0\) since \(0.0456\ge 0.01\).
    \item \textbf{Interpretation of sensitivity:} The sample statistics and \(z\)-value are identical in both checks; only \(\alpha\) changes, so the final decision becomes threshold-dependent.
    \item \textbf{Conclusion:} The sample indicates a difference at 5\% but not at 1\%.
\end{itemize}

\section{Average monthly gym visits}
A health insurer sets a benchmark of 10 gym visits per month on average. A random sample of 81 members has an average of 9.2 visits. The known long-run standard deviation is 2.4 visits. Use a 5\% significance level.

\begin{enumerate}
    \item[(a)] Test whether the true mean is lower than 10 visits.
    \item[(b)] Test whether the true mean is higher than 10 visits.
\end{enumerate}

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean monthly gym visits.
    \item \textbf{Common test statistic from the same evidence:}
    \[
    z=\frac{9.2-10}{2.4/\sqrt{81}}=\frac{-0.8}{0.2667}\approx -3.00.
    \]
    \item \textbf{(a) Lower-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=10\), \(H_1\colon \mu<10\).
        \item Tail type: left-tailed.
        \item \(p\)-value: \(P(Z\le -3.00)\approx 0.0013\).
        \item Decision: Reject \(H_0\) since \(0.0013<0.05\).
    \end{itemize}
    \item \textbf{(b) Higher-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=10\), \(H_1\colon \mu>10\).
        \item Tail type: right-tailed.
        \item \(p\)-value: \(P(Z\ge -3.00)\approx 0.9987\).
        \item Decision: Fail to reject \(H_0\) since \(0.9987\ge 0.05\).
    \end{itemize}
    \item \textbf{Comparative interpretation:} The sample evidence is fixed, but changing the alternative hypothesis changes the tail and p-value; this leads to opposite decisions.
\end{itemize}

\section{Average focus hours in deep-work sessions}
A productivity team states that workers average 15 focus hours per week in deep-work sessions. A random sample of 64 workers has an average of 14 hours. The known long-run standard deviation is 4 hours. Use a 5\% significance level.

\begin{enumerate}
    \item[(a)] Test whether the true mean is lower than 15 hours.
    \item[(b)] Test whether the true mean is different from 15 hours.
\end{enumerate}

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean focus hours per week.
    \item \textbf{Common test statistic from the same evidence:}
    \[
    z=\frac{14-15}{4/\sqrt{64}}=\frac{-1}{0.5}=-2.00.
    \]
    \item \textbf{(a) Lower-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=15\), \(H_1\colon \mu<15\).
        \item Tail type: left-tailed.
        \item \(p\)-value: \(P(Z\le -2.00)\approx 0.0228\).
        \item Decision: Reject \(H_0\) since \(0.0228<0.05\).
    \end{itemize}
    \item \textbf{(b) Different-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=15\), \(H_1\colon \mu\ne 15\).
        \item Tail type: two-tailed.
        \item \(p\)-value: \(2P(Z\ge 2.00)\approx 0.0456\).
        \item Decision: Reject \(H_0\) since \(0.0456<0.05\).
    \end{itemize}
    \item \textbf{Comparative interpretation:} The two-tailed p-value is larger because evidence must account for both tails, but here both tests still reject at 5\%.
\end{itemize}

\section{Average monthly commuting distance}
A transport report sets a benchmark of 420 kilometers per month for commuting distance. A random sample of 100 workers has an average of 430 kilometers. The known long-run standard deviation is 25 kilometers. Use a 5\% significance level.

\begin{enumerate}
    \item[(a)] Test whether the true mean is higher than 420 kilometers.
    \item[(b)] Test whether the true mean is different from 420 kilometers.
\end{enumerate}

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean monthly commuting distance.
    \item \textbf{Common test statistic from the same evidence:}
    \[
    z=\frac{430-420}{25/\sqrt{100}}=\frac{10}{2.5}=4.00.
    \]
    \item \textbf{(a) Higher-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=420\), \(H_1\colon \mu>420\).
        \item Tail type: right-tailed.
        \item \(p\)-value: \(P(Z\ge 4.00)\approx 0.0000\).
        \item Decision: Reject \(H_0\) since \(0.0000<0.05\).
    \end{itemize}
    \item \textbf{(b) Different-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=420\), \(H_1\colon \mu\ne 420\).
        \item Tail type: two-tailed.
        \item \(p\)-value: \(2P(Z\ge 4.00)\approx 0.0001\).
        \item Decision: Reject \(H_0\) since \(0.0001<0.05\).
    \end{itemize}
    \item \textbf{Comparative interpretation:} Both tests reject because the evidence is very strong in the higher direction; the two-tailed p-value is larger than the one-tailed p-value but remains below \(\alpha\).
\end{itemize}

\section{Average monthly streaming subscriptions cost}
A personal finance tracker uses 32 dollars as the average monthly subscriptions-cost benchmark. A random sample of 49 workers has an average of 32.7 dollars. The known long-run standard deviation is 7 dollars. Use a 5\% significance level.

\begin{enumerate}
    \item[(a)] Test whether the true mean is lower than 32 dollars.
    \item[(b)] Test whether the true mean is higher than 32 dollars.
    \item[(c)] Test whether the true mean is different from 32 dollars.
\end{enumerate}

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean monthly subscriptions cost.
    \item \textbf{Common test statistic from the same evidence:}
    \[
    z=\frac{32.7-32}{7/\sqrt{49}}=\frac{0.7}{1}=0.70.
    \]
    \item \textbf{(a) Lower-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=32\), \(H_1\colon \mu<32\).
        \item Tail type: left-tailed.
        \item \(p\)-value: \(P(Z\le 0.70)\approx 0.7580\).
        \item Decision: Fail to reject \(H_0\) since \(0.7580\ge 0.05\).
    \end{itemize}
    \item \textbf{(b) Higher-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=32\), \(H_1\colon \mu>32\).
        \item Tail type: right-tailed.
        \item \(p\)-value: \(P(Z\ge 0.70)\approx 0.2420\).
        \item Decision: Fail to reject \(H_0\) since \(0.2420\ge 0.05\).
    \end{itemize}
    \item \textbf{(c) Different-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=32\), \(H_1\colon \mu\ne 32\).
        \item Tail type: two-tailed.
        \item \(p\)-value: \(2P(Z\ge 0.70)\approx 0.4840\).
        \item Decision: Fail to reject \(H_0\) since \(0.4840\ge 0.05\).
    \end{itemize}
    \item \textbf{Comparative interpretation:} With weak evidence close to the benchmark, all three directional formulations fail to reject \(H_0\), showing a consistent non-signal pattern.
\end{itemize}

\section{Average response time in a customer support center}
A service center benchmark is 12.0 minutes for mean response time. A random sample of 196 calls has an average response time of 11.3 minutes. The population standard deviation is known to be 4.9 minutes. Use a 1\% significance level.

\begin{enumerate}
    \item[(a)] Test whether the true mean response time is lower than 12.0 minutes.
    \item[(b)] Test whether the true mean response time is higher than 12.0 minutes.
    \item[(c)] Test whether the true mean response time is different from 12.0 minutes.
\end{enumerate}

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean response time.
    \item \textbf{Common test statistic from the same evidence:}
    \[
    z=\frac{11.3-12.0}{4.9/\sqrt{196}}=\frac{-0.7}{0.35}=-2.00.
    \]
    \item \textbf{(a) Lower-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=12.0\), \(H_1\colon \mu<12.0\).
        \item \(p\)-value: \(P(Z\le -2.00)\approx 0.0228\).
        \item Decision at \(\alpha=0.01\): Fail to reject \(H_0\) since \(0.0228\ge 0.01\).
    \end{itemize}
    \item \textbf{(b) Higher-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=12.0\), \(H_1\colon \mu>12.0\).
        \item \(p\)-value: \(P(Z\ge -2.00)\approx 0.9772\).
        \item Decision: Fail to reject \(H_0\).
    \end{itemize}
    \item \textbf{(c) Different-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=12.0\), \(H_1\colon \mu\ne 12.0\).
        \item \(p\)-value: \(2P(Z\ge 2.00)\approx 0.0456\).
        \item Decision at \(\alpha=0.01\): Fail to reject \(H_0\).
    \end{itemize}
    \item \textbf{Synthesis interpretation:} At this strict policy threshold, all compared tests fail to reject. The direction matters for p-values, but none are small enough to cross \(\alpha=0.01\).
\end{itemize}

\section{Average number of sick days used per year}
An employer benchmark is 8 sick days used per year on average. A random sample of 144 employees has an average of 7.4 days. The known long-run standard deviation is 2.4 days. Use a 5\% significance level.

\begin{enumerate}
    \item[(a)] Test whether the true mean is higher than 8 days.
    \item[(b)] Test whether the true mean is lower than 8 days.
    \item[(c)] Test whether the true mean is different from 8 days.
\end{enumerate}

\subsection*{C5}
\begin{itemize}
    \item Let \(\mu\) be the true mean number of sick days used per year.
    \item \textbf{Common test statistic from the same evidence:}
    \[
    z=\frac{7.4-8}{2.4/\sqrt{144}}=\frac{-0.6}{0.2}=-3.00.
    \]
    \item \textbf{(a) Higher-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=8\), \(H_1\colon \mu>8\).
        \item \(p\)-value: \(P(Z\ge -3.00)\approx 0.9987\).
        \item Decision: Fail to reject \(H_0\) since \(0.9987\ge 0.05\).
    \end{itemize}
    \item \textbf{(b) Lower-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=8\), \(H_1\colon \mu<8\).
        \item \(p\)-value: \(P(Z\le -3.00)\approx 0.0013\).
        \item Decision: Reject \(H_0\) since \(0.0013<0.05\).
    \end{itemize}
    \item \textbf{(c) Different-than test}
    \begin{itemize}
        \item Hypotheses: \(H_0\colon \mu=8\), \(H_1\colon \mu\ne 8\).
        \item \(p\)-value: \(2P(Z\le -3.00)\approx 0.0026\).
        \item Decision: Reject \(H_0\) since \(0.0026<0.05\).
    \end{itemize}
    \item \textbf{Synthesis interpretation:} The lower-direction and two-tailed tests both reject, while the opposite directional test does not reject. This pattern occurs because the sample mean is substantially below the benchmark.
\end{itemize}

\end{document}
