\makeatletter
\def\input@path{{./}{../}{../../}{../../../}{../../../../}{preamble/}{../preamble/}{../../preamble/}{../../../preamble/}}
\makeatother	

\input{preamble_exams.tex}

\title{Term 3 --- Lesson 1 Practice Activity 3 (C1--C3) with Solutions}
\author{}
\date{}

\begin{document}
	
	\maketitle

	\setcounter{tocdepth}{1}
	\setcounter{secnumdepth}{2}
	\tableofcontents

	\section*{Evaluated criteria}
	\begin{itemize}
		\item C1: Identify whether the context uses a population proportion or a population mean and map the symbols correctly in context.
		\item C2: Write the null and alternative hypotheses in correct notation, keeping the null as ``parameter equals benchmark.''
		\item C3: Classify the test direction (left-tailed, right-tailed, or two-tailed) and justify it from the wording of the claim.
	\end{itemize}

	\section{Problem 1 --- Reusable bottle participation in Grade 11}
	The student council wants to know whether the true proportion of Grade 11 students who bring a reusable water bottle is now above the campaign benchmark of 60\%. During one morning check, they record a sample of \(n=150\) students and observe \(x=99\) students with reusable bottles.
	
	\textbf{Questions/Tasks.}
	\begin{enumerate}
		\item[(C1)] Identify whether this is a proportion or mean setting, define the parameter, and map the given sample information.
		\item[(C2)] Write the null and alternative hypotheses for the claim.
		\item[(C3)] Classify the tail type and justify it from the context.
	\end{enumerate}
	
	\subsection*{C1}
	This is a population proportion setting. Let \(p\) be the true proportion of Grade 11 students who bring a reusable bottle on a regular school day. The benchmark is \(p_0=0.60\). The sample information is \(n=150\), \(x=99\), and \(\hat p=99/150=0.66\).
	
	\subsection*{C2}
	\[
	H_0\colon p=0.60, \qquad H_a\colon p>0.60.
	\]
	In words, the null keeps the historical benchmark at 60\%, and the alternative represents the campaign claim that the proportion is higher than 60\%.
	
	\subsection*{C3}
	This is a right-tailed test because the claim is ``above 60\%.'' The alternative uses \(>\), so evidence is looked for in the upper tail.


	\section{Problem 2 --- On-time first submission rate in economics class}
	The teacher tracks whether students submit the first weekly task on time. The department benchmark says the true on-time proportion should be at least 85\%, and the teacher worries that the real rate may have dropped below that level. In a sample of \(n=120\) students, \(x=94\) submitted on time.
	
	\textbf{Questions/Tasks.}
	\begin{enumerate}
		\item[(C1)] Identify the parameter type and map the symbols in context.
		\item[(C2)] Write the null and alternative hypotheses.
		\item[(C3)] Classify and justify the tail type.
	\end{enumerate}
	
	\subsection*{C1}
	This is a population proportion setting. Let \(p\) be the true proportion of students who submit the first weekly task on time. The benchmark is \(p_0=0.85\). The sample summary is \(n=120\), \(x=94\), and \(\hat p=94/120\approx0.783\).
	
	\subsection*{C2}
	\[
	H_0\colon p=0.85, \qquad H_a\colon p<0.85.
	\]
	The null represents the benchmark level, and the alternative captures the concern that punctual submission is lower than the target.
	
	\subsection*{C3}
	This is a left-tailed test because the question is whether the true on-time proportion is below the benchmark. The alternative uses \(<\).


	\section{Problem 3 --- Average waiting time in the school cafeteria}
	After changing lunch-line organization, school leaders want to verify whether the true mean waiting time is now less than 9 minutes. A time study of \(n=64\) students reports a sample mean of \(\bar{x}=8.2\) minutes.
	
	\textbf{Questions/Tasks.}
	\begin{enumerate}
		\item[(C1)] Identify whether this is a mean or proportion setting and define the parameter.
		\item[(C2)] Write the null and alternative hypotheses for the claim.
		\item[(C3)] Determine and justify the test tail type.
	\end{enumerate}
	
	\subsection*{C1}
	This is a population mean setting. Let \(\mu\) be the true mean cafeteria waiting time (in minutes) for students during lunch break. The benchmark is \(\mu_0=9\). The sample information is \(n=64\) and \(\bar{x}=8.2\).
	
	\subsection*{C2}
	\[
	H_0\colon \mu=9, \qquad H_a\colon \mu<9.
	\]
	The null fixes the benchmark at 9 minutes, and the alternative reflects the improvement claim that average waiting time is lower.
	
	\subsection*{C3}
	This is a left-tailed test because the directional claim is ``less than 9 minutes.'' The alternative is expressed with \(<\).


	\section{Problem 4 --- Mean daily study time after a planning workshop}
	A Grade 11 coordinator wants to check whether the planning workshop changed average daily study time from the historical benchmark of 2.5 hours. A follow-up sample of \(n=50\) students gives \(\bar{x}=2.9\) hours.
	
	\textbf{Questions/Tasks.}
	\begin{enumerate}
		\item[(C1)] Identify the parameter type and map symbols.
		\item[(C2)] Write the null and alternative hypotheses.
		\item[(C3)] Classify and justify the tail type.
	\end{enumerate}
	
	\subsection*{C1}
	This is a population mean setting. Let \(\mu\) be the true mean daily study time (in hours) for Grade 11 students after the workshop. The benchmark is \(\mu_0=2.5\), and the sample information includes \(n=50\) and \(\bar{x}=2.9\).
	
	\subsection*{C2}
	\[
	H_0\colon \mu=2.5, \qquad H_a\colon \mu\neq2.5.
	\]
	The null states no change from the benchmark. The alternative uses ``different from'' because both increases and decreases represent a change.
	
	\subsection*{C3}
	This is a two-tailed test because the claim is about being different from 2.5 hours, not specifically higher or lower. The alternative uses \(\neq\).


	\section{Problem 5 --- Campus shuttle punctuality claim (direction check)}
	The transport office reports that at least 70\% of morning shuttle trips arrive on time. A student auditor says the real punctuality may be lower than that benchmark. In a review of \(n=180\) trips, \(x=118\) were on time.
	
	\textbf{Questions/Tasks.}
	\begin{enumerate}
		\item[(C1)] Identify the parameter and classify the setting.
		\item[(C2)] Write a hypothesis pair that matches the auditor's directional claim.
		\item[(C3)] Classify the tail type and justify why this is the correct direction.
	\end{enumerate}
	
	\subsection*{C1}
	This is a population proportion setting. Let \(p\) be the true proportion of morning shuttle trips that arrive on time. The benchmark is \(p_0=0.70\). The sample data are \(n=180\), \(x=118\), so \(\hat p=118/180\approx0.656\).
	
	\subsection*{C2}
	\[
	H_0\colon p=0.70, \qquad H_a\colon p<0.70.
	\]
	This pair matches the auditor's concern that punctuality is lower than the official benchmark.
	
	\subsection*{C3}
	Left-tailed test. The direction comes from the phrase ``may be lower,'' which corresponds to \(p<0.70\). A right-tailed or two-tailed test would not represent the stated concern as directly.


	\section{Problem 6 --- ``I am almost there'' travel-time message audit}
	A Grade 11 group studies travel messages before meetups. They define waiting time as minutes between a student texting ``I am almost there'' and actually arriving. They want to evaluate whether the true mean waiting time is greater than the practical benchmark of 7 minutes. Over several meetups, they collect \(n=55\) records and obtain \(\bar{x}=8.1\) minutes.
	
	\textbf{Questions/Tasks.}
	\begin{enumerate}
		\item[(C1)] Identify the parameter and context setup.
		\item[(C2)] Write null and alternative hypotheses in notation.
		\item[(C3)] Classify and justify the test tail.
	\end{enumerate}
	
	\subsection*{C1}
	This is a population mean setting. Let \(\mu\) be the true mean waiting time (in minutes) between sending ``I am almost there'' and arriving at the meeting point. The benchmark is \(\mu_0=7\), with sample information \(n=55\) and \(\bar{x}=8.1\).
	
	\subsection*{C2}
	\[
	H_0\colon \mu=7, \qquad H_a\colon \mu>7.
	\]
	The null keeps the practical benchmark at 7 minutes. The alternative formalizes the claim that the average delay is greater than that value.
	
	\subsection*{C3}
	Right-tailed test, because the claim is specifically that the true mean waiting time is greater than 7 minutes. The alternative uses \(>\), so the rejection region is in the upper tail.

\end{document}
