\input{../../../preamble/preamble_exams.tex}

\begin{document}
\ExamTitleBlock{11th grade}{Term 3 --- Lesson 1 Summary of Practice Activities}{ }

\section*{Contents}
\addcontentsline{toc}{section}{Contents}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{1}
\tableofcontents

\section{Summary of \texttt{G11\_T3\_L1\_C1C2C3\_practice\_activity1.tex}}
This activity develops foundational work for Criteria C1--C3 by helping students identify whether a context uses a population mean or proportion, assign correct symbols, and write hypotheses in all three possible directions (lower, higher, different). Its purpose is introductory: students first build a stable template for parameter notation and then connect each alternative-hypothesis direction to left-, right-, or two-tailed testing. The reasoning focus is classification and structure recognition before full procedural testing.

\begin{itemize}
    \item \textbf{Daily app check-ins for students and recent graduates:} Introductory proportion problem where students map context values to \(p, p_0, n, x, \alpha\), then write all three directional hypothesis pairs and classify the corresponding tail types.
    \item \textbf{Evening study time during a demanding term:} Parallel mean-based setup (with known \(\sigma\)) that reinforces the same three-direction framework, now transferring symbolic reasoning from proportions to means.
    \item \textbf{Defect rate in portable chargers:} Proportion-focused quality-control context that strengthens directional hypothesis writing under a stricter significance level, emphasizing precision in claim wording.
    \item \textbf{Homework completion time in an evening certification program:} Mean-context consolidation problem in which students repeat the full C1--C3 cycle and verify that tail classification is determined by the alternative-hypothesis direction, not by topic context.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C1C2C3\_practice\_activity2.tex}}
This redesigned follow-up activity structures C1--C3 work as a clear progression from contextual direction selection to evidence-based conclusion making. Students first identify the correct one-direction formulation from context and complete full hypothesis-test decisions using reject \(H_0\) or fail to reject \(H_0\). They then examine how conclusions can change when significance levels change while hypotheses and tail direction remain fixed. The activity ends with paired comparisons on identical data to show explicitly how choosing a one-sided versus two-sided alternative can change p-values and decisions.

\begin{itemize}
    \item \textbf{Problem 1 --- Math-support program and average algebra score:} Lower-tailed mean context where students identify \(\mu\), write \(H_0: \mu=\mu_0\) and \(H_a: \mu<\mu_0\), compute test evidence, and reach a reject-\(H_0\) decision.
    \item \textbf{Problem 2 --- Delivery app on-time rate after route training:} Higher-tailed proportion context where students map \(p\)-notation, test \(H_a: p>p_0\), and obtain a reject-\(H_0\) decision from the sample evidence.
    \item \textbf{Problem 3 --- Clinic waiting time consistency check:} Two-tailed mean context for a ``different'' claim, producing a fail-to-reject-\(H_0\) outcome so students see contrasting decisions across Stage 1.
    \item \textbf{Problem 4 --- Fuel efficiency review under different policy strictness:} Lower-tailed mean test where hypotheses and tail stay fixed, but conclusions are compared across multiple \(\alpha\) levels motivated by routine monitoring versus stricter accountability.
    \item \textbf{Problem 5 --- Scholarship completion-rate target:} Higher-tailed proportion test where the same evidence is interpreted under internal-planning and public-reporting significance standards, showing threshold-sensitive decisions.
    \item \textbf{Problem 6 --- Factory packaging-weight stability check:} Two-tailed mean quality-control test that keeps the ``different'' formulation fixed while contrasting conclusions across looser and stricter quality thresholds.
    \item \textbf{Problem 7 --- App response time (lower vs different):} Same context, same data, and same \(\alpha\) comparison of left-tailed and two-tailed formulations to highlight how alternative-hypothesis framing changes p-values and can affect rejection strength.
    \item \textbf{Problem 8 --- Daily study-time claim (higher vs different):} Same-data comparison of right-tailed and two-tailed formulations at one significance level where both tests fail to reject \(H_0\), reinforcing meaningful non-significance interpretation under different alternatives.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C4\_practice\_activity.tex}}
This file develops Criterion C4 as a deliberate four-stage hypothesis-testing ladder. Students first consolidate the complete test procedure, then investigate how tail direction and significance level alter decisions, then compare multiple alternatives on the same evidence (including near-boundary cases), and finally synthesize conclusions across several test framings in one context. Across the activity, each problem requires parameter identification, hypothesis formulation, condition checks where needed, test-statistic and p-value reasoning, comparison with \(\alpha\), and contextual interpretation.

\begin{itemize}
    \item \textbf{General Problem (Stage 1 --- full procedure foundation):} Builds a reusable full-test template for one-sample proportion and mean \(z\)-tests, including parameter setup, directional hypotheses, conditions, statistics, p-values, decisions, and contextual conclusions.
    \item \textbf{Food delivery completion rate (Stage 1 --- introductory full proportion test):} Straightforward lower-tail proportion application that walks through every core step from condition checks to contextual conclusion.
    \item \textbf{Average one-way commute time for office staff (Stage 1 --- introductory full mean test):} Straightforward higher-tail mean \(z\)-test that reinforces the same full C4 sequence in a mean context.
    \item \textbf{Public health screening attendance benchmark (Stage 2 --- direction sensitivity, lower vs higher):} Same sample evidence tested with opposite one-tailed proportion alternatives to show how direction changes tail area, p-value, and decision.
    \item \textbf{Average packaging time in a school enterprise lab (Stage 2 --- one-tailed mean direction sensitivity):} Same mean sample tested as lower-tail and higher-tail to make direction-driven decision differences explicit.
    \item \textbf{Average monthly public transport spending benchmark (Stage 2 --- significance-level sensitivity):} One two-tailed mean test interpreted at 5\% and 1\% to show that unchanged evidence can yield different decisions when only \(\alpha\) changes.
    \item \textbf{Weekly side-income amount from freelance tasks (Stage 2 --- robust two-tailed evidence across \(\alpha\)):} Two-tailed mean test where conclusions remain reject/reject at both 5\% and 1\%, contrasting with threshold-sensitive cases.
    \item \textbf{Use of monthly budgeting reminders: lower vs higher (Stage 3 --- comparative directional reasoning):} Proportion case that compares lower- and higher-tail tests on identical data, illustrating opposite-tail p-values and matched non-rejection outcomes.
    \item \textbf{Average evening study time: lower vs different (Stage 3 --- comparative directional reasoning):} Mean case showing lower-tail versus two-tailed framing on the same statistic, with both tests failing to reject.
    \item \textbf{Customer support wait-time benchmark: higher vs different (Stage 3 --- comparative directional reasoning):} Mean case showing higher-tail and two-tailed tests both rejecting while the opposite directional interpretation is unsupported.
    \item \textbf{Use of monthly budgeting reminders with near-boundary evidence (Stage 3 --- near-boundary comparison of higher/lower/different):} Three-way proportion comparison near \(\alpha=5\%\) where all tests fail to reject, highlighting careful boundary interpretation.
    \item \textbf{Battery-life compliance synthesis under multiple framings (Stage 4 --- strongest synthesis):} One context combining higher, lower, and two-tailed mean tests with dual significance-level interpretation to explain why conclusions differ across framing and strictness.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C5\_practice\_activity.tex}}
This activity is organized as a deliberate learning ladder for one-sample inference on a population mean with known \(\sigma\), moving from full-procedure foundations to direction-sensitive and significance-sensitive comparisons, and ending with synthesis cases that require students to explain why decisions differ across hypotheses even when the same sample evidence is used.

\begin{itemize}
    \item \textbf{General Problem (Stage 1 --- full workflow template):} Establishes the complete one-sample mean \(z\)-test procedure by identifying \(\mu\), writing directional hypotheses, computing \(z\), interpreting p-values, comparing with \(\alpha\), and concluding in context.
    \item \textbf{Average monthly spending on phone plans (Stage 1 --- introductory right-tailed application):} First direct right-tailed mean test using the full procedure to reinforce benchmark-above claims.
    \item \textbf{Average weekly grocery budget (Stage 1 --- introductory left-tailed application):} First direct left-tailed mean test using the same machinery for benchmark-below claims.
    \item \textbf{Average monthly commuting cost (Stage 1 --- introductory two-tailed application):} First direct two-tailed mean test to formalize difference-from-benchmark reasoning.
    \item \textbf{Average sodium intake in ready meals (Stage 3 --- left-tailed significance-level sensitivity):} Holds sample evidence fixed while comparing decisions at two \(\alpha\)-levels in a health-caution setting.
    \item \textbf{Average monthly savings in emergency funds (Stage 3 --- right-tailed significance-level sensitivity):} Uses one dataset and two significance levels to show threshold-sensitive conclusions in financial planning.
    \item \textbf{Average mathematics benchmark score (Stage 3 --- two-tailed significance-level sensitivity):} Demonstrates that unchanged data can lead to different two-tailed conclusions when academic standards use different \(\alpha\)-levels.
    \item \textbf{Average monthly gym visits (Stage 4 --- lower vs higher on the same evidence):} Compares opposite one-tailed alternatives to show direction-driven p-value asymmetry.
    \item \textbf{Average focus hours in deep-work sessions (Stage 4 --- lower vs different on the same evidence):} Contrasts left-tailed and two-tailed formulations to reveal how tail choice changes rejection outcomes.
    \item \textbf{Average monthly commuting distance (Stage 4 --- higher vs different on the same evidence):} Compares right-tailed and two-tailed tests to show when both directions support rejection under strong evidence.
    \item \textbf{Average monthly streaming subscriptions cost (Stage 4 --- lower vs higher vs different):} Three-direction comparison where all tests fail to reject, reinforcing coherent non-signal interpretation.
    \item \textbf{Average response time in a customer support center (Stage 5 --- one-direction-only rejection):} Advanced synthesis case where only the context-appropriate directional test rejects while the opposite direction and two-tailed test do not.
    \item \textbf{Average number of sick days used per year (Stage 5 --- strong-signal directional synthesis):} Advanced synthesis case where one directional test and the corresponding two-tailed test reject, while the opposite direction fails to reject, requiring explicit explanation of the evidence pattern.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C6\_practice\_activity.tex}}
This activity targets sample-size, margin-of-error, and interpretation-oriented reasoning through contextual decisions from reported test outcomes, especially when p-values are compared to different significance thresholds. Its pedagogical purpose is to train students to read statistical decisions carefully in policy and planning contexts, distinguish evidence from certainty, and articulate Type I/Type II implications. The reasoning is interpretive, comparative, and decision-theoretic.

\begin{itemize}
    \item \textbf{General Problem:} Framework problem on interpreting reject/fail-to-reject outcomes, threshold comparisons, and contextual error meanings without recalculating statistics.
    \item \textbf{Fuel-price monitoring report:} Contextual interpretation of a rejection in a right-tailed mean test, including explicit Type I error translation.
    \item \textbf{Public employment office review:} Fail-to-reject interpretation in a lower-tailed mean setting, with emphasis on what remains unproven and Type II risk.
    \item \textbf{Export quality audit with two evidence thresholds:} Single p-value interpreted at 5\% and 1\% to highlight robust rejection across stricter standards.
    \item \textbf{Digital payment adoption check under strict and lenient policy standards:} Threshold-sensitive case where conclusions differ at 1\% and 5\%.
    \item \textbf{Border operations dashboard: two independent service tests:} Multi-metric interpretation task separating conclusions for a two-tailed mean test and a one-tailed proportion test.
    \item \textbf{Education and labor program review: two different directional checks:} Dual-context interpretation with directional differences and concrete policy-error narrative.
    \item \textbf{Economic development briefing with mixed thresholds:} Comparative synthesis identifying one threshold-sensitive and one robust conclusion across independent tests.
    \item \textbf{Telecom service regulation: three independent compliance checks:} Multi-test dashboard interpretation balancing reject and fail-to-reject outcomes across domains.
    \item \textbf{Integrated transport and labor assessment with robustness checks:} Three-test robustness analysis at 10\% and 5\%, classifying stable vs sensitive findings.
    \item \textbf{National performance synthesis: three tests, multiple confidence standards:} Advanced cross-threshold reporting exercise that categorizes strong, moderate, and tentative evidence.
\end{itemize}

\section{Summary of \texttt{G11\_T3\_L1\_C7\_practice\_activity.tex}}
This file is redesigned as a five-stage pedagogical ladder for one-sample confidence intervals of a population proportion, linking earlier hypothesis-testing ideas to interval reasoning while keeping the central work on \(\hat p\), margin of error, confidence level, precision, and contextual interpretation. Students progress from foundational setup and interpretation, to confidence-level and sample-size sensitivity, then to benchmark plausibility decisions, and finally to multi-case synthesis that classifies evidence strength without reducing the activity to p-value decision rules.

\begin{itemize}
    \item \textbf{General Problem (Stage 1 --- full confidence-interval workflow):} Introduces the complete proportion-interval procedure: parameter identification, \(\hat p\), condition checks, interval construction, and non-certainty interpretation in context.
    \item \textbf{Product quality pass rate (Stage 1 --- first direct computation):} Applies the full workflow to compute a 95\% confidence interval for a quality proportion and interpret it operationally.
    \item \textbf{Service completion in a municipal office (Stage 1 --- interpretation focus):} Emphasizes correct contextual meaning of interval endpoints and what the interval does not guarantee.
    \item \textbf{App feature adoption and parameter distinction (Stage 1 --- sample vs population):} Clarifies the difference between sample proportion \(\hat p\) and unknown population proportion \(p\), then interprets the interval as an estimate for \(p\).
    \item \textbf{Payment adoption at three confidence levels (Stage 2 --- 90\%, 95\%, 99\% comparison):} Uses one dataset to show how increasing confidence widens intervals and increases margin of error.
    \item \textbf{Contactless payment reliability and interval width (Stage 2 --- width and margin-of-error analysis):} Compares widths and precision explicitly across confidence levels for the same \(\hat p\) and \(n\).
    \item \textbf{Screening compliance and confidence-level choice (Stage 2 --- decision tradeoff):} Uses context to justify selecting higher or lower confidence based on the confidence--precision tradeoff.
    \item \textbf{Retention estimation under two sample sizes (Stage 3 --- same \(\hat p\), different \(n\)):} Demonstrates how larger samples reduce standard error and narrow the interval around the same center.
    \item \textbf{Delivery reliability and stability of estimation (Stage 3 --- why larger \(n\) helps):} Requires explanation of why increasing sample size improves precision and interpretive reliability.
    \item \textbf{Planning sample size for program participation (Stage 3 --- design for target margin of error):} Connects interval precision to data-collection planning by solving for required \(n\).
    \item \textbf{Program participation benchmark plausibility (Stage 4 --- benchmark inside interval):} Uses interval inclusion to interpret a target proportion as plausible under current evidence.
    \item \textbf{On-time delivery target check (Stage 4 --- benchmark outside interval):} Uses interval exclusion to interpret a benchmark as unsupported, with a careful link to earlier testing logic.
    \item \textbf{Survey approval with two nearby benchmarks (Stage 4 --- mixed plausibility and tentative evidence):} Compares two close benchmarks and highlights how wide intervals can make conclusions less decisive.
    \item \textbf{Policy dashboard I: quality, adoption, and completion (Stage 5 --- three-case synthesis):} Compares multiple intervals, classifies evidence as strong/moderate/tentative, and evaluates target attainment across domains.
    \item \textbf{Regional digital-payment dashboard (Stage 5 --- reliability and practical consequences):} Synthesizes interval width, benchmark checks, and action priorities across regions with different estimate precision.
    \item \textbf{National performance report synthesis (Stage 5 --- strongest integrated interpretation):} Produces a report-style judgment across several proportion intervals, separating robust findings from uncertain ones and explaining implications of narrow versus wide intervals.
\end{itemize}

\end{document}
