\input{../../../preamble/preamble_exams.tex}

\begin{document}

	\newcommand{\LeftTailGraph}{%
		\begin{center}
		\begin{tikzpicture}[scale=0.55]
		\draw[thin] (-3,0) -- (3,0);
		\fill[gray!35] (-3,0) -- plot[smooth,domain=-3:-1] (\x,{1.6*exp(-0.5*\x*\x)}) -- (-1,0) -- cycle;
		\draw[thick,domain=-3:3,smooth] plot (\x,{1.6*exp(-0.5*\x*\x)});
		\end{tikzpicture}
		\end{center}
	}

	\newcommand{\RightTailGraph}{%
		\begin{center}
		\begin{tikzpicture}[scale=0.55]
		\draw[thin] (-3,0) -- (3,0);
		\fill[gray!35] (1,0) -- plot[smooth,domain=1:3] (\x,{1.6*exp(-0.5*\x*\x)}) -- (3,0) -- cycle;
		\draw[thick,domain=-3:3,smooth] plot (\x,{1.6*exp(-0.5*\x*\x)});
		\end{tikzpicture}
		\end{center}
	}

	\newcommand{\TwoTailGraph}{%
		\begin{center}
		\begin{tikzpicture}[scale=0.55]
		\draw[thin] (-3,0) -- (3,0);
		\fill[gray!35] (-3,0) -- plot[smooth,domain=-3:-1] (\x,{1.6*exp(-0.5*\x*\x)}) -- (-1,0) -- cycle;
		\fill[gray!35] (1,0) -- plot[smooth,domain=1:3] (\x,{1.6*exp(-0.5*\x*\x)}) -- (3,0) -- cycle;
		\draw[thick,domain=-3:3,smooth] plot (\x,{1.6*exp(-0.5*\x*\x)});
		\end{tikzpicture}
		\end{center}
	}
	\ExamTitleBlock{11th grade}{Term 3 — Lesson 1 Practice Activity 2 (C1–C3) with Solutions}{}

	\section*{Contents}
	\addcontentsline{toc}{section}{Contents}
	\setcounter{tocdepth}{2}
	\setcounter{secnumdepth}{1}
	\tableofcontents
	
	\ExamSection{Evaluated criteria}
	\begin{ExamCriteria}
		\item C1: Describe the concept of a statistical hypothesis in context.
		\item C2: Identify the null and alternative hypotheses using correct parameter notation.
		\item C3: Determine whether a test is one tailed or two tailed based on the claim.
	\end{ExamCriteria}
	
	\ExamSection{Introduction}
	This activity focuses on contextualized hypothesis tests where each problem has one intended claim direction.

	For every problem, use the context to select one appropriate alternative hypothesis (lower, higher, or different). Keep the null hypothesis as equality with the benchmark value, and classify a single matching tail type.

	\subsection*{C1}
	\begin{itemize}
		\item Identify whether the problem is about a population proportion or a population mean.
		\item Match the benchmark value, sample information, and significance level to the correct symbols.
		\item Use the context to determine the intended direction of the alternative hypothesis.
	\end{itemize}

	\subsection*{C2}
	\begin{itemize}
		\item Write one null--alternative pair that matches the contextual claim direction.
		\item Keep the null hypothesis as \(\text{parameter}=\text{benchmark}\).
		\item Use the correct inequality in the alternative hypothesis: \(<\), \(>\), or \(\ne\) as required by the context.
	\end{itemize}

	\subsection*{C3}
	\begin{itemize}
		\item Classify one tail type for each problem and justify it from the alternative claim.
		\item Use left-tailed for ``lower``, right-tailed for ``higher``, and two-tailed for ``different``.
	\end{itemize}

\begin{ExamProblems}

		\item
		\section{Secure login success in a budgeting app}
		A personal budgeting app recently replaced its login flow with a new security verification step. After the update, some users reported both failed attempts and, in other cases, smoother access on newer devices. The product team now wants to check whether the first-attempt login success rate has changed from the historical benchmark, since any shift could affect support costs and user retention. The platform benchmark for first-attempt login success is 91\%. In a sample of 320 attempts, 301 are successful. They will use a 5\% significance level for this check. (C1) Identify whether the claim is about a population proportion or a population mean, and match the parameter, the benchmark value of 91\%, the sample size, and the observed sample result for the first-attempt login success proportion to the symbols used in the solution. (C2) Using the 5\% significance level, write the null and alternative hypotheses for the case where the true first-attempt login success proportion is different from 91\%. (C3) Classify the test and justify it in words for the case where the true first-attempt login success proportion is different from 91\%.
		
		\subsection*{C1}
		\begin{itemize}
			\item This is a population proportion setting.
			\item Let \(p\) be the true population proportion of successful first-attempt logins across all attempts.
			\item Let \(p_0\) be the benchmark success proportion, with \(p_0=91\%\).
			\item Let \(n\) be the sample size, with \(n=320\).
			\item Let \(x\) be the observed number of successful first-attempt logins, with \(x=301\).
			\item Let \(\alpha\) be the significance level, with \(\alpha=5\%\).
			\item The alternative claim in this problem is that the proportion is different from the benchmark.
		\end{itemize}
		
		\subsection*{C2}
		\(H_0\colon p=0.91,\; H_a\colon p\ne 0.91\), \(\alpha=0.05\). In words, the null hypothesis says that the population proportion equals the benchmark, while the alternative says that the population proportion is different from the benchmark.
		
		\subsection*{C3}
		Two-tailed test because the alternative claims the parameter is different from the benchmark.\TwoTailGraph
		% --------------------------------------------------
		

		\item
		\section{Commute time to campus}
		A district transportation office is monitoring whether a new express-bus plan is meeting its promise to reduce student travel time, especially for students who attend classes while working part time. The plan was designed to keep the average commute at or below the district benchmark of 32 minutes; if the average is not lower, the office may need to redesign routes before the next term. A sample of 40 students gives an average of 34 minutes. Assume the population standard deviation is known to be 8 minutes. To be stricter because transportation subsidies are affected, they will use a 1\% significance level for this check. (C1) Identify whether the claim is about a population proportion or a population mean, and match the parameter, the benchmark value of 32 minutes, the sample size, and the observed sample result for the mean commute time to the symbols used in the solution. (C2) Using the 1\% significance level, write the null and alternative hypotheses for the case where the true mean commute time is lower than 32 minutes. (C3) Classify the test and justify it in words for the case where the true mean commute time is lower than 32 minutes.
		
		\subsection*{C1}
		\begin{itemize}
			\item This is a population mean setting.
			\item Let \(\mu\) be the true population mean commute time for all students in the district.
			\item Let \(\mu_0\) be the benchmark estimate, with \(\mu_0=32\).
			\item Let \(n\) be the sample size, with \(n=40\).
			\item Let \(\bar{x}\) be the sample mean commute time, with \(\bar{x}=34\).
			\item Let \(\sigma\) be the known population standard deviation, with \(\sigma=8\).
			\item Let \(\alpha\) be the significance level, with \(\alpha=1\%\).
			\item The alternative claim in this problem is that the mean is lower than the benchmark.
		\end{itemize}
		
		\subsection*{C2}
		\(H_0\colon \mu=32,\; H_a\colon \mu<32\), \(\alpha=0.01\). In words, the null hypothesis says that the population mean equals the benchmark, while the alternative says that the population mean is lower than the benchmark.
		
		\subsection*{C3}
		Left-tailed test because the alternative claims the parameter is lower than the benchmark.\LeftTailGraph
		% --------------------------------------------------
		

		\item
		\section{Cafeteria card usage}
		A university finance office is reviewing cafeteria payment patterns after introducing a larger card-only meal discount this term. The office is specifically concerned that card usage might now exceed the usual planning benchmark, because a higher-than-expected card share would increase transaction fees and change next semester's contract terms. The benchmark proportion of purchases paid with student cards is 70\%. In a sample of 300 purchases, 198 are paid with student cards. Because the next contract decision affects a limited budget, they will use a 2\% significance level for this check. (C1) Identify whether the claim is about a population proportion or a population mean, and match the parameter, the benchmark value of 70\%, the sample size, and the observed sample result for the card-payment proportion to the symbols used in the solution. (C2) Using the 2\% significance level, write the null and alternative hypotheses for the case where the true card-payment proportion is higher than 70\%. (C3) Classify the test and justify it in words for the case where the true card-payment proportion is higher than 70\%.
		
		\subsection*{C1}
		\begin{itemize}
			\item This is a population proportion setting.
			\item Let \(p\) be the true population proportion of cafeteria purchases paid with student cards.
			\item Let \(p_0\) be the benchmark card-payment proportion, with \(p_0=70\%\).
			\item Let \(n\) be the sample size, with \(n=300\).
			\item Let \(x\) be the observed number of purchases paid with student cards, with \(x=198\).
			\item Let \(\alpha\) be the significance level, with \(\alpha=2\%\).
			\item The alternative claim in this problem is that the proportion is higher than the benchmark.
		\end{itemize}
		
		\subsection*{C2}
		\(H_0\colon p=0.70,\; H_a\colon p>0.70\), \(\alpha=0.02\). In words, the null hypothesis says that the population proportion equals the benchmark, while the alternative says that the population proportion is higher than the benchmark.
		
		\subsection*{C3}
		Right-tailed test because the alternative claims the parameter is higher than the benchmark.\RightTailGraph
		% --------------------------------------------------
		

		\item
		\section{Quiz score average in a career-focused course}
		A department is reviewing a career-focused course after replacing several quiz items and adjusting the grading rubric midterm. Because of these changes, staff are not assuming scores would move only up or only down; they need to check whether the typical quiz score is still aligned with the long-run target. The target is an average quiz score of 68 points. A sample of 25 students has an average score of 71 points. Assume the population standard deviation is known to be 9 points. Because tutoring hours and review sessions may be expanded under stricter or more flexible policies, the department will consider three significance levels for this check: 1\%, 2\%, and 5\%. (C1) Identify whether the claim is about a population proportion or a population mean, and match the parameter, the benchmark value of 68 points, the sample size, the observed sample result, and the three stated significance levels to the symbols used in the solution. (C2) For each stated significance level, write the null and alternative hypotheses for the case where the true mean quiz score is different from 68 points. (C3) Classify the test and justify it in words for the case where the true mean quiz score is different from 68 points, and state whether the tail type changes across the stated significance levels.
		
		\subsection*{C1}
		\begin{itemize}
			\item This is a population mean setting.
			\item Let \(\mu\) be the true population mean quiz score for all students in the department.
			\item Let \(\mu_0\) be the benchmark target mean, with \(\mu_0=68\).
			\item Let \(n\) be the sample size, with \(n=25\).
			\item Let \(\bar{x}\) be the sample mean quiz score, with \(\bar{x}=71\).
			\item Let \(\sigma\) be the known population standard deviation, with \(\sigma=9\).
			\item Let \(\alpha\) represent the significance level, and here \(\alpha\) will be considered at 1\%, 2\%, and 5\%.
			\item The alternative claim in this problem is that the mean is different from the benchmark.
		\end{itemize}
		
		\subsection*{C2}
		\textbf{At \(\alpha=0.01\):} \(H_0\colon \mu=68,\; H_a\colon \mu\ne 68\).
		
		\textbf{At \(\alpha=0.02\):} \(H_0\colon \mu=68,\; H_a\colon \mu\ne 68\).
		
		\textbf{At \(\alpha=0.05\):} \(H_0\colon \mu=68,\; H_a\colon \mu\ne 68\).
		
		In words, for all three significance levels, the null hypothesis says that the population mean equals the benchmark, while the alternative says that the population mean is different from the benchmark. The hypotheses stay the same; only the significance level changes.
		
		\subsection*{C3}
		This is a two-tailed test at all three significance levels because the alternative hypothesis is \(H_a\colon \mu\ne 68\), and tail type is determined by the alternative hypothesis. As \(\alpha\) changes from 0.01 to 0.02 to 0.05, the test stays two-tailed, but the shaded rejection regions in both tails become wider for larger \(\alpha\) and narrower for smaller \(\alpha\).
		
		\begin{center}
		\begin{minipage}{0.3\textwidth}
		\centering
		\textbf{\(\alpha=0.01\)}
		\begin{tikzpicture}[scale=0.45]
		\draw[thin] (-3,0) -- (3,0);
		\fill[gray!35] (-3,0) -- plot[smooth,domain=-3:-2.35] (\x,{1.6*exp(-0.5*\x*\x)}) -- (-2.35,0) -- cycle;
		\fill[gray!35] (2.35,0) -- plot[smooth,domain=2.35:3] (\x,{1.6*exp(-0.5*\x*\x)}) -- (3,0) -- cycle;
		\draw[thick,domain=-3:3,smooth] plot (\x,{1.6*exp(-0.5*\x*\x)});
		\end{tikzpicture}
		\end{minipage}
		\begin{minipage}{0.3\textwidth}
		\centering
		\textbf{\(\alpha=0.02\)}
		\begin{tikzpicture}[scale=0.45]
		\draw[thin] (-3,0) -- (3,0);
		\fill[gray!35] (-3,0) -- plot[smooth,domain=-3:-2.1] (\x,{1.6*exp(-0.5*\x*\x)}) -- (-2.1,0) -- cycle;
		\fill[gray!35] (2.1,0) -- plot[smooth,domain=2.1:3] (\x,{1.6*exp(-0.5*\x*\x)}) -- (3,0) -- cycle;
		\draw[thick,domain=-3:3,smooth] plot (\x,{1.6*exp(-0.5*\x*\x)});
		\end{tikzpicture}
		\end{minipage}
		\begin{minipage}{0.3\textwidth}
		\centering
		\textbf{\(\alpha=0.05\)}
		\begin{tikzpicture}[scale=0.45]
		\draw[thin] (-3,0) -- (3,0);
		\fill[gray!35] (-3,0) -- plot[smooth,domain=-3:-1.75] (\x,{1.6*exp(-0.5*\x*\x)}) -- (-1.75,0) -- cycle;
		\fill[gray!35] (1.75,0) -- plot[smooth,domain=1.75:3] (\x,{1.6*exp(-0.5*\x*\x)}) -- (3,0) -- cycle;
		\draw[thick,domain=-3:3,smooth] plot (\x,{1.6*exp(-0.5*\x*\x)});
		\end{tikzpicture}
		\end{minipage}
		\end{center}
		% --------------------------------------------------
		


	\end{ExamProblems}
\end{document}
