\input{../../../preamble/preamble_exams.tex}

\begin{document}
\ExamTitleBlock{11th grade}{Term 3 --- Lesson 1 Practice Activity (C6) with Solutions}{}

\section*{Contents}
\addcontentsline{toc}{section}{Contents}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{1}
\tableofcontents

\ExamSection{Evaluated criteria}
\begin{ExamCriteria}
    \item C6: Interprets decisions from hypothesis tests in terms of the context of the problem.
\end{ExamCriteria}

\ExamSection{Problems}
\begin{ExamProblems}

\item
\section{Fuel-price monitoring report}
The city transport authority sets bus-fare reserves using a fuel-inflation figure approved in the municipal budget. That budget figure is 4.0\% per month, based on last year's treasury forecast. Before the next fare review meeting, the analytics unit sampled filling stations from all operating bus corridors and tested whether the citywide mean monthly fuel inflation is above the 4.0\% planning value. The test setup was \(H_0\colon \mu=4.0\), \(H_1\colon \mu>4.0\), at \(\alpha=0.05\), and the reported decision was \textit{reject \(H_0\)}. Interpret this decision for the fare-planning team: what are they justified in concluding about average inflation, what should they avoid claiming about certainty and about every fuel category, and what specific mistaken announcement would correspond to a Type I error in this setting?

\subsection*{C6}
\begin{itemize}
    \item The decision supports the claim that the population mean inflation is greater than 4.0\%.
    \item It does not prove with certainty that \(\mu>4.0\); there is still Type I error risk, meaning officials could incorrectly announce inflation is above 4.0\% when the true citywide mean is actually 4.0\%.
    \item It does not say every product category has inflation above 4.0\%; the result is about the population mean.
\end{itemize}

\item
\section{Public employment office review}
A labor ministry is deciding whether to expand a new job-matching platform nationwide. The baseline used in the ministry's annual plan is an average unemployment spell of 11 weeks, computed from the previous three years of administrative records. This quarter, analysts pulled a random sample of completed claimant files from pilot regions to check whether the new system has lowered the average duration below that benchmark. They tested \(H_0\colon \mu=11\), \(H_1\colon \mu<11\), with \(\alpha=0.10\), and obtained \textit{fail to reject \(H_0\)}. Explain the correct policy conclusion in plain language, state clearly what cannot be concluded from this outcome, and describe concretely what a Type II error would look like here (that is, what true improvement the ministry might miss).

\subsection*{C6}
\begin{itemize}
    \item The data do not provide sufficient evidence to support the claim that mean duration is below 11 weeks.
    \item This does not prove \(\mu=11\); it only means evidence was not strong enough at the 10\% level.
    \item Some individuals may still be unemployed for fewer than 11 weeks; the test concerns the population mean. A Type II error here would mean the platform really did reduce the true mean below 11 weeks, but the ministry failed to detect that improvement.
\end{itemize}

\item
\section{Export quality audit with two evidence thresholds}
A precision-parts exporter reports monthly defect performance to international buyers. Contracts state that bonus payments are tied to keeping the true defect proportion below 5\%, and this quarter's audit uses that 5\% figure as the reference point. Inspectors drew a sample from outgoing shipments and ran the one-sided test \(H_0\colon p=0.05\) versus \(H_1\colon p<0.05\). The p-value from the sample is 0.004, and the same result is reviewed at two evidence standards: \(\alpha=0.05\) and \(\alpha=0.01\). For each level, state the decision and the practical contract conclusion. Then compare the two levels and explain what still cannot be claimed with absolute certainty even when both decisions support the same direction.

\subsection*{C6}
\begin{itemize}
    \item At 5\%, reject \(H_0\): there is sufficient evidence to conclude the defect proportion is below 5\%.
    \item At 1\%, reject \(H_0\): there is sufficient evidence at this stricter level to conclude the defect proportion is below 5\%.
    \item The conclusion is robust across the two thresholds, but it is still not absolute proof that the true defect rate is below 5\% in every future production cycle.
\end{itemize}

\item
\section{Digital payment adoption check under strict and lenient policy standards}
A central bank tracks household payment behavior before deciding whether to fund additional cashless-service points in rural areas. The current national target in the inclusion strategy is 70\% regular digital-payment use, and this target is the benchmark for statistical review. A nationwide household sample was analyzed with \(H_0\colon p=0.70\), \(H_1\colon p>0.70\), producing a p-value of 0.041. Officials compare two rules: a strict research standard \(\alpha=0.01\) and a planning standard \(\alpha=0.05\). State which conclusion is supported under each rule, explain what policy message is acceptable at each threshold, and distinguish what may not be overclaimed when the stricter rule does not support the claim.

\subsection*{C6}
\begin{itemize}
    \item At 1\%, fail to reject \(H_0\): evidence is not strong enough to claim the population proportion exceeds 70\%.
    \item At 5\%, reject \(H_0\): evidence is sufficient to support that the population proportion exceeds 70\%.
    \item Even at 5\%, this does not guarantee every household uses digital payments weekly; the conclusion is probabilistic, not absolute certainty.
\end{itemize}

\item
\section{Border operations dashboard: two independent service tests}
A cross-border logistics agency is preparing its quarterly staffing memo. Managers require separate decisions for cargo-flow efficiency and household support coverage, so two independent tests are reported from the same quarter.

\textbf{Test A (customs wait time).} The port service contract is written around a 2.5-hour average customs wait, based on the agreed operational standard. The monitoring team sampled shipment records and tested whether the true mean wait is different from that contract value using \(H_0\colon \mu=2.5\), \(H_1\colon \mu\ne2.5\), at \(\alpha=0.05\). Reported decision: \textit{reject \(H_0\)}.

\textbf{Test B (school meal participation in border districts).} A separate social-protection unit tracks whether participation has dropped below the policy floor of 62\%. District records were tested with \(H_0\colon p=0.62\), \(H_1\colon p<0.62\), at \(\alpha=0.05\). Reported decision: \textit{fail to reject \(H_0\)}.

Interpret Test A and Test B separately. For each test, say what the agency can conclude in context, and what claim it must avoid (for example, avoid turning a mean conclusion into a statement about every single shipment or every district).

\subsection*{C6}
\begin{itemize}
    \item \textbf{Test A:} Evidence supports that the population mean customs wait time is not 2.5 hours.
    \item \textbf{Test A limitation:} The result does not identify direction unless sample information is checked separately, and it does not imply all shipments wait longer or shorter than 2.5 hours.
    \item \textbf{Test B:} At 5\%, evidence is insufficient to conclude participation is below 62\%; this does not prove participation is exactly 62\%.
\end{itemize}

\item
\section{Education and labor program review: two different directional checks}
A regional policy board is finalizing next year's education and labor allocations. Because the two programs pursue different goals, analysts run two independent tests with different directions.

\textbf{Test A (school meal participation).} The nutrition policy requires at least 62\% participation, a floor set in the regional child-welfare compact. A district survey is tested with \(H_0\colon p=0.62\), \(H_1\colon p<0.62\), at \(\alpha=0.10\), and the reported decision is \textit{reject \(H_0\)}.

\textbf{Test B (wage growth stability).} A labor-support package was budgeted under a 3.5\% annual wage-growth assumption. Economists test whether the true mean growth is different using \(H_0\colon \mu=3.5\), \(H_1\colon \mu\ne3.5\), at \(\alpha=0.10\), and they report \textit{fail to reject \(H_0\)}.

Give two independent interpretations, one for each test. Also interpret one concrete error risk: if Test A were a Type I error, what false policy story would the board tell about school meal participation, and what unnecessary action might follow from that mistake?

\subsection*{C6}
\begin{itemize}
    \item \textbf{Test A:} The data provide sufficient evidence that participation is below 62\%.
    \item \textbf{Test A error meaning:} A Type I error would mean the board wrongly concludes participation fell below 62\% and reallocates funds, even though the true participation rate is actually at the 62\% benchmark.
    \item \textbf{Test B:} There is not enough evidence to conclude wage growth differs from 3.5\%; this does not prove the mean is exactly 3.5\%.
\end{itemize}

\item
\section{Economic development briefing with mixed thresholds}
A national development office is drafting a strategy memo that combines workforce training results and agricultural credit data. The two tests are independent, but officials must compare how sensitive each conclusion is to the chosen evidence threshold.

\textbf{Test A (training certification success).} Donor agreements set an 80\% pass-rate target for certified trainees. This year's sample produces a p-value of 0.074 for testing \(H_0\colon p=0.80\) against \(H_1\colon p>0.80\). Interpret this same test at \(\alpha=0.10\) and \(\alpha=0.05\).

\textbf{Test B (average farm loan size).} A credit program is designed to keep the average approved loan below 18{,}000 dollars so that support remains focused on smallholders. For \(H_0\colon \mu=18000\) versus \(H_1\colon \mu<18000\), at \(\alpha=0.05\), the reported decision is \textit{reject \(H_0\)}.

Compare outcomes across levels and across tests. Identify which conclusion changes when the standard tightens, which one stays clear, and how that difference should affect the confidence of the office's public briefing.

\subsection*{C6}
\begin{itemize}
    \item \textbf{Test A at 10\%:} Reject \(H_0\), so evidence supports a pass proportion above 80\%.
    \item \textbf{Test A at 5\%:} Fail to reject \(H_0\), so evidence is insufficient at this stricter level; this finding is threshold-sensitive.
    \item \textbf{Test B at 5\%:} There is sufficient evidence that the population mean approved loan is below 18{,}000 dollars; this result does not imply every loan is below that value.
\end{itemize}

\item
\section{Telecom service regulation: three independent compliance checks}
A communications regulator is preparing a quarterly compliance bulletin that will be used for licensing discussions. To avoid mixing conclusions, the regulator runs three independent checks, all at \(\alpha=0.05\), each linked to a different rule or service commitment.

\textbf{Test A (broadband access).} The benchmark from the previous national census is 76\%. For \(H_0\colon p=0.76\) versus \(H_1\colon p\ne0.76\), the decision is \textit{fail to reject \(H_0\)}.

\textbf{Test B (customer response speed).} Operator licenses expect average complaint response to remain at 24 hours or less. For \(H_0\colon \mu=24\) versus \(H_1\colon \mu>24\), the decision is \textit{reject \(H_0\)}.

\textbf{Test C (network outage rate).} The technical code sets 3.0\% as the reference monthly outage rate. For \(H_0\colon p=0.03\) versus \(H_1\colon p<0.03\), the decision is \textit{fail to reject \(H_0\)}.

Interpret each test one by one. For each metric, write what can be concluded for regulators and operators, what cannot be concluded, and where a real improvement might still exist but remain statistically unconfirmed.

\subsection*{C6}
\begin{itemize}
    \item \textbf{Test A:} At the 5\% level, evidence is insufficient to conclude the broadband proportion differs from 76\%; this does not prove the true proportion equals 76\%.
    \item \textbf{Test B:} Evidence supports that average complaint response time is above 24 hours.
    \item \textbf{Test C:} Evidence is insufficient to conclude outages are below 3.0\%; a meaningful improvement may exist but remain undetected.
\end{itemize}

\item
\section{Integrated transport and labor assessment with robustness checks}
A policy task force is preparing a cabinet brief and wants to separate robust findings from findings that depend on the evidence threshold. Three independent indicators are reviewed because they affect different budget lines.

\textbf{Test A (fuel inflation for fare planning).} The benchmark in the transport budget is 4.0\% monthly fuel inflation. The test of whether the true mean is above that value gives a p-value of 0.004.

\textbf{Test B (unemployment duration for labor policy).} The labor baseline is 11 weeks. The test of whether the true mean duration is below that value gives a p-value of 0.027.

\textbf{Test C (broadband access for digital inclusion).} The reference from the census is 76\%. The two-sided test of whether the true proportion differs from 76\% gives a p-value of 0.081.

All three tests are interpreted at \(\alpha=0.10\) and \(\alpha=0.05\). Identify which conclusions are stable at both levels and which are threshold-sensitive. Then state how strongly the task force should communicate each claim in the cabinet brief, especially for the result that changes across thresholds.

\subsection*{C6}
\begin{itemize}
    \item \textbf{Test A:} Reject \(H_0\) at both 10\% and 5\%; the conclusion that mean fuel inflation is above 4.0\% is robust.
    \item \textbf{Test B:} Reject \(H_0\) at both 10\% and 5\%; the conclusion that mean unemployment duration is below 11 weeks is also robust.
    \item \textbf{Test C:} Reject \(H_0\) at 10\% but fail to reject at 5\%; the conclusion that broadband access differs from 76\% is threshold-sensitive and should be communicated more cautiously.
\end{itemize}

\item
\section{National performance synthesis: three tests, multiple confidence standards}
A national monitoring committee is writing an end-of-year performance chapter that combines evidence from digital finance, export quality, and school nutrition programs. The committee must classify which findings are strong under strict standards and which are only suggestive under looser standards, so that ministers do not overstate results.

\textbf{Test A (digital payment adoption).} Against the national 70\% inclusion target, the one-sided test for a higher true adoption proportion gives p-value 0.012.

\textbf{Test B (export defect control).} Against the 5\% contractual defect limit, the one-sided test for a lower true defect proportion gives p-value 0.043.

\textbf{Test C (school meal participation).} Against the 62\% policy floor, the one-sided test for a lower true participation proportion gives p-value 0.097.

Each test is reviewed at three thresholds: \(\alpha=0.10\), \(\alpha=0.05\), and \(\alpha=0.01\). For every test, state exactly at which thresholds the claim is supported and at which it is not. Then separate robust findings from threshold-sensitive findings and write the appropriate communication message: which results can be reported confidently, which should be reported as moderate evidence, and which should be presented as tentative.

\subsection*{C6}
\begin{itemize}
    \item \textbf{Test A (p = 0.012):} Supported at 10\% and 5\%, but not at 1\%; the evidence is strong but not at the strictest threshold.
    \item \textbf{Test B (p = 0.043):} Supported at 10\% and 5\%, but not at 1\%; this is moderate evidence and should not be presented as certainty.
    \item \textbf{Test C (p = 0.097):} Supported only at 10\%, not at 5\% or 1\%; this is the most threshold-sensitive finding and should be reported as tentative.
\end{itemize}

\end{ExamProblems}
\end{document}
