\input{../../../preamble/preamble_exams.tex}

\begin{document}
\ExamTitleBlock{11th grade}{Term 3 --- Lesson 1 Practice Activity (C7) with Solutions}{}

\section*{Contents}
\addcontentsline{toc}{section}{Contents}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{1}
\tableofcontents

\ExamSection{Evaluated criteria}
\begin{ExamCriteria}
    \item C7: Builds and interprets one-sample confidence intervals for a population proportion, compares interval precision across confidence levels and sample sizes, and uses interval evidence to judge benchmark plausibility in context.
\end{ExamCriteria}

\ExamSection{Problems}
\begin{ExamProblems}

\item
\section{General Problem: confidence interval workflow for a population proportion}
A training platform wants to estimate the proportion of enrolled learners who complete the full course.
A random sample of \(n=300\) learners includes \(x=198\) completers.
Construct and interpret a 95\% confidence interval for the population completion proportion.

\begin{enumerate}
    \item Define the parameter of interest and compute \(\hat p\).
    \item Check whether normal-approximation conditions are satisfied.
    \item Compute the standard error and margin of error.
    \item Construct the confidence interval.
    \item Interpret the interval in context, and state one thing it does \emph{not} mean.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true population proportion of enrolled learners who complete the course.
    \item Sample proportion: \(\hat p=\frac{x}{n}=\frac{198}{300}=0.66\).
    \item Conditions: \(n\hat p=300(0.66)=198\ge 10\) and \(n(1-\hat p)=300(0.34)=102\ge 10\), so the normal approximation is appropriate.
    \item Standard error: \(SE=\sqrt{\frac{\hat p(1-\hat p)}{n}}=\sqrt{\frac{0.66(0.34)}{300}}\approx 0.0274\).
    \item For 95\% confidence, \(z^*=1.96\). Margin of error: \(ME=z^*\cdot SE=1.96(0.0274)\approx 0.0536\).
    \item Interval: \(\hat p\pm ME=0.66\pm 0.0536=(0.6064,\,0.7136)\approx (0.606,\,0.714)\).
    \item Interpretation: We are 95\% confident that the true completion proportion is between about 60.6\% and 71.4\%.
    \item Limitation/non-certainty statement: this does not mean exactly 95\% of individual learners complete, and it does not guarantee the true proportion must be inside this single computed interval with absolute certainty.
\end{itemize}

\item
\section{Product quality pass rate: first direct computation}
A factory audits battery packs and records \(x=462\) pass-quality units out of \(n=500\).
Let \(p\) be the true proportion of units that meet the quality standard.
Construct a 95\% confidence interval for \(p\).

\begin{enumerate}
    \item State the parameter and compute \(\hat p\).
    \item Check interval conditions.
    \item Compute the 95\% confidence interval.
    \item Interpret the interval for quality monitoring.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true population pass-quality proportion.
    \item Sample proportion: \(\hat p=\frac{462}{500}=0.924\).
    \item Conditions: \(n\hat p=500(0.924)=462\ge 10\), \(n(1-\hat p)=500(0.076)=38\ge 10\), so normal-approximation interval is valid.
    \item \(SE=\sqrt{\frac{0.924(0.076)}{500}}\approx 0.0119\).
    \item \(ME=1.96(0.0119)\approx 0.0232\).
    \item 95\% CI: \(0.924\pm 0.0232=(0.9008,\,0.9472)\approx (0.901,\,0.947)\).
    \item Interpretation: We are 95\% confident the true pass-quality proportion is between about 90.1\% and 94.7\%.
\end{itemize}

\item
\section{Service completion in a municipal office: interpretation focus}
In a random audit of service files, \(x=344\) out of \(n=400\) applications were completed within the official time window.
Let \(p\) be the true on-time completion proportion.
Construct a 90\% confidence interval and explain what it means.

\begin{enumerate}
    \item Compute \(\hat p\) and check conditions.
    \item Construct the 90\% confidence interval.
    \item Write one correct interpretation and one incorrect interpretation to avoid.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true on-time completion proportion.
    \item \(\hat p=\frac{344}{400}=0.86\).
    \item Conditions: \(n\hat p=400(0.86)=344\ge 10\), \(n(1-\hat p)=400(0.14)=56\ge 10\).
    \item For 90\%, \(z^*=1.645\).
    \item \(SE=\sqrt{\frac{0.86(0.14)}{400}}\approx 0.0173\), so \(ME=1.645(0.0173)\approx 0.0285\).
    \item 90\% CI: \(0.86\pm 0.0285=(0.8315,\,0.8885)\approx (0.832,\,0.889)\).
    \item Correct interpretation: We are 90\% confident the true on-time completion proportion in the full service population is between about 83.2\% and 88.9\%.
    \item Incorrect interpretation to avoid: ``90\% of individual applications are guaranteed to be on time'' (the interval estimates a population proportion; it is not an individual guarantee).
\end{itemize}

\item
\section{App feature adoption and parameter distinction: sample proportion vs population proportion}
A technology company samples \(n=180\) active users and finds \(x=135\) users adopted a new feature.
Let \(p\) be the true adoption proportion among all active users.
Construct a 95\% confidence interval and explain the difference between \(\hat p\) and \(p\).

\begin{enumerate}
    \item Compute \(\hat p\) and identify \(p\).
    \item Check conditions and compute the interval.
    \item Explain why \(\hat p\) is known but \(p\) is unknown.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true feature-adoption proportion in the full active-user population.
    \item Sample proportion: \(\hat p=\frac{135}{180}=0.75\).
    \item Conditions: \(n\hat p=180(0.75)=135\ge 10\), \(n(1-\hat p)=180(0.25)=45\ge 10\).
    \item \(SE=\sqrt{\frac{0.75(0.25)}{180}}\approx 0.0323\), \(ME=1.96(0.0323)\approx 0.0633\).
    \item 95\% CI: \(0.75\pm 0.0633=(0.6867,\,0.8133)\approx (0.687,\,0.813)\).
    \item Distinction: \(\hat p=0.75\) is computed directly from the sample and is known after data collection. \(p\) is the unknown population value we want to estimate. The confidence interval uses \(\hat p\) to estimate plausible values of \(p\).
\end{itemize}

\item
\section{Payment adoption at three confidence levels: 90\%, 95\%, and 99\%}
A retailer studies digital-wallet adoption.
From \(n=300\) customers, \(x=216\) used digital wallet payment.
Using the same sample data, construct 90\%, 95\%, and 99\% confidence intervals for \(p\).

\begin{enumerate}
    \item Compute \(\hat p\) and the standard error.
    \item Construct all three intervals.
    \item Compare margins of error and interval widths.
    \item State the confidence--precision tradeoff.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true population digital-wallet adoption proportion.
    \item \(\hat p=\frac{216}{300}=0.72\).
    \item Conditions: \(300(0.72)=216\ge 10\), \(300(0.28)=84\ge 10\).
    \item \(SE=\sqrt{\frac{0.72(0.28)}{300}}\approx 0.0259\).
    \item 90\%: \(z^*=1.645\), \(ME\approx 1.645(0.0259)=0.0426\), CI \((0.677,\,0.763)\), width \(\approx 0.085\).
    \item 95\%: \(z^*=1.96\), \(ME\approx 0.0508\), CI \((0.669,\,0.771)\), width \(\approx 0.102\).
    \item 99\%: \(z^*=2.576\), \(ME\approx 0.0668\), CI \((0.653,\,0.787)\), width \(\approx 0.134\).
    \item Comparison: as confidence level increases, both margin of error and interval width increase.
    \item Tradeoff statement: higher confidence gives more cautious (wider) estimates; lower confidence gives more precise (narrower) estimates.
\end{itemize}

\item
\section{Contactless payment reliability: explicit width and margin-of-error comparison}
A transit authority samples \(n=340\) trips and finds \(x=255\) paid by contactless card.
Let \(p\) be the true contactless-payment proportion.
Compare a 95\% interval and a 99\% interval for the same sample.

\begin{enumerate}
    \item Compute \(\hat p\), \(SE\), and both margins of error.
    \item Construct both intervals.
    \item Compare widths numerically and explain which interval is more precise.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true contactless-payment proportion.
    \item \(\hat p=\frac{255}{340}=0.75\).
    \item Conditions: \(340(0.75)=255\ge 10\), \(340(0.25)=85\ge 10\).
    \item \(SE=\sqrt{\frac{0.75(0.25)}{340}}\approx 0.0235\).
    \item 95\%: \(ME=1.96(0.0235)\approx 0.0460\), CI \((0.704,\,0.796)\), width \(\approx 0.092\).
    \item 99\%: \(ME=2.576(0.0235)\approx 0.0605\), CI \((0.690,\,0.810)\), width \(\approx 0.121\).
    \item Width comparison: the 95\% interval is narrower by about \(0.121-0.092=0.029\).
    \item Precision interpretation: the 95\% interval is more precise (narrower), while the 99\% interval is more conservative (wider).
\end{itemize}

\item
\section{Screening compliance and confidence-level choice: practical tradeoff}
A health team checks screening compliance: \(x=184\) compliant cases out of \(n=250\).
Let \(p\) be the true compliance proportion.
Construct both 90\% and 99\% intervals, then choose which level to report in two contexts:
(i) routine internal monitoring, (ii) high-stakes public compliance reporting.

\begin{enumerate}
    \item Compute \(\hat p\), both intervals, and their widths.
    \item Recommend a confidence level for each context and justify.
    \item Briefly connect this to earlier significance-threshold thinking while keeping interval focus.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true screening-compliance proportion.
    \item \(\hat p=\frac{184}{250}=0.736\).
    \item Conditions: \(250(0.736)=184\ge 10\), \(250(0.264)=66\ge 10\).
    \item \(SE=\sqrt{\frac{0.736(0.264)}{250}}\approx 0.0279\).
    \item 90\%: \(ME=1.645(0.0279)\approx 0.0459\), CI \((0.690,\,0.782)\), width \(\approx 0.092\).
    \item 99\%: \(ME=2.576(0.0279)\approx 0.0718\), CI \((0.664,\,0.808)\), width \(\approx 0.144\).
    \item Recommendation: for routine internal monitoring, 90\% may be acceptable because narrower intervals support faster operational adjustments.
    \item Recommendation: for high-stakes public reporting, 99\% is often preferred because the institution wants stronger confidence before public claims.
    \item Concept link: as with stricter \(\alpha\) in earlier testing criteria, stricter confidence demands more conservative evidence statements; here this appears as wider confidence intervals rather than p-value decisions.
\end{itemize}

\item
\section{Retention estimation under two sample sizes: same \(\hat p\), different \(n\)}
A subscription service estimates retention. In both cases, \(\hat p=0.80\), but sample sizes differ:
\begin{itemize}
    \item Sample A: \(n=100\), so \(x=80\).
    \item Sample B: \(n=400\), so \(x=320\).
\end{itemize}
Construct 95\% confidence intervals for both and compare precision.

\begin{enumerate}
    \item Compute both intervals.
    \item Compare margins of error and widths.
    \item Explain what this says about stability of estimation.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true retention proportion.
    \item Both samples have \(\hat p=0.80\).
    \item Conditions hold in both cases: \(n\hat p\ge 10\) and \(n(1-\hat p)\ge 10\).
    \item Sample A (\(n=100\)): \(SE=\sqrt{\frac{0.8(0.2)}{100}}=0.04\), \(ME=1.96(0.04)=0.0784\), CI \((0.722,\,0.878)\), width \(0.1568\).
    \item Sample B (\(n=400\)): \(SE=\sqrt{\frac{0.8(0.2)}{400}}=0.02\), \(ME=1.96(0.02)=0.0392\), CI \((0.761,\,0.839)\), width \(0.0784\).
    \item Comparison: quadrupling sample size halves standard error and margin of error, so the interval is much narrower.
    \item Stability interpretation: the larger sample gives a more stable and reliable estimate around the same central sample proportion.
\end{itemize}

\item
\section{Delivery reliability and why larger \(n\) narrows intervals}
A delivery company estimates the proportion of orders arriving on time.
In two months, the observed sample proportion is the same, \(\hat p=0.82\):
\begin{itemize}
    \item Month 1: \(n=150\), \(x=123\).
    \item Month 2: \(n=600\), \(x=492\).
\end{itemize}
Construct 95\% confidence intervals and explain why the larger sample narrows the interval.

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true on-time delivery proportion.
    \item Month 1: \(\hat p=0.82\), \(SE=\sqrt{\frac{0.82(0.18)}{150}}\approx 0.0314\), \(ME\approx 1.96(0.0314)=0.0615\), CI \((0.759,\,0.882)\).
    \item Month 2: \(\hat p=0.82\), \(SE=\sqrt{\frac{0.82(0.18)}{600}}\approx 0.0157\), \(ME\approx 1.96(0.0157)=0.0308\), CI \((0.789,\,0.851)\).
    \item Reason for narrowing: \(SE=\sqrt{\frac{\hat p(1-\hat p)}{n}}\), so increasing \(n\) decreases \(SE\), which decreases \(ME=z^*SE\), producing a narrower interval.
    \item Interpretation: with more data, the company can estimate on-time reliability more precisely.
\end{itemize}

\item
\section{Planning sample size for program participation precision}
A youth program wants a 95\% confidence interval for participation proportion with margin of error at most \(0.03\).
A pilot study suggests \(\hat p\approx 0.68\).
Use
\[
 n\ge \frac{(z^*)^2\hat p(1-\hat p)}{ME^2}
\]
with \(z^*=1.96\).

\begin{enumerate}
    \item Compute the minimum required sample size.
    \item Explain why this planning step is connected to reliability concerns from earlier criteria.
\end{enumerate}

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true program-participation proportion.
    \item Given \(\hat p\approx 0.68\), desired \(ME=0.03\), \(z^*=1.96\).
    \item \(n\ge \frac{(1.96)^2(0.68)(0.32)}{(0.03)^2}=\frac{3.8416\times 0.2176}{0.0009}\approx 929.3\).
    \item Minimum whole-number sample size: \(n=930\).
    \item Interpretation: to achieve the requested precision at 95\% confidence, the team should sample at least 930 participants.
    \item Reliability connection: planning for a target margin of error is an interval-based way to manage uncertainty and improve decision reliability.
\end{itemize}

\item
\section{Program participation benchmark plausibility: benchmark inside interval}
A city wants to know whether its participation benchmark \(p_0=0.70\) is plausible.
From \(n=350\) residents, \(x=238\) participated.
Construct a 95\% confidence interval and evaluate whether 70\% is plausible.

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true city participation proportion.
    \item \(\hat p=\frac{238}{350}=0.68\).
    \item Conditions: \(350(0.68)=238\ge 10\), \(350(0.32)=112\ge 10\).
    \item \(SE=\sqrt{\frac{0.68(0.32)}{350}}\approx 0.0249\), \(ME=1.96(0.0249)\approx 0.0489\).
    \item 95\% CI: \(0.68\pm 0.0489=(0.631,\,0.729)\).
    \item Benchmark check: \(0.70\) lies inside \((0.631,0.729)\), so 70\% is plausible given this sample evidence.
    \item Inference connection: inclusion does not prove \(p=0.70\), but it indicates the benchmark is compatible with the interval evidence.
\end{itemize}

\item
\section{On-time delivery target check: benchmark outside interval}
A logistics firm sets a target of \(p_0=0.95\) for on-time delivery.
In a random sample, \(x=357\) on-time deliveries are observed out of \(n=400\).
Construct a 95\% confidence interval and evaluate the target.

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true on-time delivery proportion.
    \item \(\hat p=\frac{357}{400}=0.8925\).
    \item Conditions: \(400(0.8925)=357\ge 10\), \(400(0.1075)=43\ge 10\).
    \item \(SE=\sqrt{\frac{0.8925(0.1075)}{400}}\approx 0.0155\), \(ME=1.96(0.0155)\approx 0.0303\).
    \item 95\% CI: \(0.8925\pm 0.0303=(0.862,\,0.923)\).
    \item Benchmark check: \(0.95\) is outside (above) this interval, so the 95\% target is not plausible under the current data.
    \item Inference connection: interval exclusion suggests evidence against the benchmark claim at a level conceptually aligned with earlier hypothesis-testing ideas, but the reasoning here is confidence-interval based.
\end{itemize}

\item
\section{Survey approval with two nearby benchmarks: mixed plausibility and tentative evidence}
A survey samples \(n=100\) respondents, with \(x=68\) approving a policy.
Let \(p\) be the population approval proportion.
Construct a 95\% confidence interval and assess two benchmarks: \(p_1=0.62\) and \(p_2=0.80\).
Then discuss how interval width affects certainty.

\subsection*{C7}
\begin{itemize}
    \item Parameter: \(p\) = true policy-approval proportion.
    \item \(\hat p=\frac{68}{100}=0.68\).
    \item Conditions: \(100(0.68)=68\ge 10\), \(100(0.32)=32\ge 10\).
    \item \(SE=\sqrt{\frac{0.68(0.32)}{100}}\approx 0.0466\), \(ME=1.96(0.0466)\approx 0.0914\).
    \item 95\% CI: \(0.68\pm 0.0914=(0.589,\,0.771)\).
    \item Benchmark checks: \(0.62\) lies inside the interval (plausible), while \(0.80\) lies outside (not plausible).
    \item Tentative-evidence note: the interval is fairly wide, so although \(0.80\) is excluded, precision is limited; collecting more data could sharpen conclusions.
\end{itemize}

\item
\section{Policy dashboard I: quality, adoption, and completion (three-case synthesis)}
A ministry reviews three independent proportion indicators and corresponding targets.
Construct 95\% intervals and classify each conclusion as \textbf{strong}, \textbf{moderate}, or \textbf{tentative} based on benchmark inclusion and interval width.

\begin{itemize}
    \item Product quality pass rate: \(x=470\), \(n=500\), target \(p_0=0.90\).
    \item App feature adoption: \(x=210\), \(n=300\), target \(p_0=0.72\).
    \item Training completion: \(x=96\), \(n=150\), target \(p_0=0.70\).
\end{itemize}

\subsection*{C7}
\begin{itemize}
    \item \textbf{Case 1 (quality):} \(\hat p=0.94\), \(SE\approx 0.0106\), \(ME\approx 0.0208\), CI \((0.919,\,0.961)\). Target 0.90 is outside (below interval). Conclusion: target is exceeded; evidence is \textbf{strong} because interval is narrow.
    \item \textbf{Case 2 (adoption):} \(\hat p=0.70\), \(SE\approx 0.0265\), \(ME\approx 0.0519\), CI \((0.648,\,0.752)\). Target 0.72 is inside. Conclusion: target remains plausible; evidence is \textbf{moderate}.
    \item \textbf{Case 3 (completion):} \(\hat p=0.64\), \(SE\approx 0.0392\), \(ME\approx 0.0768\), CI \((0.563,\,0.717)\). Target 0.70 is inside but near upper side. Conclusion: target may still be plausible, but evidence is \textbf{tentative} due to wider interval.
    \item Synthesis: narrow intervals support more robust statements; wide intervals require more cautious policy language.
\end{itemize}

\item
\section{Regional digital-payment dashboard: reliability and practical consequences}
Three regions report digital-payment adoption.
For each region, construct a 95\% confidence interval, compare with target \(p_0=0.78\), and prioritize intervention.

\begin{itemize}
    \item Region North: \(x=384\), \(n=480\).
    \item Region South: \(x=132\), \(n=200\).
    \item Region East: \(x=70\), \(n=100\).
\end{itemize}

\subsection*{C7}
\begin{itemize}
    \item \textbf{North:} \(\hat p=0.80\), \(SE\approx 0.0183\), \(ME\approx 0.0358\), CI \((0.764,\,0.836)\). Target 0.78 is inside: plausible and relatively stable estimate (narrow interval).
    \item \textbf{South:} \(\hat p=0.66\), \(SE\approx 0.0335\), \(ME\approx 0.0657\), CI \((0.594,\,0.726)\). Target 0.78 is outside: target likely not met; evidence is fairly strong.
    \item \textbf{East:} \(\hat p=0.70\), \(SE\approx 0.0458\), \(ME\approx 0.0897\), CI \((0.610,\,0.790)\). Target 0.78 is inside but near upper edge: conclusion is tentative due to wide interval.
    \item Practical prioritization: immediate intervention in South, monitoring in East with more data collection, and maintenance strategy in North.
\end{itemize}

\item
\section{National performance report synthesis: strongest integrated interpretation}
A national dashboard tracks four proportion indicators using 95\% confidence intervals.
For each case, compute the interval, compare with its benchmark, classify evidence strength, and write a concise report conclusion.

\begin{itemize}
    \item Delivery reliability: \(x=455\), \(n=520\), benchmark \(p_0=0.90\).
    \item Digital payment adoption: \(x=168\), \(n=240\), benchmark \(p_0=0.75\).
    \item Screening compliance: \(x=102\), \(n=120\), benchmark \(p_0=0.80\).
    \item Course completion: \(x=54\), \(n=90\), benchmark \(p_0=0.75\).
\end{itemize}

\subsection*{C7}
\begin{itemize}
    \item \textbf{Delivery reliability:} \(\hat p=0.875\), \(SE\approx 0.0145\), \(ME\approx 0.0284\), CI \((0.847,\,0.903)\). Benchmark 0.90 lies inside but near the upper side: target is plausible, with \textbf{moderately strong} evidence due to relatively narrow interval.
    \item \textbf{Digital payment adoption:} \(\hat p=0.70\), \(SE\approx 0.0296\), \(ME\approx 0.0580\), CI \((0.642,\,0.758)\). Benchmark 0.75 is inside near upper edge: tentative-to-moderate support; more data could clarify.
    \item \textbf{Screening compliance:} \(\hat p=0.85\), \(SE\approx 0.0326\), \(ME\approx 0.0639\), CI \((0.786,\,0.914)\). Benchmark 0.80 is inside: plausible and reasonably supported.
    \item \textbf{Course completion:} \(\hat p=0.60\), \(SE\approx 0.0516\), \(ME\approx 0.101\), CI \((0.499,\,0.701)\). Benchmark 0.75 is outside (above): target likely not met, but interval is wide so precision is limited.
    \item \textbf{Integrated report statement:} interval evidence suggests strongest concern for course completion shortfall, possible concern for payment adoption near target, and generally acceptable but still monitorable performance for delivery reliability and screening compliance. Narrow intervals support firmer decisions; wide intervals warrant cautious interpretation and additional sampling.
\end{itemize}

\end{ExamProblems}
\end{document}
