\input{../../preamble/preamble_exams.tex}

\begin{document}
	\ExamTitleBlock{11th grade}{Term 3 — Lesson 1 Practice Activity (C1–C3) with Solutions}{}

	\ExamSection{Evaluated criteria}
	\begin{ExamCriteria}
		\item C1: Describe the concept of a statistical hypothesis in context.
		\item C2: Identify the null and alternative hypotheses using correct parameter notation.
		\item C3: Determine whether a test is one tailed or two tailed based on the claim.
	\end{ExamCriteria}

	\ExamSection{Problems}

	\begin{ExamProblems}

		\item
		\subsection*{Problem description}
		In the homework completion time app scenario, a school introduces a planning app and claims that the average time students need to finish nightly homework is now above 90 minutes. A random sample of Grade 11 students records daily homework time, and the parameter of interest is the population mean homework time \(\mu\).

		\subsection*{C1}
		A statistical hypothesis in this context is a testable statement about the population mean \(\mu\) based on sample homework times from students.

		\subsection*{C2}
		Let \(\mu\) denote the population mean homework time. \(H_0\colon \mu = 90\) minutes and \(H_a\colon \mu > 90\) minutes.

		\subsection*{C3}
		This is a right-tailed test because the claim supports values of \(\mu\) greater than 90 minutes.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the vaccination campaign coverage scenario, a public clinic reports that at least 80\% of teenagers in the district received a seasonal vaccine. A health survey checks a random sample of teenagers and records vaccination status, and the parameter is the population proportion \(p\) of vaccinated teenagers.

		\subsection*{C1}
		A statistical hypothesis here is a statement about the population proportion \(p\), and sample vaccination outcomes provide evidence about that population value.

		\subsection*{C2}
		Let \(p\) denote the population proportion of vaccinated teenagers. \(H_0\colon p = 0.80\) and \(H_a\colon p < 0.80\).

		\subsection*{C3}
		This is a left-tailed test because the concern is whether coverage is below 80\%.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the bottle filling machine calibration scenario, a factory sets a juice bottling machine to fill 500 mL on average. A quality engineer takes a random sample of bottles to check whether the true mean fill amount has shifted from target, and the parameter is the population mean fill volume \(\mu\).

		\subsection*{C1}
		A statistical hypothesis in this setting is a claim about the true mean bottle volume \(\mu\) for all produced bottles, assessed with sample data.

		\subsection*{C2}
		Let \(\mu\) denote the population mean fill volume. \(H_0\colon \mu = 500\) mL and \(H_a\colon \mu \neq 500\) mL.

		\subsection*{C3}
		This is a two-tailed test because both increases and decreases from 500 mL are relevant.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the mobile banking adoption scenario, a bank launches a digital campaign and claims that more than 60\% of its young clients now use the mobile app weekly. A random sample of clients is classified as weekly user or non-user, and the parameter is the population proportion \(p\) of weekly users.

		\subsection*{C1}
		A statistical hypothesis here is a claim about the population proportion \(p\) of clients who use the app weekly, evaluated with sample usage data.

		\subsection*{C2}
		Let \(p\) denote the population proportion of weekly app users. \(H_0\colon p = 0.60\) and \(H_a\colon p > 0.60\).

		\subsection*{C3}
		This is a right-tailed test because the claim supports values of \(p\) above 0.60.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the runner recovery heart rate scenario, a sports scientist states that after a new training plan, the average one-minute recovery heart rate of runners is below 110 beats per minute. A random sample of team athletes is measured after standardized workouts, and the parameter is the population mean recovery heart rate \(\mu\).

		\subsection*{C1}
		A statistical hypothesis in this context is a statement about the population mean recovery heart rate \(\mu\), tested using sample athlete measurements.

		\subsection*{C2}
		Let \(\mu\) denote the population mean one-minute recovery heart rate. \(H_0\colon \mu = 110\) bpm and \(H_a\colon \mu < 110\) bpm.

		\subsection*{C3}
		This is a left-tailed test because the claim supports values of \(\mu\) below 110 bpm.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the river water compliance rate scenario, an environmental agency says that 90\% of river samples meet a safe nitrate standard. Inspectors collect a random sample of sites and record whether each sample is compliant, and the parameter is the population proportion \(p\) of compliant sites.

		\subsection*{C1}
		A statistical hypothesis here is a claim about the true compliance proportion \(p\) across all sites, examined through sample compliance outcomes.

		\subsection*{C2}
		Let \(p\) denote the population proportion of compliant river sites. \(H_0\colon p = 0.90\) and \(H_a\colon p \neq 0.90\).

		\subsection*{C3}
		This is a two-tailed test because any change from 90\% is relevant.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the website page load performance scenario, a tech company updates its servers and claims that the mean page load time is now greater than 2.5 seconds during peak hours due to added security checks. Engineers monitor a random sample of load times, and the parameter is the population mean load time \(\mu\).

		\subsection*{C1}
		A statistical hypothesis in this setting is a statement about the population mean load time \(\mu\), with sample load times used as evidence.

		\subsection*{C2}
		Let \(\mu\) denote the population mean page load time during peak hours. \(H_0\colon \mu = 2.5\) seconds and \(H_a\colon \mu > 2.5\) seconds.

		\subsection*{C3}
		This is a right-tailed test because the claim supports values of \(\mu\) greater than 2.5 seconds.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the city bus waiting time standard scenario, a transport authority reports that average waiting time at a major station is 8 minutes. A commuter group collects random waiting-time observations to verify whether the actual mean differs from this standard, and the parameter is the population mean waiting time \(\mu\).

		\subsection*{C1}
		A statistical hypothesis here is a claim about the population mean waiting time \(\mu\), assessed with sample waiting-time observations.

		\subsection*{C2}
		Let \(\mu\) denote the population mean waiting time at the station. \(H_0\colon \mu = 8\) minutes and \(H_a\colon \mu \neq 8\) minutes.

		\subsection*{C3}
		This is a two-tailed test because both values above and below 8 minutes support a difference.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the fertilizer effect on crop height scenario, an agriculture team claims that a low-cost fertilizer keeps average plant height below 42 cm after six weeks to help reduce wind damage. A random sample of treated plants is measured, and the parameter is the population mean plant height \(\mu\).

		\subsection*{C1}
		A statistical hypothesis in this context is a statement about the population mean plant height \(\mu\), using sample plant heights as evidence.

		\subsection*{C2}
		Let \(\mu\) denote the population mean plant height after six weeks with this fertilizer. \(H_0\colon \mu = 42\) cm and \(H_a\colon \mu < 42\) cm.

		\subsection*{C3}
		This is a left-tailed test because the claim supports values of \(\mu\) below 42 cm.
		% --------------------------------------------------

		\item
		\subsection*{Problem description}
		In the checkout time by payment method scenario, a supermarket compares average checkout times for customers using mobile pay and cash and claims there is a difference between the two population means. Random samples are taken from both groups, with parameters \(\mu_1\) for mobile pay and \(\mu_2\) for cash, and the parameter of interest is \(\mu_1-\mu_2\).

		\subsection*{C1}
		A statistical hypothesis here is a statement about the population difference in mean checkout times between mobile pay and cash, evaluated with sample data from both groups.

		\subsection*{C2}
		Let \(\mu_1-\mu_2\) denote the population difference in mean checkout times between mobile pay and cash. \(H_0\colon \mu_1-\mu_2 = 0\) and \(H_a\colon \mu_1-\mu_2 \neq 0\).

		\subsection*{C3}
		This is a two-tailed test because either a positive or negative difference supports the claim of a difference.
		% --------------------------------------------------

	\end{ExamProblems}
\end{document}
