\makeatletter
\def\input@path{{./}{../}{../../}{../../../}{../../../../}{preamble/}{../preamble/}{../../preamble/}{../../../preamble/}{../../../../preamble/}}
\makeatother

\input{preamble.tex}
\input{graphs/uniform_linear_probability_distributions.tex}
\input{graphs/probability_distribution_graphs.tex}
\input{graphs/normal_distribution_graphs.tex}

\title{Grade 10 Term 1 Practice Activities Summary\\Decision Components and Payoff Tables}
\author{}
\date{}

\begin{document}
	\maketitle
	
	\setcounter{tocdepth}{1}
	\setcounter{secnumdepth}{2}
	\phantomsection
	\label{toc}
	\tableofcontents
	
	\section{G10\_T1\_L1\_C1-C2\_practice\_activity1}
	This activity develops C1--C2 skills by identifying decision components and organizing consequences in payoff tables.
	
	\begin{enumerate}
		\item \textbf{Direct alternative-state payoffs.}
		\begin{enumerate}
			\item \textbf{Local Café Investment:} 2 alternatives, 2 states.
			\item \textbf{Green Energy Production Mix:} 3 alternatives, 3 states.
			\item \textbf{Global Shipping Network Design:} 2 alternatives, 4 states.
		\end{enumerate}
		
		\item \textbf{Alternative-based variable payoffs.}
		\begin{enumerate}
			\item \textbf{Community Bakery Lunch Boxes:} 2 alternatives, 2 states.
			\item \textbf{Rural Internet Subscriptions:} 3 alternatives, 2 states.
		\end{enumerate}
		
		\item \textbf{State-based variable payoffs.}
		\begin{enumerate}
			\item \textbf{Coffee Export Routes:} 2 alternatives, 2 states.
			\item \textbf{Lake Ferry Tourism:} 2 alternatives, 3 states.
		\end{enumerate}
		
		\item \textbf{State-based fixed payoffs.}
		\begin{enumerate}
			\item \textbf{Neighborhood Shelter Support:} 2 alternatives, 2 states.
			\item \textbf{Community Hall Emergency Rentals:} 2 alternatives, 3 states.
		\end{enumerate}
		
		\item \textbf{Alternative-based fixed payoffs.}
		\begin{enumerate}
			\item \textbf{School Meal Service Contracts:} 2 alternatives, 2 states.
			\item \textbf{Mobile Clinic Service Contracts:} 3 alternatives, 2 states.
		\end{enumerate}
		
		\item \textbf{State-based fixed payoff plus alternative-based fixed payoff.}
		\begin{enumerate}
			\item \textbf{Warehouse Security Agreements:} 2 alternatives, 2 states.
			\item \textbf{Harbor Inspection Agreements:} 2 alternatives, 3 states.
		\end{enumerate}
		
		\item \textbf{Alternative-based variable payoff plus alternative-based fixed payoff.}
		\begin{enumerate}
			\item \textbf{Local Printing Contracts:} 2 alternatives, 2 states.
			\item \textbf{Cold-Storage Contracts:} 3 alternatives, 2 states.
		\end{enumerate}
		
		\item \textbf{Alternative-based variable payoff plus state-based fixed payoff.}
		\begin{enumerate}
			\item \textbf{Urban Landscaping Bids:} 2 alternatives, 2 states.
			\item \textbf{Solar Installation Bids:} 3 alternatives, 2 states.
		\end{enumerate}
		
		\item \textbf{State-based variable payoff plus alternative-based fixed payoff.}
		\begin{enumerate}
			\item \textbf{Regional Bus Charters:} 2 alternatives, 2 states.
			\item \textbf{Air Cargo Charters:} 2 alternatives, 3 states.
		\end{enumerate}
		
		\item \textbf{State-based variable payoff plus state-based fixed payoff.}
		\begin{enumerate}
			\item \textbf{Fruit Export Permits:} 2 alternatives, 2 states.
			\item \textbf{Seafood Export Licenses:} 2 alternatives, 3 states.
		\end{enumerate}
	\end{enumerate}
	
	\section{G10\_T1\_L1\_C1-C2\_practice\_activity2}
	This activity extends C1--C2 skills by interpreting decision components and building payoff tables from revenue and cost information.

	\begin{enumerate}
		\item \textbf{Food Truck Menu Upgrade:} Identify 2 alternatives, 2 states, and calculate payoff entries as revenue minus cost by state.
		\item \textbf{School Fundraiser Ticket Plan:} Identify 2 alternatives, 3 states, and calculate payoff entries as revenue minus cost by attendance state.
		\item \textbf{City Bike-Share Pricing Strategy:} Identify 3 alternatives, 2 states, and calculate payoff entries as revenue minus cost by ridership state.
		\item \textbf{Subscription Bundle Choices:} Identify 3 alternatives, 2 states, and calculate payoff entries using subscription price times sign-ups minus state-based costs.
		\item \textbf{Regional Distribution Plan:} Identify 2 alternatives, 3 states, and calculate payoff entries using state-based revenue minus delivery-based variable costs.
		\item \textbf{Farmers Market Juice Stand:} Identify 2 alternatives, 2 states, and calculate payoff entries using state-based revenue minus production cost per bottle.
		\item \textbf{After-School Tutoring Sessions:} Identify 2 alternatives, 2 states, and calculate payoff entries using sessions delivered times net fee per session.
		\item \textbf{Art Fair Booth Options:} Identify 2 alternatives, 2 states, and calculate payoff entries using kits sold times price minus alternative-based fixed setup cost.
		\item \textbf{Emergency Supply Contract:} Identify 2 alternatives, 3 states, and calculate payoff entries using state-based retainer plus kit margin minus state-based compliance fee.
	\end{enumerate}

\end{document}
