\input{../../../preamble/preamble_exams.tex}

\begin{document}
\ExamTitleBlock{10th grade}{Term 3 Practice Activity: C4 Variance and Standard Deviation}{}
\ExamSection{C4 Interprets the concept of variance and standard deviation of a probability function.}

\section*{Contents}
\tableofcontents

\section{Problem 1: Streaming Time (Uniform)}

\textbf{Problem.}
A student's daily streaming time \(X\) (minutes) is modeled by
\[
X\sim \text{Uniform}(12,24).
\]

\textbf{Questions.}
\begin{enumerate}
  \item Find \(E[X]\), \(\mathrm{Var}(X)\), and the standard deviation.
  \item What is the support interval and what does equal density mean in this context?
  \item Is a value of 22 minutes more likely than 14 minutes? Explain.
  \item Based on your variance, describe whether times are tightly clustered or widely spread.
\end{enumerate}

\textbf{Solution.}
For \(X\sim\text{Uniform}(a,b)\):
\[
E[X]=\frac{a+b}{2},\qquad \mathrm{Var}(X)=\frac{(b-a)^2}{12}.
\]
With \(a=12\), \(b=24\):
\[
E[X]=\frac{12+24}{2}=18,
\]
\[
\mathrm{Var}(X)=\frac{(24-12)^2}{12}=\frac{144}{12}=12,
\]
\[
\sigma=\sqrt{12}=2\sqrt{3}\approx 3.46.
\]
Support is \([12,24]\), and all values in this interval have the same density, so 22 and 14 are equally likely on equal-length subintervals.

% --------------------------------------------------
\section{Problem 2: Bus Wait Time (Triangular)}

\textbf{Problem.}
A bus wait time \(W\) (minutes) is modeled by a triangular distribution with minimum \(a=2\), mode \(c=6\), and maximum \(b=14\):
\[
W\sim \text{Triangular}(2,6,14).
\]

\textbf{Questions.}
\begin{enumerate}
  \item Find \(E[W]\), \(\mathrm{Var}(W)\), and the standard deviation.
  \item Describe the shape (left-skewed, right-skewed, or symmetric) and justify from parameters.
  \item Explain why values near 6 are more likely than values near 2 or 14.
  \item Interpret the spread in the bus context.
\end{enumerate}

\textbf{Solution.}
For \(\text{Triangular}(a,c,b)\):
\[
E[W]=\frac{a+b+c}{3},\qquad
\mathrm{Var}(W)=\frac{a^2+b^2+c^2-ab-ac-bc}{18}.
\]
So
\[
E[W]=\frac{2+14+6}{3}=\frac{22}{3}\approx 7.33,
\]
\[
\mathrm{Var}(W)=\frac{2^2+14^2+6^2-(2)(14)-(2)(6)-(14)(6)}{18}
=\frac{4+196+36-28-12-84}{18}=\frac{112}{18}=\frac{56}{9}.
\]
\[
\sigma=\sqrt{\frac{56}{9}}=\frac{\sqrt{56}}{3}\approx 2.49.
\]
Since mode 6 is closer to 2 than to 14, the distribution is right-skewed (longer right side).

% --------------------------------------------------
\section{Problem 3: Quiz Scores (Normal)}

\textbf{Problem.}
Quiz scores \(S\) are modeled by
\[
S\sim N(70,8^2).
\]

\textbf{Questions.}
\begin{enumerate}
  \item State \(E[S]\), \(\mathrm{Var}(S)\), and standard deviation.
  \item Give the intervals \([\mu-\sigma,\mu+\sigma]\) and \([\mu-2\sigma,\mu+2\sigma]\).
  \item Explain what symmetry means for this model.
  \item Which is more likely: scoring near 70 or near 54? Explain.
\end{enumerate}

\textbf{Solution.}
For \(N(\mu,\sigma^2)\):
\[
E[S]=\mu=70,\qquad \mathrm{Var}(S)=\sigma^2=64,\qquad \sigma=8.
\]
Intervals:
\[
[\mu-\sigma,\mu+\sigma]=[62,78],\qquad [\mu-2\sigma,\mu+2\sigma]=[54,86].
\]
The normal curve is symmetric about 70, so values equally distant from 70 have equal density; values near 70 are most likely.

% --------------------------------------------------
\section{Problem 4: Comparing Uniform Distributions}

\textbf{Problem.}
Compare these uniform models for download times (seconds):
\[
X_1\sim \text{Uniform}(4,10),\qquad X_2\sim \text{Uniform}(6,12),\qquad X_3\sim \text{Uniform}(4,14).
\]

\textbf{Questions.}
\begin{enumerate}
  \item Find \(E[X_i]\) and \(\mathrm{Var}(X_i)\) for \(i=1,2,3\).
  \item Compare \(X_1\) and \(X_2\): only location changes. How do center and spread change?
  \item Compare \(X_1\) and \(X_3\): only width changes from 6 to 10. How does variance change?
  \item Which model has the widest spread, and why?
\end{enumerate}

\textbf{Solution.}
Using \(E=\frac{a+b}{2}\), \(\mathrm{Var}=\frac{(b-a)^2}{12}\):
\[
E[X_1]=7,\quad \mathrm{Var}(X_1)=\frac{6^2}{12}=3;
\]
\[
E[X_2]=9,\quad \mathrm{Var}(X_2)=\frac{6^2}{12}=3;
\]
\[
E[X_3]=9,\quad \mathrm{Var}(X_3)=\frac{10^2}{12}=\frac{25}{3}.
\]
Shifting both endpoints by the same amount changes the mean but not spread. Increasing interval width increases variance.

% --------------------------------------------------
\section{Problem 5: Comparing Triangular Distributions}

\textbf{Problem.}
A packing process time \(T\) (minutes) follows triangular models:
\[
T_1\sim \text{Triangular}(2,5,10),\quad
T_2\sim \text{Triangular}(2,7,10),\quad
T_3\sim \text{Triangular}(4,7,12).
\]

\textbf{Questions.}
\begin{enumerate}
  \item Compute \(E[T_i]\) and \(\mathrm{Var}(T_i)\) for each model.
  \item Compare \(T_1\) and \(T_2\): only mode changes. How do center and shape change?
  \item Compare \(T_2\) and \(T_3\): all three parameters shift by +2. What stays the same and what changes?
  \item Which model is most concentrated near larger times? Explain using the peak location.
\end{enumerate}

\textbf{Solution.}
Using triangular formulas:
\[
E[T_1]=\frac{2+5+10}{3}=\frac{17}{3},\quad
\mathrm{Var}(T_1)=\frac{2^2+5^2+10^2-2\cdot5-2\cdot10-5\cdot10}{18}=\frac{49}{18}.
\]
\[
E[T_2]=\frac{2+7+10}{3}=\frac{19}{3},\quad
\mathrm{Var}(T_2)=\frac{2^2+7^2+10^2-2\cdot7-2\cdot10-7\cdot10}{18}=\frac{49}{18}.
\]
\[
E[T_3]=\frac{4+7+12}{3}=\frac{23}{3},\quad
\mathrm{Var}(T_3)=\frac{4^2+7^2+12^2-4\cdot7-4\cdot12-7\cdot12}{18}=\frac{49}{18}.
\]
Changing only the mode shifts peak and mean but here keeps variance unchanged; shifting all parameters by +2 shifts location but not spread.

% --------------------------------------------------
\section{Problem 6: Comparing Normal Distributions}

\textbf{Problem.}
Daily step counts (thousands) are modeled as:
\[
N_1\sim N(9,1.5^2),\qquad N_2\sim N(11,1.5^2),\qquad N_3\sim N(9,2.5^2).
\]

\textbf{Questions.}
\begin{enumerate}
  \item State mean and variance for each distribution.
  \item Compare \(N_1\) and \(N_2\): only mean changes. What changes on the graph?
  \item Compare \(N_1\) and \(N_3\): only standard deviation changes. What happens to spread and peak height?
  \item Which distribution gives a higher chance of being far from its mean? Explain.
\end{enumerate}

\textbf{Solution.}
\[
E[N_1]=9,\ \mathrm{Var}(N_1)=2.25;
\qquad
E[N_2]=11,\ \mathrm{Var}(N_2)=2.25;
\qquad
E[N_3]=9,\ \mathrm{Var}(N_3)=6.25.
\]
Changing \(\mu\) moves the center left or right without changing shape width. Increasing \(\sigma\) makes the curve wider and less peaked, so far-from-mean values become more likely.

% --------------------------------------------------
\section{Problem 7: Uniform with Parameters}

\textbf{Problem.}
A random time \(X\) (minutes) is
\[
X\sim \text{Uniform}(p,p+12),\qquad p\in\mathbb{R}.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Express \(E[X]\) in terms of \(p\).
  \item Express \(\mathrm{Var}(X)\) in terms of \(p\).
  \item Which parameter controls location and which controls spread?
  \item If \(p\) increases by 5, how do mean and variance change?
\end{enumerate}

\textbf{Solution.}
\[
E[X]=\frac{p+(p+12)}{2}=p+6,
\qquad
\mathrm{Var}(X)=\frac{((p+12)-p)^2}{12}=\frac{12^2}{12}=12.
\]
So \(p\) controls location; spread is fixed by width 12 and does not depend on \(p\).

% --------------------------------------------------
\section{Problem 8: Triangular with Parameters}

\textbf{Problem.}
A project completion time \(Y\) (days) is modeled by
\[
Y\sim \text{Triangular}(4,4+m,14),\qquad 0\le m\le 10.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Write \(E[Y]\) and \(\mathrm{Var}(Y)\) as functions of \(m\).
  \item State the condition on \(m\) that makes the model valid.
  \item How does the mean change as \(m\) increases?
  \item How does moving the mode affect the shape (left/right skewness)?
\end{enumerate}

\textbf{Solution.}
Here \(a=4\), \(c=4+m\), \(b=14\).
\[
E[Y]=\frac{a+b+c}{3}=\frac{4+14+(4+m)}{3}=\frac{22+m}{3}.
\]
\[
\mathrm{Var}(Y)=\frac{a^2+b^2+c^2-ab-ac-bc}{18}
=\frac{16+196+(4+m)^2-56-4(4+m)-14(4+m)}{18}
=\frac{m^2-10m+100}{18}.
\]
Validity needs \(a\le c\le b\Rightarrow 4\le 4+m\le 14\Rightarrow 0\le m\le 10\).
Mean increases linearly with \(m\). As mode moves right, the distribution becomes less right-skewed and can become left-skewed when the mode passes midpoint 9.

% --------------------------------------------------
\section{Problem 9: Normal with Parameters}

\textbf{Problem.}
A test score variable is
\[
Z\sim N(60+k,\,(2k)^2),\qquad k>0.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Express mean and variance in terms of \(k\).
  \item Determine valid values of \(k\).
  \item As \(k\) increases, how do center and spread change?
  \item Identify which part of the model controls location and which controls spread.
\end{enumerate}

\textbf{Solution.}
\[
E[Z]=60+k,
\qquad
\mathrm{Var}(Z)=(2k)^2=4k^2,
\qquad
\sigma=2k.
\]
For a valid non-degenerate normal spread, \(k>0\). Increasing \(k\) shifts mean right and also increases spread quadratically through variance.

% --------------------------------------------------
\section{Problem 10: Uniform with Parameters}

\textbf{Problem.}
Commute time \(C\) (minutes) is modeled by
\[
C\sim \text{Uniform}(10-r,10+r),\qquad r>0.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Find \(E[C]\) and \(\mathrm{Var}(C)\) in terms of \(r\).
  \item What condition on \(r\) is needed for a valid interval?
  \item As \(r\) increases, what happens to interval width and spread?
  \item For what \(r\) is the distribution narrower, and why?
\end{enumerate}

\textbf{Solution.}
\[
E[C]=\frac{(10-r)+(10+r)}{2}=10,
\]
\[
\mathrm{Var}(C)=\frac{((10+r)-(10-r))^2}{12}=\frac{(2r)^2}{12}=\frac{r^2}{3}.
\]
Need \(r>0\). Width is \(2r\), so larger \(r\) means wider interval and larger variance; smaller positive \(r\) means narrower distribution.

% --------------------------------------------------
\section{Problem 11: Triangular with Parameters}

\textbf{Problem.}
A completion score \(Q\) is modeled by
\[
Q\sim \text{Triangular}(a,a+3,a+9),\qquad a\in\mathbb{R}.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Express \(E[Q]\) and \(\mathrm{Var}(Q)\) in terms of \(a\).
  \item Is this model always valid for any real \(a\)? Explain.
  \item Which parameter controls location? Which controls spread?
  \item If \(a\) increases, what happens to center, spread, and shape?
\end{enumerate}

\textbf{Solution.}
\[
E[Q]=\frac{a+(a+3)+(a+9)}{3}=a+4.
\]
\[
\mathrm{Var}(Q)=\frac{a^2+(a+3)^2+(a+9)^2-a(a+3)-a(a+9)-(a+3)(a+9)}{18}=\frac{7}{2}.
\]
Since \(a\le a+3\le a+9\) always holds, the distribution is valid for all real \(a\). Parameter \(a\) shifts location; spread and shape remain unchanged.

% --------------------------------------------------
\section{Problem 12: Normal with Parameters}

\textbf{Problem.}
Reaction time \(R\) (seconds) is modeled by
\[
R\sim N(m,\,(1+s)^2),\qquad s> -1.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Express \(E[R]\) and \(\mathrm{Var}(R)\) in terms of \(m\) and \(s\).
  \item Give the condition(s) needed for a valid nonzero spread.
  \item Explain how increasing \(m\) changes the distribution.
  \item Explain how increasing \(s\) changes the distribution width and likelihood of values far from the mean.
\end{enumerate}

\textbf{Solution.}
\[
E[R]=m,
\qquad
\mathrm{Var}(R)=(1+s)^2,
\qquad
\sigma=1+s.
\]
A valid standard deviation needs \(\sigma>0\Rightarrow 1+s>0\Rightarrow s>-1\).
Increasing \(m\) shifts the center only. Increasing \(s\) increases spread, making the bell wider and less concentrated around the mean.

% --------------------------------------------------
\section{Problem 13: Uniform with Parameters}

\textbf{Problem.}
A delivery time \(U\) (minutes) for a local route is modeled by
\[
U\sim \text{Uniform}(h-2w,\,h+w),\qquad w>0.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Derive \(E[U]\), \(\mathrm{Var}(U)\), and \(\sigma_U\) in terms of \(h\) and \(w\).
  \item Identify which parameter controls the center and which controls the interval width.
  \item Compare Case A: \((h,w)=(30,4)\) and Case B: \((h,w)=(35,4)\). Does the distribution shift, widen, or both?
  \item Compare Case C: \((h,w)=(30,4)\) and Case D: \((h,w)=(30,7)\). How do center, width, and spread change?
  \item If both \(h\) and \(w\) increase by 2 at the same time, explain whether the model shifts, widens, or both.
\end{enumerate}

\textbf{Solution.}
For \(U\sim\text{Uniform}(a,b)\), \(E[U]=\frac{a+b}{2}\), \(\mathrm{Var}(U)=\frac{(b-a)^2}{12}\).
Here \(a=h-2w\), \(b=h+w\), so
\[
E[U]=\frac{(h-2w)+(h+w)}{2}=h-\frac{w}{2},
\]
\[
\mathrm{Var}(U)=\frac{\big((h+w)-(h-2w)\big)^2}{12}=\frac{(3w)^2}{12}=\frac{3w^2}{4},
\]
\[
\sigma_U=\sqrt{\frac{3w^2}{4}}=\frac{\sqrt{3}}{2}\,w \quad (w>0).
\]
The center \(h-\frac{w}{2}\) depends on both \(h\) and \(w\), while the width \(b-a=3w\) depends only on \(w\).
Case A \(\to\) B keeps \(w\) fixed, so width and spread stay the same, and the model shifts right by 5.
Case C \(\to\) D keeps \(h\) fixed and increases \(w\), so width grows from \(12\) to \(21\), spread increases, and the center moves from \(28\) to \(26.5\) (left).
If \(h\) and \(w\) both increase by 2, then width increases by \(3(2)=6\) (widens) and
\[
E[U]'-(E[U])=\left(h+2-\frac{w+2}{2}\right)-\left(h-\frac{w}{2}\right)=1,
\]
so it also shifts right by 1.

% --------------------------------------------------
\section{Problem 14: Triangular with Parameters}

\textbf{Problem.}
A machine-cycle time \(T\) (seconds) is modeled by
\[
T\sim \text{Triangular}(k,\,k+r,\,k+3r),\qquad r>0.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Derive \(E[T]\), \(\mathrm{Var}(T)\), and \(\sigma_T\) in terms of \(k\) and \(r\).
  \item State the validity conditions on \(k\) and \(r\), and explain why they are sufficient.
  \item How do \(k\) and \(r\) separately affect center and spread?
  \item Compare shape with the symmetric triangular model having the same endpoints \(k\) and \(k+3r\). Is this model symmetric, left-skewed, or right-skewed?
  \item For fixed \(k\), compare \(r=2\) and \(r=5\): discuss changes in mean, variance, and skewness direction.
\end{enumerate}

\textbf{Solution.}
For \(\text{Triangular}(a,c,b)\):
\[
E[T]=\frac{a+b+c}{3},\qquad
\mathrm{Var}(T)=\frac{a^2+b^2+c^2-ab-ac-bc}{18}.
\]
With \(a=k\), \(c=k+r\), \(b=k+3r\):
\[
E[T]=\frac{k+(k+r)+(k+3r)}{3}=k+\frac{4r}{3}.
\]
For variance,
\[
\mathrm{Var}(T)=\frac{k^2+(k+3r)^2+(k+r)^2-k(k+3r)-k(k+r)-(k+3r)(k+r)}{18}
=\frac{7r^2}{18}.
\]
Thus
\[
\sigma_T=\sqrt{\frac{7r^2}{18}}=r\sqrt{\frac{7}{18}}.
\]
Validity requires \(a\le c\le b\), i.e.
\[
k\le k+r\le k+3r,
\]
which holds whenever \(r\ge 0\); with \(r>0\), spread is nonzero, so the model is valid and non-degenerate for any real \(k\).
Parameter \(k\) shifts location only. Parameter \(r\) increases both mean offset (\(\frac{4r}{3}\)) and spread (\(\propto r^2\)).
For endpoints \(k\) and \(k+3r\), symmetry would need mode at midpoint \(k+\frac{3r}{2}\), but here mode is \(k+r\), left of midpoint; therefore the model is right-skewed.
Comparing \(r=2\) and \(r=5\) (fixed \(k\)): mean increases by \(\frac{4}{3}(5-2)=4\), variance scales from \(\frac{28}{18}\) to \(\frac{175}{18}\), and skewness direction stays right-skewed.

% --------------------------------------------------
\section{Problem 15: Normal with Parameters}

\textbf{Problem.}
The daily amount \(N\) (GB) of network data used by a class is modeled by
\[
N\sim N\!\left(50+3p,\,(2+q)^2\right),\qquad p\in\mathbb{R},\ q>-2.
\]

\textbf{Questions.}
\begin{enumerate}
  \item Derive \(E[N]\), \(\mathrm{Var}(N)\), and \(\sigma_N\) in terms of \(p\) and \(q\).
  \item Explain the role of \(p\) and \(q\): which controls location and which controls spread?
  \item Compare Case A: \((p,q)=(0,1)\) and Case B: \((p,q)=(-2,1)\). What changes in center, concentration around the mean, and tail likelihood?
  \item Compare Case C: \((p,q)=(0,1)\) and Case D: \((p,q)=(0,4)\). What changes in center, spread, concentration near the mean, and likelihood of values far from the mean?
  \item Between models \(M_1:(p,q)=(1,0)\) and \(M_2:(p,q)=(-1,3)\), which model is more likely to produce values at least 15 GB above its own mean? Justify conceptually.
\end{enumerate}

\textbf{Solution.}
For \(N(\mu,\sigma^2)\), \(E[N]=\mu\), \(\mathrm{Var}(N)=\sigma^2\), \(\sigma>0\).
So
\[
E[N]=50+3p,\qquad
\mathrm{Var}(N)=(2+q)^2,\qquad
\sigma_N=2+q \quad (q>-2).
\]
Parameter \(p\) changes only location (horizontal shift). Parameter \(q\) changes spread: larger \(q\) gives larger \(\sigma\), wider curve, lower peak, and heavier tails.
Case A \(\to\) B keeps \(q=1\), so \(\sigma=3\) is unchanged: concentration and tail behavior relative to each mean stay the same, but mean shifts from \(50\) to \(44\) (left by 6).
Case C \(\to\) D keeps \(p=0\), so mean stays 50, but \(\sigma\) increases from \(3\) to \(6\): values are less concentrated near 50 and far-from-mean values become more likely.
For \(M_1\), \(\sigma=2\); for \(M_2\), \(\sigma=5\). A value 15 GB above its own mean is \(15/\sigma\) standard deviations above mean: \(7.5\sigma\) for \(M_1\) versus \(3\sigma\) for \(M_2\). The \(3\sigma\) event is much less extreme, so \(M_2\) is more likely.

\end{document}
