\input{../../../preamble/preamble_exams.tex}

\begin{document}
	\ExamTitleBlock{11th grade}{Term 3 — Lesson 1 Activity (C1--C3): Designing Hypothesis Tests from Context}{}

	\ExamSection{Evaluated criteria}
	\begin{ExamCriteria}
		\item C1: Describe the concept of a statistical hypothesis in context.
		\item C2: Identify the null and alternative hypotheses using correct parameter notation.
		\item C3: Determine whether a test is one tailed or two tailed based on the claim.
	\end{ExamCriteria}

	\ExamSection{Introduction}
	A hypothesis test starts with a contextual benchmark claim about a population. In this activity, you will design that claim yourself: first decide whether your context is about a population mean or a population proportion, then translate the real-world claim into formal hypotheses. The main goal is to build the structure of the test clearly and justify the tail direction from the wording of the alternative claim.

	\subsection*{General problem}
	A benchmark claim may refer either to a population proportion or to a population mean, and in this activity you will create the benchmark from the context.

	For a proportion setting, define a population characteristic that is measurable as success/failure and let the parameter be the true population proportion \(p\). Propose a plausible benchmark value \(p_0\) and explain why that benchmark makes sense in your context.

	For a mean setting, define a measurable quantitative variable and let the parameter be the true population mean \(\mu\). Propose a plausible benchmark value \(\mu_0\) and explain why that benchmark is reasonable for your context.

	In both settings, describe how sample information would be collected, including a sample plan or sample size, invent a plausible sample outcome that could be compared with the benchmark, and choose a significance level.

	\textbf{Questions/Tasks.}
	\begin{enumerate}
		\item[(C1)] For your chosen context, identify whether it is a proportion setting or a mean setting; define the parameter in context; propose and narratively justify a benchmark value; explain how sample information would be collected; invent a plausible sample result; and state the significance level.
		\item[(C2)] Write the null and alternative hypotheses with correct notation, keeping the null as ``parameter equals benchmark''.
		\item[(C3)] Classify the test as left-tailed, right-tailed, or two-tailed, and justify the classification from the wording of the alternative claim.
	\end{enumerate}

	\ExamSection{Problems}
	\begin{ExamProblems}
		\item
		\section{Worked example 1 --- After-school tutoring attendance}
		A school coordination team wants to monitor the participation of Grade 11 students in optional after-school tutoring sessions before major exams. Teachers are concerned that attendance may be lower than expected because many students have sports training or family responsibilities in the afternoon. For planning purposes, the school proposes a benchmark claim that 60\% of Grade 11 students attend at least one tutoring session per week.

		To investigate this claim, the team decides to randomly select 120 Grade 11 students from class lists and ask each student whether they attended at least one tutoring session in the current week. A plausible invented sample outcome is that 64 of the 120 sampled students report attending at least one session. The team chooses a significance level of 5\%.

		\subsection*{C1}
		\begin{itemize}
			\item This is a population proportion setting.
			\item Let \(p\) be the true population proportion of Grade 11 students who attend at least one tutoring session per week.
			\item The benchmark is \(p_0=0.60\).
			\item This benchmark is narratively justified because school planners expect roughly three out of five students to use tutoring during exam preparation weeks.
			\item Sample information would be collected by selecting a random sample from Grade 11 enrollment lists and recording attendance status (yes/no) for each selected student.
			\item The sample plan uses \(n=120\) students.
			\item The invented sample outcome is \(x=64\) students who attended at least one session.
			\item The significance level is \(\alpha=0.05\).
			\item The benchmark and invented sample data are plausible choices for investigating whether attendance is lower than the expected standard.
		\end{itemize}

		\subsection*{C2}
		\(H_0\colon p=0.60,\; H_a\colon p<0.60\), \(\alpha=0.05\).

		\subsection*{C3}
		Left-tailed test because the alternative claim uses ``lower than'' the benchmark.

		\item
		\section{Worked example 2 --- Daily study time at home}
		A learning support committee wants to evaluate whether Grade 11 students are studying more than the minimum expected daily time at home during the term. The committee sets a benchmark of 90 minutes per day as the target average study time, based on the school's study-planning recommendations and homework load.

		To gather evidence, the committee proposes taking a random sample of 40 Grade 11 students and asking them to keep a one-week study log. From those logs, the daily averages per student would be computed, and then the sample mean would be summarized. A plausible invented sample result is a sample mean of \(\bar{x}=98\) minutes per day. Assume the known population standard deviation for daily study time is \(\sigma=22\) minutes, and use a significance level of 1\%.

		\subsection*{C1}
		\begin{itemize}
			\item This is a population mean setting.
			\item Let \(\mu\) be the true population mean daily study time (in minutes) for Grade 11 students.
			\item The benchmark is \(\mu_0=90\) minutes.
			\item This benchmark is narratively justified because the school recommends around 90 minutes of independent study to keep pace with coursework.
			\item Sample information would be collected through one-week study logs from randomly selected students, then summarized as a sample mean.
			\item The sample plan uses \(n=40\) students.
			\item The invented sample outcome is \(\bar{x}=98\) minutes per day.
			\item A known population standard deviation is included as \(\sigma=22\) minutes.
			\item The significance level is \(\alpha=0.01\).
			\item The benchmark and invented sample data are plausible choices for investigating whether average study time is higher than the recommended level.
		\end{itemize}

		\subsection*{C2}
		\(H_0\colon \mu=90,\; H_a\colon \mu>90\), \(\alpha=0.01\).

		\subsection*{C3}
		Right-tailed test because the alternative claim uses ``higher than'' the benchmark.

		\item
		\section{Worked example 3 --- Travel time to school in Bogotá}
		A youth mobility project in Bogotá wants to check whether current commute conditions for Grade 11 students are different from a planning benchmark used in a previous semester. The project uses a benchmark average travel time of 35 minutes from home to school, based on earlier route reports.

		For the new check, surveyors propose sampling 50 Grade 11 students from several neighborhoods and asking each student to record their home-to-school travel time for five consecutive school days. The project would average each student's five-day values and then compute the overall sample mean. A plausible invented sample result is \(\bar{x}=38\) minutes. Assume the known population standard deviation is \(\sigma=9\) minutes, and use a significance level of 5\%.

		\subsection*{C1}
		\begin{itemize}
			\item This is a population mean setting.
			\item Let \(\mu\) be the true population mean home-to-school travel time (in minutes) for Grade 11 students in the target area.
			\item The benchmark is \(\mu_0=35\) minutes.
			\item This benchmark is narratively justified because it comes from past planning reports and is used to monitor changes in mobility conditions.
			\item Sample information would be collected by selecting students across neighborhoods and recording daily travel times over five school days before calculating the sample mean.
			\item The sample plan uses \(n=50\) students.
			\item The invented sample outcome is \(\bar{x}=38\) minutes.
			\item A known population standard deviation is included as \(\sigma=9\) minutes.
			\item The significance level is \(\alpha=0.05\).
			\item The benchmark and invented sample data are plausible choices for investigating whether travel time is different from the historical planning value.
		\end{itemize}

		\subsection*{C2}
		\(H_0\colon \mu=35,\; H_a\colon \mu\ne 35\), \(\alpha=0.05\).

		\subsection*{C3}
		Two-tailed test because the alternative claim uses ``different from'' the benchmark.

		\item
		\section{Activity question 1 --- School context}
		Create a hypothesis test about a measurable feature of your school (for example, average homework time, average cafeteria waiting time, or proportion of students who bring reusable water bottles).
		\begin{enumerate}
			\item[(C1)] Describe the context; state whether it is a population mean or population proportion setting; define the parameter; propose a benchmark claim; justify the benchmark narratively; explain how you would collect sample information; invent a plausible sample result; and state a significance level.
			\item[(C2)] Write \(H_0\) and \(H_a\) using correct notation, with the null in the form ``parameter equals benchmark''.
			\item[(C3)] Classify your test as left-tailed, right-tailed, or two-tailed, and justify the classification from the wording of your alternative claim.
		\end{enumerate}

		\item
		\section{Activity question 2 --- Home context}
		Create a hypothesis test about a measurable feature of life at home (for example, average nightly sleep time, proportion of nights with at least one hour of study, or average daily screen time at home).
		\begin{enumerate}
			\item[(C1)] Describe the context; state whether it is a population mean or population proportion setting; define the parameter; propose a benchmark claim; justify the benchmark narratively; explain how you would collect sample information; invent a plausible sample result; and state a significance level.
			\item[(C2)] Write \(H_0\) and \(H_a\) using correct notation, with the null in the form ``parameter equals benchmark''.
			\item[(C3)] Classify your test as left-tailed, right-tailed, or two-tailed, and justify the classification from the wording of your alternative claim.
		\end{enumerate}

		\item
		\section{Activity question 3 --- Neighborhood context}
		Create a hypothesis test about a measurable feature of your neighborhood (for example, average time to reach a bus stop, proportion of households that separate recyclable waste, or average weekly time spent in neighborhood sports or cultural spaces).
		\begin{enumerate}
			\item[(C1)] Describe the context; state whether it is a population mean or population proportion setting; define the parameter; propose a benchmark claim; justify the benchmark narratively; explain how you would collect sample information; invent a plausible sample result; and state a significance level.
			\item[(C2)] Write \(H_0\) and \(H_a\) using correct notation, with the null in the form ``parameter equals benchmark''.
			\item[(C3)] Classify your test as left-tailed, right-tailed, or two-tailed, and justify the classification from the wording of your alternative claim.
		\end{enumerate}

		\item
		\section{Activity question 4 --- Your Grade 11 group context}
		Create a hypothesis test about a measurable feature of your own Grade 11 group (for example, average quiz score in a subject, proportion of students who submit assignments on time, or average number of reading pages completed per week).
		\begin{enumerate}
			\item[(C1)] Describe the context; state whether it is a population mean or population proportion setting; define the parameter; propose a benchmark claim; justify the benchmark narratively; explain how you would collect sample information; invent a plausible sample result; and state a significance level.
			\item[(C2)] Write \(H_0\) and \(H_a\) using correct notation, with the null in the form ``parameter equals benchmark''.
			\item[(C3)] Classify your test as left-tailed, right-tailed, or two-tailed, and justify the classification from the wording of your alternative claim.
		\end{enumerate}

		\item
		\section{Activity question 5 --- Personal context}
		Create a hypothesis test about a measurable feature related to yourself (for example, your average daily study minutes across a month, the proportion of days you complete your planned study checklist, or your average time spent exercising per week).
		\begin{enumerate}
			\item[(C1)] Describe the context; state whether it is a population mean or population proportion setting; define the parameter; propose a benchmark claim; justify the benchmark narratively; explain how you would collect sample information; invent a plausible sample result; and state a significance level.
			\item[(C2)] Write \(H_0\) and \(H_a\) using correct notation, with the null in the form ``parameter equals benchmark''.
			\item[(C3)] Classify your test as left-tailed, right-tailed, or two-tailed, and justify the classification from the wording of your alternative claim.
		\end{enumerate}

		\item
		\section{Activity question 6 --- Bogotá context}
		Create a hypothesis test about a measurable feature of Bogotá that can be investigated at your level (for example, average time students spend commuting, proportion of residents who use public transportation for daily travel, or average weekly participation in public cultural activities).
		\begin{enumerate}
			\item[(C1)] Describe the context; state whether it is a population mean or population proportion setting; define the parameter; propose a benchmark claim; justify the benchmark narratively; explain how you would collect sample information; invent a plausible sample result; and state a significance level.
			\item[(C2)] Write \(H_0\) and \(H_a\) using correct notation, with the null in the form ``parameter equals benchmark''.
			\item[(C3)] Classify your test as left-tailed, right-tailed, or two-tailed, and justify the classification from the wording of your alternative claim.
		\end{enumerate}

	\end{ExamProblems}

\end{document}
