\makeatletter
\def\input@path{{./}{../}{../../}{../../../}{../../../../}{preamble/}{../preamble/}{../../preamble/}{../../../preamble/}}
\makeatother

\input{preamble.tex}

\title{Grade 10 - Term 3 \\ Learning Evidence 4 Activity (C8--C9)}
\author{}
\date{}

\begin{document}
	\maketitle
	
	\section*{Learning Objective and Assessed Criteria}
	\textbf{Learning objective.} Create interactive computational tools for modeling continuous probability distributions by graphing PDFs, calculating probabilities under curves, estimating parameters from data, and comparing fitted distributions with empirical histograms.
	
	\vspace{0.2cm}
	\textbf{Assessed criteria.}
	\begin{itemize}
		\item \textbf{C8.} Creates an interactive webpage that represents continuous probability distributions and allows users to modify parameters and calculate probabilities under the PDF.
		\item \textbf{C9.} Uses AI-assisted programming to create a script that estimates distribution parameters from data and compares the fitted PDF with the empirical histogram.
	\end{itemize}

	\section{Background from C1--C7}
	In C1--C4, you modeled continuous random variables with linear and piecewise probability density functions (PDFs), ensuring non-negativity and total area equal to 1. In C5--C7, you worked with the normal distribution, standard-normal calculations, and inverse-normal reasoning.
	
	In this activity, you will extend that work into a computational project that combines probability modeling, visualization, and AI-assisted coding.

	\section{Task Overview}
	You will build a student-friendly interactive webpage and an AI-assisted parameter estimation script.
	\begin{itemize}
		\item Part A (C8): interactive visualization and probability calculations for multiple continuous distributions.
		\item Part B (C9): data-driven parameter estimation and fit comparison (histogram vs fitted PDF).
	\end{itemize}
	
	\textbf{Important:} This is a task document. Do not submit only screenshots or only code fragments. You must provide all required files and evidence listed below.

	\section{Mandatory Requirements for C8 (Interactive Webpage)}
	Your webpage must include interactive PDF graphs for the following distributions:
	\begin{enumerate}
		\item Uniform distribution.
		\item Triangular distribution.
		\item Increasing or decreasing linear distribution.
		\item Piecewise distribution.
		\item Normal distribution.
	\end{enumerate}
	
	For each distribution, your webpage must:
	\begin{itemize}
		\item Allow the user to enter or modify all defining parameters.
		\item Generate the corresponding PDF from the parameters.
		\item Display the PDF graph interactively.
		\item Compute probabilities under the curve for:
		\begin{itemize}
			\item probability below a point,
			\item probability above a point,
			\item probability between two points.
		\end{itemize}
		\item Visually shade the selected probability region.
		\item Display the numerical probability value.
		\item Update the graph and values whenever parameters or bounds change.
	\end{itemize}

	\subsection*{Distribution-Specific C8 Conditions}
	\begin{itemize}
		\item \textbf{Uniform:} input lower endpoint and upper endpoint; generate a constant PDF.
		\item \textbf{Triangular:} input lower endpoint, upper endpoint, and mode; generate a triangular PDF.
		\item \textbf{Linear:} define an increasing or decreasing linear PDF on a finite interval; ensure total area under the PDF is 1.
		\item \textbf{Piecewise:} define at least one piecewise PDF using intervals and formulas; graph all pieces correctly.
		\item \textbf{Normal:} input mean and standard deviation; graph the normal PDF.
	\end{itemize}

	\section{Mandatory Requirements for C9 (AI-Assisted Parameter Estimation)}
	Use AI to help you create a script that receives data and estimates distribution parameters.
	
	The script must support at least:
	\begin{enumerate}
		\item Uniform distribution (estimate lower and upper endpoints).
		\item Triangular distribution (estimate lower endpoint, upper endpoint, and mode).
		\item Normal distribution (estimate mean and standard deviation).
	\end{enumerate}
	
	Your webpage must allow the user to upload or input data and choose which distribution to fit.
	
	For each fit, display:
	\begin{enumerate}
		\item estimated parameters,
		\item fitted PDF,
		\item histogram of uploaded data,
		\item histogram and fitted PDF on the same graph,
		\item short interpretation of whether the fitted distribution appears reasonable.
	\end{enumerate}

	\subsection*{Required Documentation of AI Use}
	You may use AI tools, but you must document your process. Submit:
	\begin{enumerate}
		\item prompt(s) used,
		\item AI-generated code or code fragments,
		\item short explanation of what you modified or verified,
		\item short explanation of how you tested correctness.
	\end{enumerate}

	\section{Validation Checklist (Minimum for Acceptance)}
	Your submission is valid only if all of the following are true:
	\begin{enumerate}
		\item PDF changes when parameters are modified.
		\item Shaded probability region changes when bounds are modified.
		\item Numerical probability updates correctly.
		\item Uploaded/entered data is processed successfully.
		\item Estimated parameters are displayed.
		\item Histogram appears on the same graph as the fitted PDF.
		\item You can explain the meaning of each parameter.
		\item The webpage works for at least one example of each required distribution type.
	\end{enumerate}

	\section{Required Deliverables}
	Submit all of the following:
	\begin{enumerate}
		\item Webpage files.
		\item AI-assisted parameter estimation script.
		\item At least three test datasets:
		\begin{itemize}
			\item one uniform-like dataset,
			\item one triangular-like dataset,
			\item one normal-like dataset.
		\end{itemize}
		\item Screenshots showing the webpage working.
		\item Short written explanation of how the webpage works.
		\item Short reflection on AI prompts and corrections made to generated code.
	\end{enumerate}

	\section{Evaluation Checklist}
	Use this checklist before submission.
	\begin{itemize}
		\item[\(\square\)] I included all required distributions in C8.
		\item[\(\square\)] I included all required probability calculations and shaded regions.
		\item[\(\square\)] My graphs update dynamically when values change.
		\item[\(\square\)] My C9 script estimates parameters for uniform, triangular, and normal models.
		\item[\(\square\)] I plotted histogram and fitted PDF together.
		\item[\(\square\)] I documented AI prompts, outputs, edits, and tests.
		\item[\(\square\)] I prepared all deliverables in organized folders.
	\end{itemize}

	\section{Submission Instructions}
	\begin{enumerate}
		\item Organize your files in clearly named folders (webpage, script, datasets, evidence).
		\item Verify that all links, scripts, and assets run on the school computer.
		\item Include your written explanation and AI-use reflection as PDF or text file.
		\item Submit before the deadline indicated by your teacher.
	\end{enumerate}
	
	\vspace{0.2cm}
	\textbf{Reminder:} Do not submit solved mathematics exercises. Submit a functional project with evidence of development, testing, and responsible AI use.

\end{document}
