\input{../../preamble/preamble_exams.tex}
\begin{document}
	\ExamTitleBlock{11th grade}{Term 2 Midterm Mock (Hard): Confidence Interval Planning (Solutions)}{}

\section*{Contents}
\addcontentsline{toc}{section}{Contents}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{1}
\tableofcontents

	\ExamSection{Evaluated criteria}
\addcontentsline{toc}{section}{Evaluated criteria}
	\begin{ExamCriteria}
		\item C1: Compute the sample mean and sample standard deviation.
		\item C2: Distinguish between sample statistics and population parameters.
		\item C3: Explain why interval estimation is preferred over a single point estimate.
		\item C4: Interpret the meaning of a X\% confidence interval in context.
		\item C5: Construct a X\% confidence interval using known population standard deviation.
		\item C6: Determine the required sample size to achieve a target margin of error.
		\item C7: Construct and interpret a confidence interval for a population proportion.
	\end{ExamCriteria}

	\ExamSection{Problems}
\addcontentsline{toc}{section}{Problems}
	\begin{ExamProblems}
		\newpage
		\item
		\textbf{Problem description}
		A regional rail authority compares average monthly maintenance cost per trainset (thousand USD) in two depots after introducing predictive maintenance software.
		The grouped data report maintenance-cost values together with their frequencies, where each cost value occurs in a certain number of trainsets.

		\textbf{Depot A:} The data correspond to 27 trainsets. Each maintenance-cost value below is measured in thousand USD, and the stated frequency indicates how many trainsets had that value. Cost = 72 (thousand USD) occurs in 5 trainsets; cost = 78 (thousand USD) occurs in 7 trainsets; cost = 85 (thousand USD) occurs in 6 trainsets; cost = 92 (thousand USD) occurs in 5 trainsets; and cost = 100 (thousand USD) occurs in 4 trainsets.

		\textbf{Depot B:} The data correspond to 28 trainsets. Each maintenance-cost value below is measured in thousand USD, and the stated frequency indicates how many trainsets had that value. Cost = 68 (thousand USD) occurs in 4 trainsets; cost = 75 (thousand USD) occurs in 6 trainsets; cost = 83 (thousand USD) occurs in 8 trainsets; cost = 90 (thousand USD) occurs in 6 trainsets; and cost = 98 (thousand USD) occurs in 4 trainsets.

		From long-run accounting records, the population standard deviation is known to be $\sigma = 14$ thousand USD for both depots.
		\begin{enumerate}
			\item Calculate the sample mean and sample standard deviation for each depot using the provided grouped data (C1).
			\item Construct 90\%, 95\%, and 99\% confidence intervals for each depot mean using known $\sigma$, and report the margin of error for each interval (C5).
			\item Compute required sample sizes for target margins of error $E=3.5$ and $E=2.8$ at 90\%, 95\%, and 99\%, and determine additional observations needed for each depot (C6).
		\end{enumerate}

		\subsection*{C1}
		\textbf{Depot A}
		\[
		\sum fx = 72(5)+78(7)+85(6)+92(5)+100(4)=2276,
		\qquad \bar{x}_A=\frac{2276}{27}=84.30
		\]
		\[
		\sum f(x-\bar{x}_A)^2=2323.48,
		\quad s_A^2=\frac{2323.48}{26}=89.36,
		\quad s_A\approx 9.45
		\]
		\textbf{Depot B}
		\[
		\sum fx = 68(4)+75(6)+83(8)+90(6)+98(4)=2318,
		\qquad \bar{x}_B=\frac{2318}{28}=82.79
		\]
		\[
		\sum f(x-\bar{x}_B)^2=2475.64,
		\quad s_B^2=\frac{2475.64}{27}=91.69,
		\quad s_B\approx 9.58
		\]

		\subsection*{C5}
		For Depot A, $SE_A=\frac{14}{\sqrt{27}}\approx2.694$.
		\[
		E_{A,90}=1.645(2.694)=4.43,\ \ CI_{90}:\ 84.30\pm4.43=[79.87,88.73]
		\]
		\[
		E_{A,95}=1.96(2.694)=5.28,\ \ CI_{95}:\ 84.30\pm5.28=[79.02,89.58]
		\]
		\[
		E_{A,99}=2.576(2.694)=6.94,\ \ CI_{99}:\ 84.30\pm6.94=[77.36,91.24]
		\]
		For Depot B, $SE_B=\frac{14}{\sqrt{28}}\approx2.646$.
		\[
		E_{B,90}=1.645(2.646)=4.35,\ \ CI_{90}:\ 82.79\pm4.35=[78.43,87.14]
		\]
		\[
		E_{B,95}=1.96(2.646)=5.19,\ \ CI_{95}:\ 82.79\pm5.19=[77.60,87.97]
		\]
		\[
		E_{B,99}=2.576(2.646)=6.82,\ \ CI_{99}:\ 82.79\pm6.82=[75.97,89.60]
		\]

		\subsection*{C6}
		Use $n=\left(\frac{z^*\sigma}{E}\right)^2$ and round up.
		For $E=3.5$:
		\[
		n_{90}=\left(\frac{1.645\cdot14}{3.5}\right)^2=43.32\Rightarrow44,
		n_{95}=61.47\Rightarrow62,
		n_{99}=106.15\Rightarrow107
		\]
		Additional observations:
		Depot A ($n=27$): $17,35,80$; Depot B ($n=28$): $16,34,79$.
		For $E=2.8$:
		\[
		n_{90}=67.68\Rightarrow68,
		n_{95}=96.05\Rightarrow97,
		n_{99}=165.86\Rightarrow166
		\]
		Additional observations:
		Depot A: $41,70,139$; Depot B: $40,69,138$.

		\newpage
		\item
		\textbf{Problem description}
		A digital payments platform tracks average fraud-review handling time (minutes) for two independent teams.
		The grouped data are reported using class intervals, and the frequency for each interval indicates how many cases fall within that time range.

		\textbf{Team A:} The data correspond to 38 cases. Each interval below is measured in minutes, and the stated frequency is the number of cases in that interval. 42--46 minutes corresponds to 6 cases; 47--51 minutes corresponds to 9 cases; 52--56 minutes corresponds to 10 cases; 57--61 minutes corresponds to 8 cases; and 62--66 minutes corresponds to 5 cases.

		\textbf{Team B:} The data correspond to 40 cases. Each interval below is measured in minutes, and the stated frequency is the number of cases in that interval. 40--44 minutes corresponds to 5 cases; 45--49 minutes corresponds to 8 cases; 50--54 minutes corresponds to 11 cases; 55--59 minutes corresponds to 9 cases; and 60--64 minutes corresponds to 7 cases.

		Assume known population standard deviation $\sigma=6.5$ minutes.
		\begin{enumerate}
			\item Calculate the sample mean and sample standard deviation for each team using the provided grouped data (C1).
			\item Construct 90\%, 95\%, and 99\% confidence intervals for each team mean using known $\sigma$, and report the margin of error for each interval (C5).
			\item Compute required sample sizes for $E=1.2$ and $E=0.9$ at 90\%, 95\%, and 99\%, and determine additional cases needed for each team (C6).
		\end{enumerate}

		\subsection*{C1}
		Use class midpoints.
		For Team A, midpoints are 44,49,54,59,64.
		\[
		\sum fm=44(6)+49(9)+54(10)+59(8)+64(5)=2037,
		\quad \bar{x}_A=\frac{2037}{38}=53.61
		\]
		\[
		\sum f(m-\bar{x}_A)^2=1520.86,
		\quad s_A^2=\frac{1520.86}{37}=41.10,
		\quad s_A\approx6.41
		\]
		For Team B, midpoints are 42,47,52,57,62.
		\[
		\sum fm=42(5)+47(8)+52(11)+57(9)+62(7)=2105,
		\quad \bar{x}_B=\frac{2105}{40}=52.63
		\]
		\[
		\sum f(m-\bar{x}_B)^2=1610.94,
		\quad s_B^2=\frac{1610.94}{39}=41.31,
		\quad s_B\approx6.42
		\]

		\subsection*{C5}
		\[
		SE_A=\frac{6.5}{\sqrt{38}}=1.055,
		\ E_{A,90}=1.645(1.055)=1.73,
		\ E_{A,95}=1.96(1.055)=2.07,
		\ E_{A,99}=2.576(1.055)=2.72
		\]
		\[
		CI_{A,90}=[51.87,55.34],\quad CI_{A,95}=[51.54,55.67],\quad CI_{A,99}=[50.89,56.32]
		\]
		\[
		SE_B=\frac{6.5}{\sqrt{40}}=1.028,
		\ E_{B,90}=1.69,
		\ E_{B,95}=2.01,
		\ E_{B,99}=2.65
		\]
		\[
		CI_{B,90}=[50.93,54.32],\quad CI_{B,95}=[50.61,54.64],\quad CI_{B,99}=[49.98,55.27]
		\]

		\subsection*{C6}
		For $E=1.2$:
		\[
		n_{90}=\left(\frac{1.645\cdot6.5}{1.2}\right)^2=79.36\Rightarrow80,
		n_{95}=112.69\Rightarrow113,
		n_{99}=194.65\Rightarrow195
		\]
		Additional cases needed:
		Team A ($n=38$): $42,75,157$; Team B ($n=40$): $40,73,155$.
		For $E=0.9$:
		\[
		n_{90}=141.08\Rightarrow142,
		n_{95}=200.34\Rightarrow201,
		n_{99}=346.04\Rightarrow347
		\]
		Additional cases needed:
		Team A: $104,163,309$; Team B: $102,161,307$.

		\newpage
		\item
		\textbf{Problem description}
		An emergency logistics center records warehouse dispatch completion times (hours) for one expansion zone.
		A random sample of 18 dispatches is:
		31, 29, 34, 28, 33, 35, 30, 32, 36, 27, 34, 31, 29, 33, 30, 37, 28, 32.
		Known population standard deviation is $\sigma=5.8$ hours.
		\begin{enumerate}
			\item Calculate the sample mean and sample standard deviation for the dispatch-time sample (C1).
			\item Construct 90\%, 95\%, and 99\% confidence intervals for the population mean completion time using known $\sigma$, and report the margin of error for each interval (C5).
			\item Compute required sample sizes for target errors $E=2.0$ and $E=1.5$ at 90\%, 95\%, and 99\%, and determine additional dispatches needed (C6).
		\end{enumerate}

		\subsection*{C1}
		\[
		\bar{x}=\frac{569}{18}=31.61
		\]
		\[
		\sum (x_i-\bar{x})^2=142.28,
		\quad s^2=\frac{142.28}{17}=8.37,
		\quad s=2.89
		\]

		\subsection*{C5}
		\[
		SE=\frac{5.8}{\sqrt{18}}=1.367
		\]
		\[
		E_{90}=1.645(1.367)=2.25,\quad CI_{90}=31.61\pm2.25=[29.36,33.86]
		\]
		\[
		E_{95}=1.96(1.367)=2.68,\quad CI_{95}=31.61\pm2.68=[28.93,34.29]
		\]
		\[
		E_{99}=2.576(1.367)=3.52,\quad CI_{99}=31.61\pm3.52=[28.09,35.13]
		\]

		\subsection*{C6}
		For $E=2.0$:
		\[
		n_{90}=\left(\frac{1.645\cdot5.8}{2.0}\right)^2=22.75\Rightarrow23,
		n_{95}=32.31\Rightarrow33,
		n_{99}=55.83\Rightarrow56
		\]
		Additional dispatches from current $n=18$: $5,15,38$.
		For $E=1.5$:
		\[
		n_{90}=40.44\Rightarrow41,
		n_{95}=57.43\Rightarrow58,
		n_{99}=99.26\Rightarrow100
		\]
		Additional dispatches: $23,40,82$.

		\newpage
		\item
		\textbf{Problem description}
		A tax administration compares average daily audit value recovered (thousand USD) by two independent field units.

		\textbf{Unit A (15 days):} 112, 108, 115, 121, 117, 109, 114, 118, 116, 113, 120, 111, 119, 107, 122.

		\textbf{Unit B (18 days):} 105, 109, 111, 114, 108, 112, 116, 110, 113, 107, 115, 118, 106, 109, 117, 111, 114, 108.

		Assume known $\sigma=9.2$ thousand USD in both units.
		\begin{enumerate}
			\item Calculate the sample mean and sample standard deviation for each unit using the provided daily values (C1).
			\item Construct 90\%, 95\%, and 99\% confidence intervals for each unit mean using known $\sigma$, and report the margin of error for each interval (C5).
			\item Compute required sample sizes for $E=3.0$ and $E=2.2$ at 90\%, 95\%, and 99\%, and determine additional observations needed for each unit (C6).
		\end{enumerate}

		\subsection*{C1}
		\[
		\bar{x}_A=\frac{1722}{15}=114.80,
		\quad s_A\approx4.77
		\]
		\[
		\bar{x}_B=\frac{2003}{18}=111.28,
		\quad s_B\approx3.85
		\]

		\subsection*{C5}
		For Unit A, $SE_A=\frac{9.2}{\sqrt{15}}=2.375$.
		\[
		E_{A,90}=3.91,\ CI_{A,90}=[110.89,118.71]
		\]
		\[
		E_{A,95}=4.66,\ CI_{A,95}=[110.14,119.46]
		\]
		\[
		E_{A,99}=6.12,\ CI_{A,99}=[108.68,120.92]
		\]
		For Unit B, $SE_B=\frac{9.2}{\sqrt{18}}=2.168$.
		\[
		E_{B,90}=3.57,\ CI_{B,90}=[107.71,114.84]
		\]
		\[
		E_{B,95}=4.25,\ CI_{B,95}=[107.03,115.53]
		\]
		\[
		E_{B,99}=5.59,\ CI_{B,99}=[105.69,116.86]
		\]

		\subsection*{C6}
		For $E=3.0$:
		\[
		n_{90}=25.44\Rightarrow26,\quad n_{95}=36.14\Rightarrow37,\quad n_{99}=62.44\Rightarrow63
		\]
		Additional days needed:
		Unit A ($n=15$): $11,22,48$; Unit B ($n=18$): $8,19,45$.
		For $E=2.2$:
		\[
		n_{90}=47.28\Rightarrow48,\quad n_{95}=67.17\Rightarrow68,\quad n_{99}=116.11\Rightarrow117
		\]
		Additional days needed:
		Unit A: $33,53,102$; Unit B: $30,50,99$.

		\newpage
		\item
		\textbf{Problem description}
		A central bank supervision office evaluates a fintech lender.
		One team reports from a recent compliance review: average repayment delay was 4.7 days with spread 2.1 days, and a 95\% confidence interval for the mean delay was [4.1, 5.3].
		A second team cites long-run portfolio values from the full national borrower base: true average delay = 4.9 days and true standard deviation = 2.4 days.
		The policy division also references a strategic threshold: ``acceptable mean delay below 5.0 days.''

		Explicit analysis tasks:
		1) Classify every numerical quantity above as a statistic from observed data, a fixed characteristic of the full borrower base, or a policy benchmark, and justify (C2).
		2) Using the reported interval for mean repayment delay, answer all of the following (C3): (i) state the point estimate for the mean repayment delay; (ii) explain, in this supervision context, why reporting only that point estimate is insufficient compared to reporting the interval; (iii) compute the interval width and explain what it says about precision; (iv) propose two concrete changes to the data-collection or measurement design that would narrow the interval, and justify briefly.
		3) State which mean values are ruled out/plausible under [4.1, 5.3] and explain ``95\% confidence'' in repeated-trial language (C4).

		\subsection*{C2}
		Observed-data statistics: sample mean $4.7$, sample spread $2.1$, and interval endpoints $4.1$ and $5.3$.
		Full-population characteristics: true mean delay $\mu=4.9$ and true standard deviation $\sigma=2.4$.
		Policy benchmark: acceptable-delay threshold $5.0$ days.

		\subsection*{C3}
		(i) Point estimate: $\hat{\mu}=4.7$ days.

		(ii) Why point estimate alone is insufficient: $4.7$ does not show uncertainty; the interval [4.1, 5.3] shows a range of plausible means and makes risk-aware supervision possible.

		(iii) Width: $5.3-4.1=1.2$ days. This is a moderate precision level; a narrower width would indicate greater precision.

		(iv) Two changes to narrow the interval: increase sample size (reduces standard error) and improve measurement consistency (reduces variability, e.g., standardized delay-timestamp rules).

		\subsection*{C4}
		Plausible means under this report are between 4.1 and 5.3 days; means below 4.1 or above 5.3 are ruled out by this 95\% interval procedure.
		In repeated comparable samples, about 95\% of intervals built this same way would capture the true mean delay.

		\newpage
		\item
		\textbf{Problem description}
		A national logistics ministry monitors truck turnaround time at border ports.
		A pilot review from Port North gives mean 11.8 hours and 90\% CI [10.9, 12.7].
		A separate audit from Port South gives mean 12.4 hours and 90\% CI [11.6, 13.2].
		Historic all-port values for the old regime were true average turnaround = 12.1 hours and true standard deviation = 3.0 hours.
		A policy KPI requires maintaining average turnaround at or below 12.0 hours.

		Explicit analysis tasks:
		1) Classify observed-data quantities, full-system characteristics, and policy quantities (C2).
		2) Using the reported intervals for mean turnaround time, answer all of the following (C3): (i) state the point estimate for Port North and the point estimate for Port South; (ii) explain, in this policy-decision context, why reporting only those point estimates is insufficient compared to reporting intervals; (iii) compare precision by computing the widths of the North and South intervals and identify which report is more precise; (iv) propose two concrete changes to the data-collection or measurement design that would narrow the interval(s), and justify briefly.
		3) For each port, state ruled-out/plausible means and explain operational meaning of 90\% confidence (C4).

		\subsection*{C2}
		Observed-data statistics: means 11.8 and 12.4 with intervals [10.9, 12.7] and [11.6, 13.2].
		Full-system characteristics (historic regime): true mean 12.1 and true standard deviation 3.0.
		Policy quantity: KPI target 12.0 hours.

		\subsection*{C3}
		(i) Point estimates: Port North $=11.8$, Port South $=12.4$ hours.

		(ii) Why intervals are needed: point estimates alone cannot show uncertainty around each port's true mean, which is crucial for policy decisions near the KPI cutoff.

		(iii) Interval widths:
		\[
		W_N=12.7-10.9=1.8,\qquad W_S=13.2-11.6=1.6
		\]
		Port South is slightly more precise (narrower interval).

		(iv) Two ways to narrow intervals: increase sample size at each port and reduce process/measurement variability (standardized timing protocol).

		\subsection*{C4}
		Port North plausible means: 10.9 to 12.7; ruled out: outside that range.
		Port South plausible means: 11.6 to 13.2; ruled out: outside that range.
		Operationally, 90\% confidence means this interval-construction method would capture each true port mean in about 90\% of repeated comparable studies.

		\newpage
		\item
		\textbf{Problem description}
		A social protection agency studies benefit-payment processing time across two digital channels.
		Channel A monitoring report gives mean 6.3 days with CI [5.8, 6.8] at 95\%.
		Channel B audit report gives mean 6.0 days with CI [5.2, 6.8] at 95\%.
		Legacy national values before automation were true average processing time = 6.7 days and true standard deviation = 1.9 days.
		The ministry also publishes two benchmarks: warning level 6.5 days, excellent level 6.0 days.

		Explicit analysis tasks:
		1) Categorize all values into observed-data statistics, fixed national characteristics, and policy benchmarks (C2).
		2) Using the reported intervals for mean processing time, answer all of the following (C3): (i) state the point estimate for Channel A and the point estimate for Channel B; (ii) explain, in this social-protection context, why reporting only those point estimates is insufficient compared to reporting intervals; (iii) compare precision by computing the widths of the Channel A and Channel B intervals and identify which report is more precise; (iv) propose two concrete changes to the data-collection or measurement design that would narrow the interval(s), and justify briefly.
		3) Determine plausible/ruled-out means per channel and explain repeated-trial meaning of 95\% confidence (C4).

		\subsection*{C2}
		Observed-data statistics: means 6.3 and 6.0 with intervals [5.8, 6.8] and [5.2, 6.8].
		Fixed national characteristics (legacy): true mean 6.7 and true standard deviation 1.9.
		Policy benchmarks: warning 6.5 days and excellent 6.0 days.

		\subsection*{C3}
		(i) Point estimates: Channel A $=6.3$, Channel B $=6.0$ days.

		(ii) Point estimates alone are insufficient because they do not show uncertainty; intervals show plausible ranges and support safer operational decisions.

		(iii) Widths:
		\[
		W_A=6.8-5.8=1.0,\qquad W_B=6.8-5.2=1.6
		\]
		Channel A is more precise.

		(iv) Two changes to narrow intervals: increase sample size and improve data-quality consistency (uniform timestamp and case-classification rules).

		\subsection*{C4}
		Channel A plausible means: 5.8 to 6.8; ruled out: outside this interval.
		Channel B plausible means: 5.2 to 6.8; ruled out: outside this interval.
		Repeated-trial meaning: about 95\% of intervals built with this method would contain the true channel mean.

		\newpage
		\item
		\textbf{Problem description}
		A sovereign wealth monitoring office evaluates monthly return volatility reporting from two external fund managers.
		Manager Alpha reports a mean monthly return of 1.4\% with 95\% CI [0.8\%, 2.0\%].
		Manager Beta reports a mean of 1.1\% with 95\% CI [0.2\%, 2.0\%].
		Historical long-run values from the full investment universe are true average monthly return = 1.3\% and true standard deviation = 3.5\%.
		The board uses two policy thresholds: ``underperforming'' if mean return $<1.0\%$ and ``stretch target'' if mean return $\ge 1.8\%$.

		Explicit analysis tasks:
		1) Classify all quantities into observed-data statistics, fixed universe characteristics, and policy thresholds (C2).
		2) Using the reported intervals for mean monthly return, answer all of the following (C3): (i) state the point estimate for Manager Alpha and the point estimate for Manager Beta; (ii) explain, in this strategic-allocation context, why reporting only those point estimates is insufficient compared to reporting intervals; (iii) compare precision by computing the widths of the Alpha and Beta intervals and identify which report is more precise; (iv) propose two concrete changes to the data-collection or measurement design that would narrow the interval(s), and justify briefly.
		3) For each manager, state ruled-out and plausible values and explain repeated-trial meaning of 95\% confidence (C4).

		\subsection*{C2}
		Observed-data statistics: means 1.4\% and 1.1\%, with intervals [0.8\%, 2.0\%] and [0.2\%, 2.0\%].
		Fixed universe characteristics: true mean 1.3\% and true standard deviation 3.5\%.
		Policy thresholds: 1.0\% (underperforming) and 1.8\% (stretch target).

		\subsection*{C3}
		(i) Point estimates: Alpha $=1.4\%$, Beta $=1.1\%$.

		(ii) Intervals are needed because point estimates do not quantify uncertainty around each manager's true mean return.

		(iii) Widths:
		\[
		W_{\alpha}=2.0-0.8=1.2\%,\qquad W_{\beta}=2.0-0.2=1.8\%
		\]
		Manager Alpha is more precise.

		(iv) Two changes to narrow intervals: use larger sample windows and improve data consistency (standardized reporting periods and return definitions).

		\subsection*{C4}
		Alpha plausible means: 0.8\% to 2.0\%; ruled out: outside this range.
		Beta plausible means: 0.2\% to 2.0\%; ruled out: outside this range.
		Repeated-trial meaning: about 95\% of intervals from this method would contain the true manager mean return.

		\newpage
		\item
		\textbf{Problem description}
		A public procurement platform reports supplier on-time document submission rates for two regions.
		Region East sample: 420 suppliers with 246 on-time.
		Region West sample: 380 suppliers with 186 on-time.
		\begin{enumerate}
			\item Construct 90\%, 95\%, and 99\% confidence intervals for each region's population proportion, and report the margin of error for each interval (C7).
		\end{enumerate}

		\subsection*{C7}
		\textbf{Region East}
		\[
		\hat{p}_E=\frac{246}{420}=0.5857,
		\quad SE_E=\sqrt{\frac{0.5857(0.4143)}{420}}=0.0240
		\]
		\[
		E_{90}=1.645(0.0240)=0.0395,\quad CI_{90}=[0.5462,0.6253]
		\]
		\[
		E_{95}=1.96(0.0240)=0.0471,\quad CI_{95}=[0.5386,0.6328]
		\]
		\[
		E_{99}=2.576(0.0240)=0.0619,\quad CI_{99}=[0.5238,0.6476]
		\]

		\textbf{Region West}
		\[
		\hat{p}_W=\frac{186}{380}=0.4895,
		\quad SE_W=\sqrt{\frac{0.4895(0.5105)}{380}}=0.0256
		\]
		\[
		E_{90}=1.645(0.0256)=0.0422,\quad CI_{90}=[0.4473,0.5317]
		\]
		\[
		E_{95}=1.96(0.0256)=0.0503,\quad CI_{95}=[0.4392,0.5397]
		\]
		\[
		E_{99}=2.576(0.0256)=0.0661,\quad CI_{99}=[0.4234,0.5555]
		\]

		\newpage
		\item
		\textbf{Problem description}
		A municipal e-governance program evaluates digital tax filing completion by business size subgroup in one city.
		Small firms: $n=250$, completed $x=142$.
		Medium firms: $n=200$, completed $x=96$.
		Large firms: $n=150$, completed $x=57$.
		\begin{enumerate}
			\item Construct 90\% and 95\% confidence intervals for each subgroup proportion, and report the margin of error for each interval (C7).
		\end{enumerate}

		\subsection*{C7}
		\textbf{Small firms}
		\[
		\hat{p}_S=\frac{142}{250}=0.568,
		\quad SE_S=\sqrt{\frac{0.568(0.432)}{250}}=0.0313
		\]
		\[
		E_{90}=1.645(0.0313)=0.0515,\quad CI_{90}=[0.5165,0.6195]
		\]
		\[
		E_{95}=1.96(0.0313)=0.0614,\quad CI_{95}=[0.5066,0.6294]
		\]

		\textbf{Medium firms}
		\[
		\hat{p}_M=\frac{96}{200}=0.480,
		\quad SE_M=\sqrt{\frac{0.48(0.52)}{200}}=0.0353
		\]
		\[
		E_{90}=1.645(0.0353)=0.0581,\quad CI_{90}=[0.4219,0.5381]
		\]
		\[
		E_{95}=1.96(0.0353)=0.0692,\quad CI_{95}=[0.4108,0.5492]
		\]

		\textbf{Large firms}
		\[
		\hat{p}_L=\frac{57}{150}=0.380,
		\quad SE_L=\sqrt{\frac{0.38(0.62)}{150}}=0.0396
		\]
		\[
		E_{90}=1.645(0.0396)=0.0652,\quad CI_{90}=[0.3148,0.4452]
		\]
		\[
		E_{95}=1.96(0.0396)=0.0777,\quad CI_{95}=[0.3023,0.4577]
		\]

		\newpage
		\item
		\textbf{Problem description}
		A pharmaceutical procurement unit tracks defective shipment proportion in two independent quarterly audits.
		Quarter 1: $n_1=520$, defectives $x_1=57$.
		Quarter 2: $n_2=610$, defectives $x_2=52$.
		\begin{enumerate}
			\item Construct 95\% and 99\% confidence intervals for each quarter proportion, and report the margin of error for each interval (C7).
		\end{enumerate}

		\subsection*{C7}
		\textbf{Quarter 1}
		\[
		\hat{p}_1=\frac{57}{520}=0.1096,
		\quad SE_1=\sqrt{\frac{0.1096(0.8904)}{520}}=0.0137
		\]
		\[
		E_{95}=1.96(0.0137)=0.0269,\quad CI_{95}=[0.0828,0.1365]
		\]
		\[
		E_{99}=2.576(0.0137)=0.0353,\quad CI_{99}=[0.0743,0.1449]
		\]

		\textbf{Quarter 2}
		\[
		\hat{p}_2=\frac{52}{610}=0.0852,
		\quad SE_2=\sqrt{\frac{0.0852(0.9148)}{610}}=0.0113
		\]
		\[
		E_{95}=1.96(0.0113)=0.0222,\quad CI_{95}=[0.0631,0.1074]
		\]
		\[
		E_{99}=2.576(0.0113)=0.0291,\quad CI_{99}=[0.0561,0.1144]
		\]

		\newpage
		\item
		\textbf{Problem description}
		A national energy agency studies the proportion of households that adopted rooftop solar systems in two districts during the same year.
		District North: $n_1=480$, adopters $x_1=168$.
		District South: $n_2=540$, adopters $x_2=221$.
		\begin{enumerate}
			\item Construct 90\%, 95\%, and 99\% confidence intervals for each district proportion, and report the margin of error for each interval (C7).
		\end{enumerate}

		\subsection*{C7}
		\textbf{District North}
		\[
		\hat{p}_N=\frac{168}{480}=0.3500,
		\quad SE_N=\sqrt{\frac{0.35(0.65)}{480}}=0.0218
		\]
		\[
		E_{90}=1.645(0.0218)=0.0358,\quad CI_{90}=[0.3142,0.3858]
		\]
		\[
		E_{95}=1.96(0.0218)=0.0427,\quad CI_{95}=[0.3073,0.3927]
		\]
		\[
		E_{99}=2.576(0.0218)=0.0561,\quad CI_{99}=[0.2939,0.4061]
		\]

		\textbf{District South}
		\[
		\hat{p}_S=\frac{221}{540}=0.4093,
		\quad SE_S=\sqrt{\frac{0.4093(0.5907)}{540}}=0.0212
		\]
		\[
		E_{90}=1.645(0.0212)=0.0348,\quad CI_{90}=[0.3745,0.4441]
		\]
		\[
		E_{95}=1.96(0.0212)=0.0415,\quad CI_{95}=[0.3678,0.4507]
		\]
		\[
		E_{99}=2.576(0.0212)=0.0545,\quad CI_{99}=[0.3548,0.4638]
		\]

	\end{ExamProblems}

\end{document}
