\input{../../preamble/preamble_exams.tex}
\begin{document}
	\ExamTitleBlock{11th grade - Activity Criteria 6}{}{}
	\vspace{-1.5cm}

	\begin{itemize}
		\item C6: Determine the required sample size to achieve a target margin of error.
		\item The criterion is evaluated across the 6 problems. It is considered passed when it is correctly activated in at least 4 of the 6 problems.
	\end{itemize}
		

	\begin{enumerate}
		\item \textbf{Sample Size Determination for Estimating a Mean Commute Time} \\
		A transportation planner wants to estimate the mean commute time (in minutes) for a metropolitan rail line. Historical records from previous years report a long-run average commute time of $\mu = 42$ minutes. Past studies show the population standard deviation is $\sigma = 8$ minutes. The planner wants a 99\% confidence estimate with a maximum error of $E = 2$ minutes. Determine the minimum sample size needed.
		
		\item \textbf{Sample Size Determination for Estimating a Mean Weekly Wage} \\
		A labor analyst wants to estimate the mean weekly wage (in dollars) for a specific industry. Long-term administrative records for this industry report an average weekly wage of $\mu = 820$ dollars. The population standard deviation is known to be $\sigma = 15$ dollars. The analyst needs a 95\% confidence estimate with a maximum error of $E = 3$ dollars. Determine the minimum sample size required.
		
		\item
		\textbf{Sample Size Determination for Estimating Mean Crop Yield} \\
		An agriculture board wants to estimate the mean yield (in tons per hectare) of a crop. Historical production reports for the region indicate a long-run average yield of $\mu = 6.8$ tons per hectare. The population standard deviation is $\sigma = 4.5$ tons per hectare. The board wants a 90\% confidence estimate with a maximum error of $E = 1$ ton per hectare. Determine the minimum sample size required.
		
		\item
		\textbf{Additional Sample Size Needed to Reduce Margin of Error for a Mean} \\
		A utilities department wants to estimate the mean monthly electricity bill (in dollars). The population standard deviation is $\sigma = 5.2$ dollars, and $n_{\text{current}} = 49$ bills have been recorded. From these 49 bills, a sample mean of $\bar{x} = 120.5$ dollars and a sample standard deviation of $s = 5.0$ dollars were obtained. How many additional observations are required to reduce the margin of error to $E = 1.0$ dollar at 95\% confidence?
		
		\item
		\textbf{Additional Sample Size Needed to Reduce Margin of Error for a Mean Exam Score} \\
		A school district is estimating the mean final exam score for a course. The population standard deviation is $\sigma = 6$ points, and $n_{\text{current}} = 25$ scores are available. From these scores, a sample mean of $\bar{x} = 78.2$ points and a sample standard deviation of $s = 5.8$ points were obtained. How many additional observations are required to reduce the margin of error to $E = 1.2$ points at 90\% confidence?
		
		\item
		\textbf{Additional Sample Size Needed to Reduce Margin of Error for Mean Water Use} \\
		A water authority is estimating the mean daily household water use (in liters). The population standard deviation is $\sigma = 18$ liters, and $n_{\text{current}} = 64$ households have been surveyed. From these households, a sample mean of $\bar{x} = 210$ liters and a sample standard deviation of $s = 17.2$ liters were obtained. How many additional observations are required to reduce the margin of error to $E = 4$ liters at 99\% confidence?
		
	\end{enumerate}
\end{document}
